---
subtitle: 'COMBINED | Cheat Sheet'
---

\newpage

# Combined Study Data {#sec-combined}

```{r}
#| label: SETUP
#| warning : false
#| message : false

library(Hmisc) # %nin% operator
library(janitor) #compare df cols 
library(mosaic) #simple descriptives [favstats]
library(multimode) #test for multimodality
library(fitdistrplus) #fitting distributions
library(performance) #multimodality
library(kableExtra) #printing tables 
library(vcd) #mosaicplots
library(ggpubr) #arrange plots
library(ggstatsplot) #sanity checks 
library(sjPlot) #model plots
library(ggeasy) #easy plot changes
library(statsExpressions) #easy plot changes

#modelling
library(lme4)
library(ordinal) #ordinal and mixed ordinal models
library(nnet) #multinomial regression
library(jtools) # misc helpers 
library(brms) #bayesian regression models

#all the things
library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

#SET THEME
theme_set(theme_minimal())
```

## SETUP

```{r}
#| label: IMPORT-SUBJECTS
#| warning : false
#| message : false

# HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
# mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
# setwd(mbp)

##:::::::: IMPORT SUBJECT DATA 

in_sgc3a <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds") %>% mutate(
  study = "SGC3A"
) #%>% dplyr::select(-absolute_score)

in_sgc3b <- read_rds("analysis/SGC3B/data/2-scored-data/sgc3b_scored_participants.rds") %>% mutate(
  study = "SGC3B"
) #%>% dplyr::select(-absolute_score)

in_sgc4a <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_participants.rds") %>% mutate(
  study = "SGC4A",
  pretty_mode = "online-replication"
)
in_sgc4b <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_participants.rds") %>% mutate(
  study = "SGC4B",
  pretty_mode = "online-replication"
)

# TODO TEMPORARY ONLY REPLACE WITH SCORED DATA 
in_sgc4c <- read_rds("analysis/SGC4C/data/0-session-level/sgc4c_participants.rds") %>% mutate(
  pretty_mode = "online-replication"
)

in_sgc5 <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5_scored_participants.rds") %>% mutate(
  study = "SGC5A",
  pretty_mode = "online-replication"
) %>% dplyr::select(-absolute_score)

#COMPARE DF COLUMNS
# janitor::compare_df_cols(sgc3a, sgc4c)
#SGC5 HAS ABSOLUTE SCORE BUT THEY SHOULDN'T 

##:::::::: IMPORT ITEM DATA 

in_sgc3a_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(study = "SGC3A")
  
in_sgc3b_items <- read_rds("analysis/SGC3B/data/2-scored-data/sgc3b_scored_items.rds") %>% mutate(study = "SGC3B")
  
in_sgc4a_items <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_items.rds") %>% mutate(study = "SGC4A")
  
in_sgc4b_items <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_items.rds") %>% mutate(study = "SGC4B")

# TODO TEMPORARY ONLY REPLACE WITH SCORED DATA 
in_sgc4c_items <- read_rds("analysis/SGC4C/data/0-session-level/sgc4c_items.rds") %>% mutate(study = "SGC4C")
  
in_sgc5_items <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5a_scored_items.rds") %>% mutate(study = "SGC5")
  
##:::::::: IMPORT OSPAN DATA 
ospan <- read_csv("analysis/combined/data/fall21_scored_ospan.csv") %>% mutate(
  subject = SUBJECTID
)

# MERGE #TODO ADD IN SGC4C
df_all <- rbind(in_sgc3a, in_sgc3b, in_sgc4a, in_sgc4b, in_sgc5) %>%  dplyr::select(subject,term,study,condition,pretty_condition)
```

### Conditions per Study

```{r}
paste("SG3A")
addmargins(table(in_sgc3a$term, in_sgc3a$pretty_condition))

paste("SG3B")
addmargins(table(in_sgc3b$term, in_sgc3b$pretty_condition))

paste("SG4A")
addmargins(table(in_sgc4a$term, in_sgc4a$pretty_condition))

paste("SG4B")
addmargins(table(in_sgc4b$term, in_sgc4b$pretty_condition))

paste("SG4C")
addmargins(table(in_sgc4c$term, in_sgc4c$pretty_condition))

paste("SG5")
addmargins(table(in_sgc5$term, in_sgc5$pretty_condition))

paste("ALL")
addmargins(table(df_all$study, df_all$condition))

```

### Are Subjects Unique?

```{r}
#CHECK FOR DUPLICATES 
n_total <- nrow(df_all)
paste(n_total, "subjects records imported ")
n_dist <- nrow(df_all %>% distinct(subject))
paste(n_dist, "unique subjects")
subs <- df_all$subject
dups <- subs[duplicated(subs)] 
n_dups <- length(dups)
paste(n_dups, "duplicated subjects")
```

### De-Duplicate?

#### SGC3A 
```{r}
# table(in_sgc3a$term, in_sgc3a$condition)
# 
# #REPORT sgc3a ONLINE REPLICATION [fall21, winter22]
# # sub_df <- sgc3a %>% filter(term %in% c("fall21"))
# sub_df <- in_sgc3a %>% filter(mode == "asynch")
# item_df <- in_sgc3a_items %>% filter(subject %in% sub_df$subject) %>% filter(q %nin% c(6,9)) %>% mutate(
#   state = recode_factor(score_SCALED, #for ordinal
#                          "-1" = "orth-like",
#                          "-0.5" = "unknown",
#                          "0" = "unknown",
#                          "0.5" = "tri-like",
#                          "1" = "tri-like"),
#   state = as.ordered(state),
#   score_SCALED = as.ordered(score_SCALED)
# )
# nrow(item_df) == nrow(sub_df) * 13
# table(sub_df$condition)
# 
# sgc3a <- sub_df
# sgc3a_items <- item_df
# 
# table(sgc3a$term, sgc3a$condition) %>% addmargins()

```


#### SGC4A 
```{r}

table(in_sgc4a$term, in_sgc4a$condition)

#sgc4a as ONLINE ONLY [winter22]
#discard 17,18 as pilot?

sub_df <- in_sgc4a %>% filter(term == "winter22")
item_df <- in_sgc4a_items %>% filter(subject %in% sub_df$subject) %>% filter(q %nin% c(6,9)) %>% mutate(
  state = recode_factor(score_SCALED, #for ordinal
                         "-1" = "orth-like",
                         "-0.5" = "unknown",
                         "0" = "unknown",
                         "0.5" = "tri-like",
                         "1" = "tri-like"),
  state = as.ordered(state),
  score_SCALED = as.ordered(score_SCALED)
)
nrow(item_df) == nrow(sub_df) * 13
table(sub_df$condition)

sgc4a_A <- sub_df %>% filter(condition %in% c("111","113")) %>% droplevels()
sgc4a_A_items <- item_df %>% filter(subject %in% sgc4a_A$subject) %>% droplevels()

sgc4a_B <- sub_df %>% filter(condition %in% c("111","114","115")) %>% droplevels()
sgc4a_B_items <- item_df %>% filter(subject %in% sgc4a_B$subject) %>% droplevels()

table(sgc4a_A$term, sgc4a_A$condition) %>% addmargins()
table(sgc4a_B$term, sgc4a_B$condition) %>% addmargins()
```


### EVALUATE DE-DUP

```{r}
# new_df_subjects <- rbind(new_3a,new_3a_online,new_3b,new_4a,new_4b,new_5a)
# 
# #CHECK FOR DUPLICATES 
# x_total <- nrow(new_df_subjects)
# paste(x_total, "subjects should be unique")
# x_dist <- nrow(new_df_subjects %>% distinct(subject))
# paste(x_dist, "unique subjects")
# x_subs <- new_df_subjects$subject
# x_dups <- x_subs[duplicated(x_subs)] 
# xn_dups <- length(x_dups)
# paste(xn_dups, "duplicated subjects")
# 
# pilot_3b = 15+29
# pilot_3b3 = 5+12+3+11
# n_discard = 6+52+5+12+3+11 + 15+29 #hardcoded number of discards 
# #1197 unique subjects  + 327 [de-duplicated] + (sgc3b,4a pilots) = 
# 
# #number of distinct in original == number of distinct now + n discarded 
# n_dist ==  x_dist + n_discard 

```

## SGC3A _ IN PERSON

```{r}

#SETUP DATA HERE 
sub_df <-  in_sgc3a %>% filter(mode == "lab-synch") 
item_df <- in_sgc3a_items %>% filter(subject %in% sub_df$subject) %>% 
  mutate (
    state = recode_factor(score_SCALED, #for ordinal
                         "-1" = "orth-like",
                         "-0.5" = "unknown",
                         "0" = "unknown",
                         "0.5" = "tri-like",
                         "1" = "tri-like"),
    state = as.ordered(state))


q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))
```


#### Describe
```{r}
table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)
```

```{r}
#::::::::::::SUBJECT LEVEL
ab.dist <- gf_density(~s_NABS, fill= ~condition, data = sub_df) %>%
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Absolute Score")

result <- two_sample_test(data = sub_df, x = condition, y = s_NABS,
                          type = "nonparametric", var.equal = FALSE,
                          alternative = "less",
                          k = 2L, conf.level = 0.89, effsize.type = "g",
                          bf.prior = 0.707, tr = 0.2, nboot = 100L)

ab.box <- ggbetweenstats(data = sub_df, x = condition, y = s_NABS,
               type = "nonparametric", var.equal = FALSE, 
               results.subtitle = FALSE,
               subtitle = result$expression[[1]],
               title = "Total Absolute Score [one-tailed]")

scal.dist <- gf_density(~s_SCALED, fill= ~condition, data = sub_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Scaled Score")

result <- two_sample_test(data = sub_df, x = condition, y = s_SCALED,
                          type = "nonparametric", var.equal = FALSE,
                          alternative = "less",
                          k = 2L, conf.level = 0.89, effsize.type = "g",
                          bf.prior = 0.707, tr = 0.2, nboot = 100L)

scal.box <- ggbetweenstats(data = sub_df, x = condition, y = s_SCALED,
               type = "nonparametric", var.equal = FALSE,
               results.subtitle = FALSE,
               subtitle = result$expression[[1]],
               title = "Total Scaled Score [one-tailed]")

stat.dist <- gf_props(~state, fill= ~condition, data = item_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total States")

stat.box <- ggbarstats(data = item_df, x = state, y = condition,
               results.subtitle = FALSE, label = "percentage",
               title = "Total States")

#::::::::::::Q1 
# q1ab.box <- 
  ggbarstats(data = item_df %>% filter(q==1), x = condition, y = score_niceABS,
               # results.subtitle = FALSE,
               title = "Q1 Absolute Score")

q1scal.box <- ggbarstats(data = item_df %>% filter(q==1), x = score_SCALED, y = condition,
               results.subtitle = FALSE,
               title = "Q1 Scaled Score")

q1stat.box <- ggbarstats(data = item_df %>% filter(q==1), x = state, y = condition,
               results.subtitle = FALSE,
               title = "Q1 State")

#Subject Level Totals
cowplot::plot_grid(ab.dist, scal.dist, stat.dist, ab.box, scal.box, stat.box,  nrow = 2)

#Q1 Totals
cowplot::plot_grid(q1ab.box, q1scal.box, q1stat.box,  nrow = 1)

```



#### NON PARAMETRIC T TESTS

```{r}

#::::::::::::NONPARAMETRIC TESTS

# paste("Q1 Absolute Score")
# #WILCOXON RANK SUM TEST  [non parametric t-test]
# wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
#             alternative = "less")
# 
# paste("Q1 Scaled Score")
# wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
#             alternative = "less")


#WILCOXON RANK SUM TEST  [non parametric t-test]
paste("Test Phase Absolute")
wilcox.test(item_test_NABS ~ pretty_condition, 
            data = sub_df,
            alternative = "less")

paste("Test Phase Scaled")
wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Absolute")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Scaled")
wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")
```


#### NON PARAMETRIC ANOVA

```{r}

#::::::::::::NONPARAMETRIC TESTS

# paste("Q1 Absolute Score")
# #KRUSKAL WALLIS [non parametric ANOVA]
# kruskal.test(formula = item_q1_NABS ~ pretty_condition, data = sub_df,
#             alternative = "less")
# 
# paste("Q1 Scaled Score")
# wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
#             alternative = "less")


# #WILCOXON RANK SUM TEST  [non parametric t-test]
# paste("Test Phase Absolute")
# kruskal.test(item_test_NABS ~ pretty_condition,
#             data = sub_df)
# 
# #POSTHOC
# # wilcox.test(item_test_NABS ~ pretty_condition, 
# #             data = sub_df,
# #             alternative = "less")
# 
# paste("Test Phase Scaled")
# kruskal.test(item_test_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Absolute")
# kruskal.test(s_NABS ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Scaled")
# kruskal.test(s_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
```


#### Logistic Regression on ABSOLUTE SCORE
```{r}
#::::::::::::LOGISTIC REGRESSION MODELS

paste("Logistic regression of q1 score")
q1.lr <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(q1.lr)
plot_model(q1.lr)

paste("MIXED Logistic regression of test phase score")
test.ml <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(test.ml)
plot_model(test.ml)

paste("MIXED Logistic regression of all items score")
all.ml <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(all.ml)
plot_model(all.ml)
```

#### Ordinal Regression on SCALED SCORE
```{r}

#::::::::::::ORDINAL REGRESSION MODELS

paste("Ordinal regression of q1 score")
q1.om <- clm(as.ordered(score_SCALED) ~ condition, data = q1)
summary(q1.om)
(ctable <- coef(summary(q1.om)))
(ci <- confint(q1.om)) 
sjPlot::plot_model(q1.om, type = "eff")
sjPlot::plot_model(q1.om)

paste("MIXED Ordinal regression of test phase score")
test.mom <- clmm(score_SCALED ~ condition + (1|subject), data = test)
summary(test.mom)
(ctable <- coef(summary(test.mom)))
(ci <- confint(test.mom)) 
sjPlot::plot_model(test.mom, type = "eff")
sjPlot::plot_model(test.mom)

paste("MIXED Ordinal regression of all phase score")
all.mom <- clmm(score_SCALED ~ condition + (1|subject), data = all)
summary(all.mom)
(ctable <- coef(summary(all.mom)))
(ci <- confint(all.mom)) 
sjPlot::plot_model(all.mom, type = "pred")
sjPlot::plot_model(all.mom)

```


#### Ordinal Regression on STATE
```{r}

#::::::::::::ORDINAL REGRESSION MODELS

paste("Ordinal regression of q1 state")
q1.om <- clm(state ~ condition, data = q1)
summary(q1.om)
(ctable <- coef(summary(q1.om)))
(ci <- confint(q1.om)) 
sjPlot::plot_model(q1.om, type = "eff")
sjPlot::plot_model(q1.om)

paste("MIXED Ordinal regression of test phase state")
test.mom <- clmm(state ~ condition + (1|subject), data = test)
summary(test.mom)
(ctable <- coef(summary(test.mom)))
(ci <- confint(test.mom)) 
sjPlot::plot_model(test.mom, type = "eff")
sjPlot::plot_model(test.mom)

paste("MIXED Ordinal regression of all phase stae")
all.mom <- clmm(state ~ condition + (1|subject), data = all)
summary(all.mom)
(ctable <- coef(summary(all.mom)))
(ci <- confint(all.mom)) 
sjPlot::plot_model(all.mom, type = "pred")
sjPlot::plot_model(all.mom)

```

#### Multinomial Regression on Q1 STATE
```{r}

#::::::::::::MULTINOMIAL REGRESSION MODELS
paste("Multinomial regression of q1 state")
q1.catm <- multinom(state ~ condition, data = q1)
summary(q1.catm)
(ctable <- coef(summary(q1.catm)))
(ci <- confint(q1.catm)) 
sjPlot::plot_model(q1.catm, type = "eff")
sjPlot::plot_model(q1.catm)

paste("TODO MIXED Multinomial regression of test phase STATE")
### BAYESIAN ###
paste("TODO MIXED Multinomial regression of all phase STATE")

```





## MODEL SGC3A _ ONLINE REPLICATION

```{r}

#STUDY LEVEL DATAFRAMES
# sgc3a, sgc3a_items
# sgc4a_A, sgc4a_A_items
# sgc4a_B, sgc4a_B_items

#SETUP DATA HERE 
sub_df <-  sgc3a #set dataframe
item_df <- sgc3a_items # set dataframe
  
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))
```


#### Describe
```{r}
table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)
```

```{r}
#::::::::::::SUBJECT LEVEL
ab.dist <- gf_density(~s_NABS, fill= ~condition, data = sub_df) %>%
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Absolute Score")

ab.box <- ggbetweenstats(data = sub_df, x = condition, y = s_NABS,
               type = "nonparametric", var.equal = FALSE, 
               title = "Total Absolute Score")

scal.dist <- gf_density(~s_SCALED, fill= ~condition, data = sub_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Scaled Score")

scal.box <- ggbetweenstats(data = sub_df, x = condition, y = s_SCALED,
               type = "nonparametric", var.equal = FALSE,
               title = "Total Scaled Score")

stat.dist <- gf_props(~state, fill= ~condition, data = item_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total States")

stat.box <- ggbarstats(data = item_df, x = state, y = condition,
               results.subtitle = FALSE, label = "percentage",
               title = "Total States")


#::::::::::::Q1 
q1ab.box <- ggbarstats(data = item_df %>% filter(q==1), x = score_niceABS, y = condition,
               results.subtitle = FALSE,
               title = "Q1 Absolute Score")

q1scal.box <- ggbarstats(data = item_df %>% filter(q==1), x = score_SCALED, y = condition,
               results.subtitle = FALSE,
               title = "Q1 Scaled Score")

q1stat.box <- ggbarstats(data = item_df %>% filter(q==1), x = state, y = condition,
               results.subtitle = FALSE,
               title = "Q1 State")

#Subject Level Totals
cowplot::plot_grid(ab.dist, scal.dist, stat.dist, ab.box, scal.box, stat.box,  nrow = 2)

#Q1 Totals
cowplot::plot_grid(q1ab.box, q1scal.box, q1stat.box,  nrow = 1)

```



#### NON PARAMETRIC T TESTS

```{r}

#::::::::::::NONPARAMETRIC TESTS

# paste("Q1 Absolute Score")
# #WILCOXON RANK SUM TEST  [non parametric t-test]
# wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
#             alternative = "less")
# 
# paste("Q1 Scaled Score")
# wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
#             alternative = "less")


#WILCOXON RANK SUM TEST  [non parametric t-test]
paste("Test Phase Absolute")
wilcox.test(item_test_NABS ~ pretty_condition, 
            data = sub_df,
            alternative = "less")

paste("Test Phase Scaled")
wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Absolute")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Scaled")
wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")
```


#### NON PARAMETRIC ANOVA

```{r}

#::::::::::::NONPARAMETRIC TESTS

# paste("Q1 Absolute Score")
# #KRUSKAL WALLIS [non parametric ANOVA]
# kruskal.test(formula = item_q1_NABS ~ pretty_condition, data = sub_df,
#             alternative = "less")
# 
# paste("Q1 Scaled Score")
# wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
#             alternative = "less")


# #WILCOXON RANK SUM TEST  [non parametric t-test]
# paste("Test Phase Absolute")
# kruskal.test(item_test_NABS ~ pretty_condition,
#             data = sub_df)
# 
# #POSTHOC
# # wilcox.test(item_test_NABS ~ pretty_condition, 
# #             data = sub_df,
# #             alternative = "less")
# 
# paste("Test Phase Scaled")
# kruskal.test(item_test_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Absolute")
# kruskal.test(s_NABS ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Scaled")
# kruskal.test(s_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
```


#### Logistic Regression on ABSOLUTE SCORE
```{r}
#::::::::::::LOGISTIC REGRESSION MODELS

paste("Logistic regression of q1 score")
q1.lr <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(q1.lr)
plot_model(q1.lr)

paste("MIXED Logistic regression of test phase score")
test.ml <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(test.ml)
plot_model(test.ml)

paste("MIXED Logistic regression of all items score")
all.ml <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(all.ml)
plot_model(all.ml)
```

#### Ordinal Regression on SCALED SCORE
```{r}

#::::::::::::ORDINAL REGRESSION MODELS

paste("Ordinal regression of q1 score")
q1.om <- clm(as.ordered(score_SCALED) ~ condition, data = q1)
summary(q1.om)
(ctable <- coef(summary(q1.om)))
(ci <- confint(q1.om)) 
sjPlot::plot_model(q1.om, type = "eff")
sjPlot::plot_model(q1.om)

paste("MIXED Ordinal regression of test phase score")
test.mom <- clmm(score_SCALED ~ condition + (1|subject), data = test)
summary(test.mom)
(ctable <- coef(summary(test.mom)))
(ci <- confint(test.mom)) 
sjPlot::plot_model(test.mom, type = "eff")
sjPlot::plot_model(test.mom)

paste("MIXED Ordinal regression of all phase score")
all.mom <- clmm(score_SCALED ~ condition + (1|subject), data = all)
summary(all.mom)
(ctable <- coef(summary(all.mom)))
(ci <- confint(all.mom)) 
sjPlot::plot_model(all.mom, type = "pred")
sjPlot::plot_model(all.mom)

```


#### Ordinal Regression on STATE
```{r}

#::::::::::::ORDINAL REGRESSION MODELS

paste("Ordinal regression of q1 state")
q1.om <- clm(state ~ condition, data = q1)
summary(q1.om)
(ctable <- coef(summary(q1.om)))
(ci <- confint(q1.om)) 
sjPlot::plot_model(q1.om, type = "eff")
sjPlot::plot_model(q1.om)

paste("MIXED Ordinal regression of test phase state")
test.mom <- clmm(state ~ condition + (1|subject), data = test)
summary(test.mom)
(ctable <- coef(summary(test.mom)))
(ci <- confint(test.mom)) 
sjPlot::plot_model(test.mom, type = "eff")
sjPlot::plot_model(test.mom)

paste("MIXED Ordinal regression of all phase stae")
all.mom <- clmm(state ~ condition + (1|subject), data = all)
summary(all.mom)
(ctable <- coef(summary(all.mom)))
(ci <- confint(all.mom)) 
sjPlot::plot_model(all.mom, type = "pred")
sjPlot::plot_model(all.mom)

```

#### Multinomial Regression on Q1 STATE
```{r}

#::::::::::::MULTINOMIAL REGRESSION MODELS
paste("Multinomial regression of q1 state")
q1.catm <- multinom(state ~ condition, data = q1)
summary(q1.catm)
(ctable <- coef(summary(q1.catm)))
(ci <- confint(q1.catm)) 
sjPlot::plot_model(q1.catm, type = "eff")
sjPlot::plot_model(q1.catm)

paste("TODO MIXED Multinomial regression of test phase STATE")
### BAYESIAN ###
paste("TODO MIXED Multinomial regression of all phase STATE")

```


## MODEL SGC4A _ TRIANGULAR OPTIMIZATION

```{r}

#STUDY LEVEL DATAFRAMES
# sgc3a, sgc3a_items
# sgc4a_A, sgc4a_A_items
# sgc4a_B, sgc4a_B_items

#SETUP DATA HERE 
sub_df <-  sgc4a_A #set dataframe
item_df <- sgc4a_A_items # set dataframe
  
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))
```


#### Describe
```{r}
table(sub_df$condition)
favstats(s_NABS ~ condition, data = sub_df)
favstats(s_SCALED ~ condition, data = sub_df)
```

```{r}
#::::::::::::SUBJECT LEVEL
ab.dist <- gf_density(~s_NABS, fill= ~condition, data = sub_df) %>%
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Absolute Score")

ab.box <- ggbetweenstats(data = sub_df, x = condition, y = s_NABS,
               type = "nonparametric", var.equal = FALSE, 
               title = "Total Absolute Score")

scal.dist <- gf_density(~s_SCALED, fill= ~condition, data = sub_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total Scaled Score")

scal.box <- ggbetweenstats(data = sub_df, x = condition, y = s_SCALED,
               type = "nonparametric", var.equal = FALSE,
               title = "Total Scaled Score")

stat.dist <- gf_props(~state, fill= ~condition, data = item_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend() + 
  labs(title = "Total States")

stat.box <- ggbarstats(data = item_df, x = state, y = condition,
               results.subtitle = FALSE, label = "percentage",
               title = "Total States")


#::::::::::::Q1 
q1ab.box <- ggbarstats(data = item_df %>% filter(q==1), x = score_niceABS, y = condition,
               results.subtitle = FALSE,
               title = "Q1 Absolute Score")

q1scal.box <- ggbarstats(data = item_df %>% filter(q==1), x = score_SCALED, y = condition,
               results.subtitle = FALSE,
               title = "Q1 Scaled Score")

q1stat.box <- ggbarstats(data = item_df %>% filter(q==1), x = state, y = condition,
               results.subtitle = FALSE,
               title = "Q1 State")

#Subject Level Totals
cowplot::plot_grid(ab.dist, scal.dist, stat.dist, ab.box, scal.box, stat.box,  nrow = 2)

#Q1 Totals
cowplot::plot_grid(q1ab.box, q1scal.box, q1stat.box,  nrow = 1)

```



#### NON PARAMETRIC T TESTS

```{r}

#::::::::::::NONPARAMETRIC TESTS

paste("Q1 Absolute Score")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("Q1 Scaled Score")
wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Absolute")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("All Task Scaled")
wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")
```


#### NON PARAMETRIC ANOVA

```{r}

#::::::::::::NONPARAMETRIC TESTS

# paste("Q1 Absolute Score")
# #KRUSKAL WALLIS [non parametric ANOVA]
# kruskal.test(formula = item_q1_NABS ~ pretty_condition, data = sub_df,
#             alternative = "less")
# 
# paste("Q1 Scaled Score")
# wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
#             alternative = "less")


# #WILCOXON RANK SUM TEST  [non parametric t-test]
# paste("Test Phase Absolute")
# kruskal.test(item_test_NABS ~ pretty_condition,
#             data = sub_df)
# 
# #POSTHOC
# # wilcox.test(item_test_NABS ~ pretty_condition, 
# #             data = sub_df,
# #             alternative = "less")
# 
# paste("Test Phase Scaled")
# kruskal.test(item_test_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Absolute")
# kruskal.test(s_NABS ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
# #             alternative = "less")
# 
# paste("All Task Scaled")
# kruskal.test(s_SCALED ~ pretty_condition,
#             data = sub_df)
# #POSTHOC
# # wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
# #             alternative = "less")
```


#### Logistic Regression on Q1 ABSOLUTE SCORE
```{r}
#::::::::::::LOGISTIC REGRESSION MODELS

paste("Logistic regression of q1 score")
q1.lr <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summary(q1.lr)
plot_model(q1.lr)

paste("MIXED Logistic regression of all items score")
all.ml <- glmer( score_niceABS ~ condition + (1|subject) + (1|q), data = all, family = "binomial")
summary(all.ml)
plot_model(all.ml)
```

#### Ordinal Regression on SCALED SCORE
```{r}
# 
# #::::::::::::ORDINAL REGRESSION MODELS
# 
# paste("Ordinal regression of q1 score")
# q1.om <- clm(as.ordered(score_SCALED) ~ condition, data = q1)
# summary(q1.om)
# (ctable <- coef(summary(q1.om)))
# (ci <- confint(q1.om)) 
# sjPlot::plot_model(q1.om, type = "eff")
# sjPlot::plot_model(q1.om)
# 
# paste("MIXED Ordinal regression of test phase score")
# test.mom <- clmm(score_SCALED ~ condition + (1|subject), data = test)
# summary(test.mom)
# (ctable <- coef(summary(test.mom)))
# (ci <- confint(test.mom)) 
# sjPlot::plot_model(test.mom, type = "eff")
# sjPlot::plot_model(test.mom)
# 
# paste("MIXED Ordinal regression of all phase score")
# all.mom <- clmm(score_SCALED ~ condition + (1|subject), data = all)
# summary(all.mom)
# (ctable <- coef(summary(all.mom)))
# (ci <- confint(all.mom)) 
# sjPlot::plot_model(all.mom, type = "pred")
# sjPlot::plot_model(all.mom)

```


#### Ordinal Regression on STATE
```{r}

# #::::::::::::ORDINAL REGRESSION MODELS
# 
# paste("Ordinal regression of q1 state")
# q1.om <- clm(state ~ condition, data = q1)
# summary(q1.om)
# (ctable <- coef(summary(q1.om)))
# (ci <- confint(q1.om)) 
# sjPlot::plot_model(q1.om, type = "eff")
# sjPlot::plot_model(q1.om)
# 
# paste("MIXED Ordinal regression of test phase state")
# test.mom <- clmm(state ~ condition + (1|subject), data = test)
# summary(test.mom)
# (ctable <- coef(summary(test.mom)))
# (ci <- confint(test.mom)) 
# sjPlot::plot_model(test.mom, type = "eff")
# sjPlot::plot_model(test.mom)
# 
# paste("MIXED Ordinal regression of all phase stae")
# all.mom <- clmm(state ~ condition + (1|subject), data = all)
# summary(all.mom)
# (ctable <- coef(summary(all.mom)))
# (ci <- confint(all.mom)) 
# sjPlot::plot_model(all.mom, type = "pred")
# sjPlot::plot_model(all.mom)

```

#### Multinomial Regression on Q1 STATE
```{r}

#::::::::::::MULTINOMIAL REGRESSION MODELS
paste("Multinomial regression of q1 state")
q1.catm <- multinom(state ~ condition, data = q1)
summary(q1.catm)
(ctable <- coef(summary(q1.catm)))
(ci <- confint(q1.catm)) 
sjPlot::plot_model(q1.catm, type = "eff")
sjPlot::plot_model(q1.catm)

paste("TODO MIXED Multinomial regression of test phase STATE")
### BAYESIAN ###
paste("TODO MIXED Multinomial regression of all phase STATE")

```



## WIP

### OSPAN SETUP

### Find OSPAN subjects

```{r}
# 
# #OSPAN RECORDS
# n_ospan <- nrow(ospan)
# paste(n_ospan, "scored OSPAN records")
# 
# #MERGE OSPAN DATA WITH SGC DATA 
# ospan_subjects <- new_df_subjects %>% filter(
#   subject %in% ospan$subject
# )
# n_found <- nrow(ospan_subjects)
# paste(n_found, "records found")
# n_missing <- n_ospan - n_found
# paste(n_missing, "ospan records don't have subjects")
# #where did the others go? maybe no clean records?
# paste("Some of these subjects failed attention check. Others were part of SGC3B online pilot")
# 
# #GET MERGED OSPAN
# df_ospan <- merge(ospan_subjects, ospan)
# #CONFIRM ALL DISTINCT 
# df_ospan %>% distinct(subject) %>% nrow() == nrow(df_ospan)
# 
# table(df_ospan$condition)
# 
# 
# #HOW MANY OSPAN IN SGC3A?
# sgc3a_ospan <- sgc3a %>% filter(subject %in% subs_ospan$subject)
# paste(nrow(sgc3a_ospan)," subjects in SGC3A")
# 
# sgc3b_ospan <- sgc3b %>% filter(subject %in% subs_ospan$subject)
# paste(nrow(sgc3b_ospan)," subjects in SGC3B")
# 
# sgc4a_ospan <- sgc4a %>% filter(subject %in% subs_ospan$subject)
# paste(nrow(sgc4a_ospan)," subjects in SGC4A")
# 
# 

```

**Investigate subjects for SGC3A, SGC3B**

```{r}
# #
# #validate how many subjects per study
# table(df_subjects$study)
# 
# #COMBINE SGC3A AND SGC3B
# df3 <- rbind(sgc3a, sgc3b)
# n_total <- nrow(df3)
# paste(n_total, "subjects total SGC3A + SGC3B")
# n_dist <- nrow(df3 %>% distinct(subject))
# paste(n_dist, "unique subjects")
# subs <- df3$subject
# dups <- subs[duplicated(subs)] 
# n_dups <- length(dups)
# paste(n_dups, "duplicated subjects")
# 
# #DATAFRAME of DUPLICATES
# df_dups <- df3 %>% filter(subject %in% dups)
# #what studies?
# unique(df_dups$condition)
# 
# 
# #filter out sgc3a subjects from 3b study file
# df_3b <- sgc3b
# df_3b <- sgc3b %>% filter(mode == 'lab-synch')
# table(df_3b$condition)
# 
# 
# 
# #double check; only sgc3as onlines are not in sgc3b
# check <- sgc3a 
# 
# 
# #CHECK 
# #need to fill cells for SGC3B 
# 
# df_a <- sgc3a %>% filter(subject %nin% df_3b$subject)
# table(df_a$condition , df_a$mode)
# 
# table(sgc3a$condition, sgc3a$mode)
# 
# table(sgc3a$term)
# 
# 
# 
# 

```

```{r}

#LOOK AT DUPLICATES
# n_total <- nrow(df_subjects)
# paste(n_total, "subjects total")
# n_dist <- nrow(df_subjects %>% distinct(subject))
# paste(n_dist, "unique subjects")
# subs <- df_subjects$subject
# dups <- subs[duplicated(subs)] 
# n_dups <- length(dups)
# paste(n_dups, "duplicated subjects")
# 
# #DATAFRAME of DUPLICATES
# df_dups <- df_subjects %>% filter(subject %in% dups)
# #what studies?
# unique(df_dups$condition)
# table(df_dups$condition, df_dups$study)

#NEED 62 for SGC4A

```

```{r}

# df_3b <- df_3b %>% mutate(
#   IMPASSE = recode_factor(condition, 
#                            "111" = "none",
#                            "121" = "impasse",
#                            "211" = "none",
#                            "221" = "impasse",
#                            "311" = "none",
#                            "321" = "impasse"),
#    EXPLICIT = recode_factor(condition, 
#                            "111" = "none",
#                            "121" = "none",
#                            "211" = "img",
#                            "221" = "img",
#                            "311" = "ixn",
#                            "321" = "ixn")
# )
# 
# 
# gf_dhistogram(~item_test_NABS, data = df_3b) %>% 
#   gf_facet_grid(IMPASSE ~ EXPLICIT)
# 
# gf_boxplot(item_test_NABS ~ condition, data = df_3b) %>% 
#   gf_facet_grid(IMPASSE ~ EXPLICIT)
# 
# 
# 
# m <- lm(item_test_NABS ~ IMPASSE*EXPLICIT, data = df_3b)
# summ(m)
# check_model(m)
# 
# 
# 
# #KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
# kruskal.test(item_test_NABS ~ pretty_condition, data = df_3b)
# 
# #POSTHOC 
# pairwise.wilcox.test(df_3b$item_test_NABS, df_3b$pretty_condition,
#   p.adjust.method = "holm"
# )

```

```{r}
# 
# #MERGE OSPAN DATA WITH SGC DATA 
# df_ospan <- df_subjects %>% filter(
#   subject %in% ospan$subject
# )
# 
# df_ospan <- merge(df_ospan, ospan)
# 
# #RENAME for reuse
# df_all <- df_subjects
# # df_subjects <- df_ospan
# 
# #RECODE other study control conditions to CONTROL
# df <- df_all %>% mutate(
#   pretty_condition = recode_factor(pretty_condition, "Orth-Full" = "control",
#                                    "point" = "control")
# )
# 
# #CLEANUP
# # rm(ospan, sgc3a, sgc4a, sgc4b, sgc5, df_ospan)
# 
# temp <- df_all %>% dplyr::select(s_NABS, condition, study, subject) %>% rename(score = s_NABS)
# temp2 <- sgc4c %>% dplyr::select(absolute_score, condition, study, subject) %>% rename(score = "absolute_score")
# df <- rbind(temp, temp2)
```

```{r}
# table(df$study, df$condition)
# 
# paste("Subjects: ", nrow(df))
# paste("unique subjects ",nrow(unique(df$subject)))
# 
# #get list of unique subjects 
# u <- df %>% dplyr::select(subject) %>% distinct() 
# d <- df$subject %nin% u$subject
# 
# #unique subjects
# u <- unique(df$subject)
# #duplicated subjects
# d <- df$subject[duplicated(df$subject)]
# 
# length(d) 
# length(u) 
# length(d) + length(u) == nrow(df)
# 
# dups <- df %>% filter(subject %in% d)
# 
# table(dups$study, dups$condition)
# 
# #TODO ... NEED TO RUN 4A CONTROL 62
```


### Describe OSPAN

```{r}
#| label: DESC-PARTICIPANTS
# 
# title = "OSPAN Participants by Condition"
# cols = c("Condition","n")
# cont <- table(df_ospan$pretty_condition)
# cont %>%  addmargins() %>% kbl(caption = title, col.names = cols) %>% kable_classic()

```


```{r}
#| label: DESC-SUBJ-OSPAN
 
# title = "Descriptive Statistics of OSPAN Task Accuracy"
# ospan.stats <- rbind(
#   "MATH" = df_ospan %>% dplyr::select(math_acc) %>% unlist() %>% favstats(),
#   "ORDER" = df_ospan %>%  dplyr::select(order_acc) %>% unlist() %>% favstats(),
#   "WEIGHTED" = df_ospan %>% dplyr::select(weighted) %>% unlist() %>% favstats()
#   
# ) 
# ospan.stats %>% kbl (caption = title) %>% kable_classic() %>% 
#   footnote(general = "MATH = %correct of all math questions; 
#            ORDER = % correct of OSPAN ordering (out of 30); WEIGHTED = math*ospan number correct", general_title = "Note: ",footnote_as_chunk = T) 

```


```{r}
#| label: VIS-SUBJ-OSPAN

# #GGFORMULA | DENSITY HISTOGRAM 
#   gf_dhistogram(~weighted, data = df_ospan) + 
#   labs(x = "OSPAN (weighted) score",
#        y = "% of subjects",
#        title = "Distribution of OSPAN SCORE",
#        subtitle = "") 
# 
# ##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
# p <- gghistogram(df_ospan, x = "weighted", binwidth = 1,
#    add = "mean", rug = TRUE,
#    fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
#    add_density = TRUE)
# facet(p, facet.by=c("pretty_condition")) +
#   labs( title = "Distribution of OSPAN Score",
#         subtitle ="The Distribution of OSPAN scores is similar across conditions",
#         x = "OSPAN (weighted) score", y = "number of subjects") +
#   theme_minimal() + theme(legend.position = "blank")
# 
# 
# #PLOT EMPIRICIAL CUMULATIVE DISTRIBUTION FUNCTION
# ggplot(data = df_ospan, aes(weighted)) + 
#   stat_ecdf() + 
#   facet_wrap(~pretty_condition) + 
#   labs( title = "Empirical Cumulative Density Function — OSPAN",
#         x = "OSPAN (weighted) score", 
#         y = "Cumulative Probability") + theme_minimal()

```

```{r}

#IXN PLOT

# ospan.stats <- favstats(weighted, groups = condition)


```

#### NONPARAMETRIC TEST? nope no factorial anova

```{r}
# 
# #IMPORT SUBJECTS 
# df_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(
#   study = "SGC3A"
# ) #%>% dplyr::select(-absolute_score)
# 
# #FILTER GET ONLY neeed records
# 
# ospan_items <- df_items %>% filter(subject %in% df_ospan$subject)
# nrow(ospan_items)
# #VERIFY ITEMS ARE FOR SSPAN SUBJECTS
# nrow(ospan_items) == nrow(df_ospan) * 15
# 
# 
# #STANDARDIZE OSPAN
# df_ospan$z_ospan = zscore(df_ospan$weighted)
# histogram(df_ospan$z_ospan)
# 
# #OSPAN MEDIAN SPLIT
# med_ospan = median(df_ospan$weighted)
# df_ospan$low_ospan <- as.factor(df_ospan$weighted < med_ospan)
# 
# 
# #JOIN TO ospan to items 
# df_test <- left_join(df_ospan , ospan_items, by = 'subject') %>% 
#   dplyr::select(subject, condition.y, q, score_niceABS,  score_SCALED, low_ospan, z_ospan, weighted, math_acc, order_acc) %>% rename(condition = condition.y)
# 
# 
# items <- df_test
# q1 <- items %>% filter(q==1)
# test <- items %>% filter(q %nin% c(1,2,3,4,5,6,9))
# all <- items 
# 
# paste("Logistic regression of q1 score")
# m <- glm( score_niceABS ~ condition, data = q1, family = 'binomial')
# summ(m)
# 
# paste("Logistic regression of test phase score")
# m <- glmer( score_niceABS ~ condition*low_ospan+ (1|subject), data = test, family='binomial')
# summ(m)
# 
# paste("Logistic regression of all items score")
# m <- glmer( score_niceABS ~ condition*low_ospan+ (1|subject), data = all, family = "binomial")
# summ(m)

```



#### MODELS

```{r}
# 
# #CONDITION ONLY MODEL
# m1 <- lm( item_test_SCALED ~ condition, data = df_ospan)
# summary(m1)
# 
# #OSPAN ONLY MODEL
# m2 <- lm( item_test_SCALED ~ weighted, data = df_ospan)
# summary(m2)
# 
# #STANDARDIZE OSPAN
# df_ospan$z_ospan = zscore(df_ospan$weighted)
# histogram(df_ospan$z_ospan)
# 
# #OSPAN MEDIAN SPLIT
# med_ospan = median(df_ospan$weighted)
# df_ospan$low_ospan <- as.factor(df_ospan$weighted < med_ospan)
# 
# #STANDARDIZED OSPAN ONLY MODEL
# m3 <- lm( item_test_SCALED ~ low_ospan, data = df_ospan)
# summ(m3)
# 
# #MULTIPLE WITH STANDARDIZED OSPAN
# mr <- lm(item_test_SCALED ~ condition + low_ospan, data = df_ospan)
# summ(mr)
# anova(mr)
# 
# library(interactions)
# cat_plot(mr, item_test_SCALED, pred = condition, modx = low_ospan)
# df <- df_ospan %>% mutate(
#   condition = recode_factor(condition,
#                             "111"="111","121"="121")
# )
# library(ggstatsplot)
# grouped_ggbetweenstats(data = df,
#                        y = item_test_SCALED,
#                        x = condition,
#                        grouping.var = low_ospan)
# 
# #SINGLE WITH STANDARDIZED OSPAN
# mr2 <- lm(item_test_SCALED ~ condition + low_ospan, data = df_ospan)
# summary(mr2)
# 
# gf_histogram(~item_test_SCALED, data = df_ospan) %>% 
#   gf_facet_grid(low_ospan ~ condition, labeller = label_both)
# 
# 
# gf_jitter(item_test_SCALED ~ weighted, color = ~pretty_condition, data = df_ospan) %>% 
#   gf_facet_wrap(~pretty_condition)
# 
# gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = df_ospan) %>% 
#   gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

```

**OPSAN score does NOT predict performance in the control condition.**

```{r}
# 
# control <- df_ospan %>% filter(condition == "111")
# 
# gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = control) %>% 
#   gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)
# 
# #test performance predicted by OSPAN score
# m <- lm(item_test_NABS ~ weighted, data = control)
# summary(m)

```

**OPSAN score does NOT predict performance in the control condition.**

```{r}

# impasse <- df_ospan %>% filter(condition == "121")
# 
# gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = impasse) %>% 
#   gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)
# 
# #test performance predicted by OSPAN score
# m <- lm(item_test_NABS ~ low_ospan, data = impasse)
# summary(m)

```

**EXPLORE**

```{r}

# test <- df_ospan %>% filter(condition == "121")
# table(test$condition)
# 
# m <- lm(item_test_NABS ~ low_ospan, data = test)
# summ(m)
# 
# gf_boxplot(item_test_NABS ~ low_ospan, data = test)
```
#### describe ospan totals 
```{r}

#330 total 
# —————————
#126 lab 
#204 online  
  #  71 no ospan 
  # 133 ospan 
# sgc3a_lab <- df_all %>% filter(mode=="lab-synch") %>% filter(study=="SGC3A")
# sgc3a_online <- df_all %>% filter(mode=="asynch") %>% filter(study=="SGC3A")
# sgc3a_online_noospan <- sgc3a_online %>% filter(subject %nin% ospan$subject)
# sgc3a_onspan <- df_subjects
# 
# m <- lm( item_test_NABS ~ condition, data = sgc3a_online_noospan)
# summary(m)
# 
# m2 <- lm(item_test_NABS ~ condition, data = df_subjects)
# summary(m2)
# 
# 
# m3 <- lm(item_test_NABS ~ low_ospan, data = df_subjects)
# summary(m3)
# 


```

```{r}
# library(lpme)
# m <- modereg(Y = df_ospan$DV_percent_NABS, W= df_ospan$condition, 
#         bw = 2, nstart = 2, PLOT = TRUE)
# 
# summary(m)
```



