---
subtitle: 'COMBINED | 3 Description'
---

\newpage

# Description {#sec-combined-description}


*The purpose of this notebook is describe the distributions of dependent the OSPAN task data.*


```{r}
#| label: SETUP
#| warning : false
#| message : false


library(Hmisc) # %nin% operator
library(mosaic) #simple descriptives [favstats]
library(multimode) #test for multimodality
library(fitdistrplus) #fitting distributions
library(performance) #multimodality
library(kableExtra) #printing tables 
library(vcd) #mosaicplots
library(ggpubr) #arrange plots
library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

```

```{r}
#| label: IMPORT-DATA
#| warning : false
#| message : false

# HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
setwd(mbp)

#IMPORT SUBJECTS 
sgc3a <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds") %>% mutate(
  study = "SGC3A"
) %>% dplyr::select(-absolute_score)
sgc4a <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_participants.rds") %>% mutate(
  study = "SGC4A",
  pretty_mode = "online-replication"
)
sgc4b <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_participants.rds") %>% mutate(
  pretty_mode = "online-replication"
)
sgc5 <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5_scored_participants.rds") %>% mutate(
  study = "SGC5A",
  pretty_mode = "online-replication"
) %>% dplyr::select(-absolute_score)

#MERGE 
df_subjects <- rbind(sgc3a, sgc4a, sgc4b, sgc5)

#IMPORT OSPAN DATA 
ospan <- read_csv("analysis/combined/data/fall21_scored_ospan.csv") %>% mutate(
  subject = SUBJECTID
)

#MERGE OSPAN DATA WITH SGC DATA 
df_ospan <- df_subjects %>% filter(
  subject %in% ospan$subject
)

df_ospan <- merge(df_ospan, ospan)

#RENAME for reuse
df_all <- df_subjects
df_subjects <- df_ospan

#CLEANUP
# rm(ospan,sgc3a, sgc4a, sgc4b, sgc5, df_ospan)
```






### OSPAN Participants

```{r}
#| label: DESC-PARTICIPANTS

title = "OSPAN Participants by Condition"
cols = c("Condition","n")
cont <- table(df_subjects$condition)
cont %>%  addmargins() %>% kbl(caption = title, col.names = cols) %>% kable_classic()

```
## OSPAN

TODO OSPAN description


```{r}
#| label: DESC-SUBJ-OSPAN
 
title = "Descriptive Statistics of OSPAN Task Accuracy"
ospan.stats <- rbind(
  "MATH" = df_subjects %>% dplyr::select(math_acc) %>% unlist() %>% favstats(),
  "ORDER" = df_subjects %>%  dplyr::select(order_acc) %>% unlist() %>% favstats(),
  "WEIGHTED" = df_subjects %>% dplyr::select(weighted) %>% unlist() %>% favstats()
  
) 
ospan.stats %>% kbl (caption = title) %>% kable_classic() %>% 
  footnote(general = "MATH = %correct of all math questions; 
           ORDER = % correct of OSPAN ordering (out of 30); WEIGHTED = math*ospan number correct", general_title = "Note: ",footnote_as_chunk = T) 

```
For (online) OSPAN weighted scores (n = `r ospan.stats["WEIGHTED",]$n`) range from `r round(ospan.stats["WEIGHTED",]$min,2)` to `r round(ospan.stats["WEIGHTED",]$max,2)` with a mean score of (M = `r round(ospan.stats["WEIGHTED",]$mean,2)`, SD = `r round(ospan.stats["WEIGHTED",]$sd,2)`).

```{r}
#| label: VIS-SUBJ-OSPAN

#GGFORMULA | DENSITY HISTOGRAM 
  gf_dhistogram(~weighted, data = df_subjects) + 
  labs(x = "OSPAN (weighted) score",
       y = "% of subjects",
       title = "Distribution of OSPAN SCORE",
       subtitle = "") + 
  theme_minimal()


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_subjects, x = "weighted", binwidth = 1,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of OSPAN Score",
        subtitle ="The Distribution of OSPAN scores is similar across conditions",
        x = "OSPAN (weighted) score", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")


#PLOT EMPIRICIAL CUMULATIVE DISTRIBUTION FUNCTION
ggplot(data = df_subjects, aes(weighted)) + 
  stat_ecdf() + 
  facet_wrap(~pretty_condition) + 
  labs( title = "Empirical Cumulative Density Function — OSPAN",
        x = "OSPAN (weighted) score", 
        y = "Cumulative Probability") + theme_minimal()

```

## PEEKING 

```{r}


#CONDITION ONLY MODEL
m1 <- lm( item_test_SCALED ~ condition, data = df_subjects)
summary(m1)

#OSPAN ONLY MODEL
m2 <- lm( item_test_SCALED ~ weighted, data = df_subjects)
summary(m2)

#STANDARDIZE OSPAN
df_subjects$z_ospan = zscore(df_subjects$weighted)
histogram(df_subjects$z_ospan)

#OSPAN MEDIAN SPLIT
med_ospan = median(df_subjects$weighted)
df_subjects$low_ospan <- as.factor(df_subjects$weighted < med_ospan)

#STANDARDIZED OSPAN ONLY MODEL
m3 <- lm( item_test_SCALED ~ low_ospan, data = df_subjects)
summary(m3)

#MULTIPLE WITH STANDARDIZED OSPAN
mr <- lm(item_test_SCALED ~ condition + low_ospan, data = df_subjects)
summary(mr)
anova(mr)

#SINGLE WITH STANDARDIZED OSPAN
mr2 <- lm(item_test_SCALED ~ condition + low_ospan, data = df_subjects)
summary(mr2)

gf_histogram(~item_test_SCALED, data = df_subjects) %>% 
  gf_facet_grid(low_ospan ~ condition, labeller = label_both)


gf_jitter(item_test_SCALED ~ weighted, color = ~pretty_condition, data = df_subjects) %>% 
  gf_facet_wrap(~pretty_condition)

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = df_subjects) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

```


**OPSAN score does NOT predict performance in the control condition. **
```{r}

control <- df_subjects %>% filter(condition == "111")

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = control) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

#test performance predicted by OSPAN score
m <- lm(item_test_NABS ~ weighted, data = control)
summary(m)

```

**OPSAN score does NOT predict performance in the control condition. **
```{r}

impasse <- df_subjects %>% filter(condition == "121")

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = impasse) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

#test performance predicted by OSPAN score
m <- lm(item_test_NABS ~ low_ospan, data = impasse)
summary(m)

```


```{r}

#330 total 
# —————————
#126 lab 
#204 online  
  #  71 no ospan 
  # 133 ospan 
sgc3a_lab <- df_all %>% filter(mode=="lab-synch") %>% filter(study=="SGC3A")
sgc3a_online <- df_all %>% filter(mode=="asynch") %>% filter(study=="SGC3A")
sgc3a_online_noospan <- sgc3a_online %>% filter(subject %nin% ospan$subject)
sgc3a_onspan <- df_subjects

m <- lm( item_test_NABS ~ condition, data = sgc3a_online_noospan)
summary(m)

m2 <- lm(item_test_NABS ~ condition, data = df_subjects)
summary(m2)


m3 <- lm(item_test_NABS ~ low_ospan, data = df_subjects)
summary(m3)



```


## RESOURCES

-   https://rpkgs.datanovia.com/ggpubr/reference/index.html

```{r}
#| label: session
sessionInfo()
```


