---
subtitle: 'COMBINED | Cheat Sheep'
---

\newpage

**todo check welch? non parametric anova
**check cumulative vs etc. ordinal regression
**check multinomial interpretation [and what it really means]
**todo review Agresti for categorical data analysis


# Combined Study Data {#sec-combined}

**TODO** - TODO DOUBLE CHECK PAIRED WILCOXON ETC\*\*
**non parametric anova? factorial anova?*

```{r}
#| label: SETUP
#| warning : false
#| message : false


library(Hmisc) # %nin% operator
library(janitor) #compare df cols 
library(mosaic) #simple descriptives [favstats]
library(multimode) #test for multimodality
library(fitdistrplus) #fitting distributions
library(performance) #multimodality
library(kableExtra) #printing tables 
library(vcd) #mosaicplots
library(ggpubr) #arrange plots
library(ggstatsplot) #sanity checks 
library(sjPlot) #model plots
library(ggeasy) #easy plot changes



#modelling
library(lme4)
library(ordinal) #mixed ordinal models
library(jtools) # misc helpers 
library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

#SET THEME
theme_set(theme_minimal())
```

## SETUP

```{r}
#| label: IMPORT-SUBJECTS
#| warning : false
#| message : false

# HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
#mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
#setwd(mbp)

##:: IMPORT SUBJEC DATA 

sgc3a <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds") %>% mutate(
  study = "SGC3A"
) #%>% dplyr::select(-absolute_score)

sgc3b <- read_rds("analysis/SGC3B/data/2-scored-data/sgc3b_scored_participants.rds") %>% mutate(
  study = "SGC3B"
) #%>% dplyr::select(-absolute_score)

sgc4a <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_participants.rds") %>% mutate(
  study = "SGC4A",
  pretty_mode = "online-replication"
)
sgc4b <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_participants.rds") %>% mutate(
  study = "SGC4B",
  pretty_mode = "online-replication"
)

# TODO TEMPORARY ONLY REPLACE WITH SCORED DATA 
sgc4c <- read_rds("analysis/SGC4C/data/0-session-level/sgc4c_participants.rds") %>% mutate(
  pretty_mode = "online-replication"
)

sgc5 <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5_scored_participants.rds") %>% mutate(
  study = "SGC5A",
  pretty_mode = "online-replication"
) %>% dplyr::select(-absolute_score)

#COMPARE DF COLUMNS
# janitor::compare_df_cols(sgc3a, sgc4c)
#SGC5 HAS ABSOLUTE SCORE BUT THEY SHOULDN'T 

##:: IMPORT ITEM DATA 

sgc3a_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(study = "SGC3A")
  
sgc3b_items <- read_rds("analysis/SGC3B/data/2-scored-data/sgc3b_scored_items.rds") %>% mutate(study = "SGC3B")
  
sgc4a_items <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_items.rds") %>% mutate(study = "SGC4A")
  
sgc4b_items <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_items.rds") %>% mutate(study = "SGC4B")

# TODO TEMPORARY ONLY REPLACE WITH SCORED DATA 
sgc4c_items <- read_rds("analysis/SGC4C/data/0-session-level/sgc4c_items.rds") %>% mutate(study = "SGC4C")
  
sgc5_items <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5a_scored_items.rds") %>% mutate(study = "SGC5")
  
##:: IMPORT OSPAN DATA 
ospan <- read_csv("analysis/combined/data/fall21_scored_ospan.csv") %>% mutate(
  subject = SUBJECTID
)


# MERGE #TODO ADD IN SGC4C
df_subjects <- rbind(sgc3a, sgc3b, sgc4a, sgc4b, sgc5) %>%  dplyr::select(subject,term,study,condition)
```

## Data Validation

### Conditions per Study

```{r}
paste("SG3A")
addmargins(table(sgc3a$term, sgc3a$pretty_condition))

paste("SG3B")
addmargins(table(sgc3b$term, sgc3b$pretty_condition))

paste("SG4A")
addmargins(table(sgc4a$term, sgc4a$pretty_condition))

paste("SG4B")
addmargins(table(sgc4b$term, sgc4b$pretty_condition))

paste("SG4C")
addmargins(table(sgc4c$term, sgc4c$pretty_condition))

paste("SG5")
addmargins(table(sgc5$term, sgc5$pretty_condition))
```

### Are Subjects Unique?

```{r}
#CHECK FOR DUPLICATES 
n_total <- nrow(df_subjects)
paste(n_total, "subjects records imported ")
n_dist <- nrow(df_subjects %>% distinct(subject))
paste(n_dist, "unique subjects")
subs <- df_subjects$subject
dups <- subs[duplicated(subs)] 
n_dups <- length(dups)
paste(n_dups, "duplicated subjects")
```

## DE-DUPLICATE

### 3A in person \[spring only\]

*3A \[in person\] SPRING 18 ONLY \~ 35 PER CELL*

```{r}


# OPTION 1: sgc3a as spring 18 only [designate fall 18 as sgc3b only]
sub_df <- sgc3a %>% filter(term %in% c("spring18"))
item_df <- sgc3a_items %>% filter(subject %in% sub_df$subject) %>% mutate(
  score_SCALED = as.ordered(score_SCALED)
)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_density(~item_test_NABS, fill= ~condition, data = sub_df) %>%
  gf_facet_grid(condition ~ .) + easy_remove_legend()

gf_density(~item_test_SCALED, fill= ~condition, data = sub_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#::::::::::::WITH STATISTICS
ggbetweenstats(data = sub_df, x = condition, y = item_test_NABS )
ggbetweenstats(data = sub_df, x = condition, y = item_test_SCALED )

```

```{r}

#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("TEST SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_test_NABS ~ pretty_condition, 
            data = sub_df,
            alternative = "less")

wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("ALL SCORE")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

#::::::::::::LOGISTIC REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(m)
plot_model(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

#::::::::::::ORDINAL REGRESSION MODELS
paste("Ordinal regression of q1 score")
m <- clm(score_SCALED ~ condition, data = q1)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 

sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

paste("Ordinal regression of test phase score")
m <- clm(score_SCALED ~ condition, data = test)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 

sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

paste("Ordinal regression of all phase score")
m <- clm(score_SCALED ~ condition, data = all)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 

sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

```


### 3A in person \[fall + spring \]

*3A \[in person\] FALL AND SPRING 18 ONLY \~ 35 PER CELL*

```{r}

# OPTION 1: sgc3a as FALL AND SPRING only [designate fall 18 as sgc3b only]
sub_df <- sgc3a %>% filter(term %in% c("fall17","spring18"))
item_df <- sgc3a_items %>% filter(subject %in% sub_df$subject) %>% mutate(
  score_SCALED = as.ordered(score_SCALED)
)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_density(~item_test_NABS, fill= ~condition, data = sub_df) %>%
  gf_facet_grid(condition ~ .) + easy_remove_legend()

gf_density(~item_test_SCALED, fill= ~condition, data = sub_df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#::::::::::::WITH STATISTICS
ggbetweenstats(data = sub_df, x = condition, y = item_test_NABS )
ggbetweenstats(data = sub_df, x = condition, y = item_test_SCALED )

```

```{r}

#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("TEST SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_test_NABS ~ pretty_condition, 
            data = sub_df,
            alternative = "less")

wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("ALL SCORE")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

#::::::::::::LOGISTIC REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(m)
plot_model(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

#::::::::::::ORDINAL REGRESSION MODELS
paste("Ordinal regression of q1 score")
m <- clm(score_SCALED ~ condition, data = q1)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 

sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

paste("Ordinal regression of test phase score")
m <- clm(score_SCALED ~ condition, data = test)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 

sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

paste("Ordinal regression of all phase score")
m <- clm(score_SCALED ~ condition, data = all)
summary(m)
(ctable <- coef(summary(m)))
(ci <- confint(m)) 
#exponentiate
(e <- exp(cbind(OR = coef(m), ci)))


sjPlot::plot_model(m, type = "pred")
sjPlot::plot_model(m)

```
- Next we see the estimates for the 4 intercepts, which are sometimes called cutpoints. 
- The intercepts indicate where the latent variable is cut to make the three groups that we observe in our data. 
- Note that this latent variable is continuous. In general, these are not used in the interpretation of the results. 
- The cutpoints are closely related to thresholds, which are reported by other statistical packages.
- for k groups there will be k-1 intercepts (cutpoints)
- confirm that the CI does not include 0 (the units are ordered logits [ordered log odds])
- as with logistic regression we exponentiate the coefficients and confints to get odds ratio 
- 

Interpretation: 
_For subjects in the impasse condition, the odds of giving less orthogonal responses (more uncertain or triangular) are 2.16 times that of those in the control condition. 




### 3A online

```{r}

# [1]sgc3a as fall21 winter 22

sub_df <-  sgc3a %>% filter(term %in% c("winter22", "fall21"))
item_df <- sgc3a_items %>% filter(subject %in% sub_df$subject)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE TEST PHASE SCORES
# gf_dhistogram(~item_test_NABS, fill= ~condition, data = sub_df) %>% 
#   gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)

gf_boxplot(item_test_SCALED ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)


#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_q1_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(item_q1_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("TEST SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_test_NABS ~ pretty_condition, 
            data = sub_df,
            alternative = "less")

wilcox.test(item_test_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")

paste("ALL SCORE")
wilcox.test(s_NABS ~ pretty_condition, data = sub_df,
            alternative = "less")

wilcox.test(s_SCALED ~ pretty_condition, data = sub_df,
            alternative = "less")


#::::::::::::REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)


#ORDINAL LOGISTIC REGRESSION
## TODO : double check type of cumulative proportional odds 

#CREATE DATAFRAME OF Q1
df <- q1 %>% mutate(scaled = as.ordered(score_SCALED))

#MODEL POLYR
m <- polr(scaled ~ condition , data = df, Hess=TRUE)
summary(m)
(ctable <- coef(summary(m)))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
(ctable <- cbind(ctable, "p value" = p))
(ci <- confint(m)) # default method gives profiled CIs
performance(m)

library(ordinal)

#MODEL ORDINAL
m1 <- clm(scaled ~ condition, data = df)
summary(m1)
(ctable <- coef(summary(m1)))
(ci <- confint(m1)) # default method gives profiled CIs
performance(m1)
sjPlot::plot_model(m1, type = "pred")
sjPlot::plot_model(m1)

#TEST PHASE
df <- test %>%  mutate(scaled = as.ordered(score_SCALED))

m2 <- clmm(scaled ~ condition + (1|subject), data = df)
summary(m2)
(ctable <- coef(summary(m2)))
(ci <- confint(m2)) # default method gives profiled CIs
performance(m2)
sjPlot::plot_model(m2, type = "est")
sjPlot::plot_model(m2, type = "pred")





#FIT MODEL
# #exponentiate coefficients and CIs 
# ci <- confint(m)
# ci
# e <- coef(m)
# e
# # exp(cbind(e,ci))
# 
# # Retrieve predictions as probabilities 
# # (for each level of the predictor)
# # p.control <- predict(m,data.frame(condition="111"),type="response")
# # paste("Probability of success in control,", p.control)
# # p.impasse <- predict(m,data.frame(condition="121"),type="response")
# # paste("Probability of success in impasse,", p.impasse)
# 
# # Plot Predicted data and original data points
# # ggplot(df, aes(x=condition, y=accuracy)) + 
# #   geom_point() +
# #   stat_smooth(method="glm", color="green", se=FALSE,
# #                 method.args = list(family=binomial))
#   
# #TO PLOT ALL EFFECTS
# library(effects)
# plot(allEffects(m))
# 
# #SJPLOT
# library(sjPlot)
# plot_model(m, )
# 
# 
# #CONVERT TO PROBABILITIES
# newdat <- data.frame(condition=c("111","121"))
# prob <- (phat <- predict(object = m, newdat, type="p"))
# prob
# 

```

### 3B

*3B in person Fall 17*

```{r}

# [1]sgc3b fall 17 only 

sub_df <-  sgc3b %>% filter(term %in% c("fall17"))
item_df <- sgc3b_items %>% filter(subject %in% sub_df$subject)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE TEST PHASE SCORES
# gf_dhistogram(~item_test_NABS, fill= ~condition, data = sub_df) %>% 
#   gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)

gf_boxplot(item_test_SCALED ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)


#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

paste("TEST SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_test_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_test_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_test_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_test_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

paste("ALL SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#::::::::::::REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

```

```{r}

# [1]sgc3a as fall 17 and spring 18 only 
new_3a <- sgc3a %>% filter(term %in% c("spring18", "fall17"))
table(new_3a$condition)
favstats(item_test_NABS ~ condition, data = new_3a)

#::::::::::::VISUALIZE TEST PHASE SCORES
gf_dhistogram(~item_test_NABS, fill= ~condition, data = new_3a) %>% 
  gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = new_3a) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)


#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_q1_NABS ~ pretty_condition, data = new_3a,
            alternative = "less")

paste("TEST SCORE")
#WILCOXON RANK SUM TEST  [non parametric t-test]
wilcox.test(item_test_NABS ~ pretty_condition, data = new_3a,
            alternative = "less")

paste("ALL SCORE")
wilcox.test(s_NABS ~ pretty_condition, data = new_3a,
            alternative = "less")

#::::::::::::REGRESSION MODELS

new_sgc3a_items <- sgc3a_items %>% filter(subject %in% new_3a$subject)
nrow(sgc3a_items) == nrow(new_3a) * 15

q1   <- new_sgc3a_items %>% filter(q ==1)
test <- new_sgc3a_items %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- new_sgc3a_items %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

```

### 4A

**4A WI22 \~ 80 PER CELL**

```{r}

# [1] winter 22

sub_df <-  sgc4a %>% filter(term=="winter22")
item_df <- sgc4a_items %>% filter(subject %in% sub_df$subject)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE TEST PHASE SCORES
# gf_dhistogram(~item_test_NABS, fill= ~condition, data = sub_df) %>% 
#   gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)

gf_boxplot(item_test_SCALED ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)


#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

# paste("TEST SCORE")
# #KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
# kruskal.test(item_test_NABS ~ pretty_condition, data = sub_df)
# 
# #POSTHOC 
# pairwise.wilcox.test(sub_df$item_test_NABS, sub_df$pretty_condition,
#   p.adjust.method = "holm", alternative = "two.sided"
# )
# 
# #KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
# kruskal.test(item_test_SCALED ~ pretty_condition, data = sub_df)
# 
# #POSTHOC 
# pairwise.wilcox.test(sub_df$item_test_SCALED, sub_df$pretty_condition,
#   p.adjust.method = "holm", alternative = "two.sided"
# )

paste("ALL SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#::::::::::::REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
# test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

# paste("Logistic regression of test phase score")
# m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
# summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

```

### 4B

**4B WI22 ONLY \~ 90 PER CELL**

```{r}

# [1]sgc4a 
sub_df <- sgc4b %>% filter(term=="winter22")
item_df <- sgc4b_items %>% filter(subject %in% sub_df$subject)
nrow(item_df) == nrow(sub_df) * 15

table(sub_df$condition)
favstats(item_test_NABS ~ condition, data = sub_df)
favstats(item_test_SCALED ~ condition, data = sub_df)

#::::::::::::VISUALIZE TEST PHASE SCORES
# gf_dhistogram(~item_test_NABS, fill= ~condition, data = sub_df) %>% 
#   gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)

gf_boxplot(item_test_SCALED ~ condition, fill= ~condition, data = sub_df) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)


#::::::::::::NONPARAMETRIC TESTS
paste("Q1 SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_q1_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$item_q1_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

# paste("TEST SCORE")
# #KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
# kruskal.test(item_test_NABS ~ pretty_condition, data = sub_df)
# 
# #POSTHOC 
# pairwise.wilcox.test(sub_df$item_test_NABS, sub_df$pretty_condition,
#   p.adjust.method = "holm", alternative = "two.sided"
# )
# 
# #KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
# kruskal.test(item_test_SCALED ~ pretty_condition, data = sub_df)
# 
# #POSTHOC 
# pairwise.wilcox.test(sub_df$item_test_SCALED, sub_df$pretty_condition,
#   p.adjust.method = "holm", alternative = "two.sided"
# )

paste("ALL SCORE")
#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_NABS ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_NABS, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(s_SCALED ~ pretty_condition, data = sub_df)

#POSTHOC 
pairwise.wilcox.test(sub_df$s_SCALED, sub_df$pretty_condition,
  p.adjust.method = "holm", alternative = "two.sided"
)

#::::::::::::REGRESSION MODELS
q1   <- item_df %>% filter(q ==1)
# test <- item_df %>% filter(q %nin% c(1,2,3,4,5,6,9))
all  <- item_df %>% filter(q %nin% c(6,9))

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = "binomial")
summ(m)

# paste("Logistic regression of test phase score")
# m <- glmer( score_niceABS ~ condition + (1|subject), data = test, family = "binomial")
# summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition + (1|subject), data = all, family = "binomial")
summ(m)

```

### 5A

**5A WI22 ONLY INCOMPLETE MISSING CONTROL**

```{r}

# [1]sgc5a 
new_5a <- sgc5 #%>% filter(term=="winter22")
table(sgc5$condition, sgc5$term)
favstats( item_test_NABS ~ condition, data = sgc5)

table(sgc5$condition, sgc5$term)

#WILCOXON RANK SUM TEST  [non parametric t-test]
# wilcox.test(item_test_NABS ~ pretty_condition, data = df_temp,
#             alternative = "less")

gf_dhistogram(~item_test_NABS, fill= ~condition, data = sgc5) %>% 
  gf_facet_grid(condition ~ .)

gf_boxplot(item_test_NABS ~ condition, fill= ~condition, data = sgc5) %>% 
  gf_jitter(width = 0.1, alpha = 0.5, color = ~condition)

#HOW MANY 5A OSPAN?
#[should be 0]
a5_ospan <- ospan %>% filter(subject %in% sgc5$subject)
nrow(a5_ospan)
#none 

```

### EVALUATE DE-DUP

```{r}
new_df_subjects <- rbind(new_3a,new_3a_online,new_3b,new_4a,new_4b,new_5a)

#CHECK FOR DUPLICATES 
x_total <- nrow(new_df_subjects)
paste(x_total, "subjects should be unique")
x_dist <- nrow(new_df_subjects %>% distinct(subject))
paste(x_dist, "unique subjects")
x_subs <- new_df_subjects$subject
x_dups <- x_subs[duplicated(x_subs)] 
xn_dups <- length(x_dups)
paste(xn_dups, "duplicated subjects")

pilot_3b = 15+29
pilot_3b3 = 5+12+3+11
n_discard = 6+52+5+12+3+11 + 15+29 #hardcoded number of discards 
#1197 unique subjects  + 327 [de-duplicated] + (sgc3b,4a pilots) = 

#number of distinct in original == number of distinct now + n discarded 
n_dist ==  x_dist + n_discard 

```

## OSPAN SETUP

### Find OSPAN subjects

```{r}

#OSPAN RECORDS
n_ospan <- nrow(ospan)
paste(n_ospan, "scored OSPAN records")

#MERGE OSPAN DATA WITH SGC DATA 
ospan_subjects <- new_df_subjects %>% filter(
  subject %in% ospan$subject
)
n_found <- nrow(ospan_subjects)
paste(n_found, "records found")
n_missing <- n_ospan - n_found
paste(n_missing, "ospan records don't have subjects")
#where did the others go? maybe no clean records?
paste("Some of these subjects failed attention check. Others were part of SGC3B online pilot")

#GET MERGED OSPAN
df_ospan <- merge(ospan_subjects, ospan)
#CONFIRM ALL DISTINCT 
df_ospan %>% distinct(subject) %>% nrow() == nrow(df_ospan)

table(df_ospan$condition)
```

**Investigate subjects for SGC3A, SGC3B**

```{r}
#
#validate how many subjects per study
table(df_subjects$study)

#COMBINE SGC3A AND SGC3B
df3 <- rbind(sgc3a, sgc3b)
n_total <- nrow(df3)
paste(n_total, "subjects total SGC3A + SGC3B")
n_dist <- nrow(df3 %>% distinct(subject))
paste(n_dist, "unique subjects")
subs <- df3$subject
dups <- subs[duplicated(subs)] 
n_dups <- length(dups)
paste(n_dups, "duplicated subjects")

#DATAFRAME of DUPLICATES
df_dups <- df3 %>% filter(subject %in% dups)
#what studies?
unique(df_dups$condition)


#filter out sgc3a subjects from 3b study file
df_3b <- sgc3b
df_3b <- sgc3b %>% filter(mode == 'lab-synch')
table(df_3b$condition)



#double check; only sgc3as onlines are not in sgc3b
check <- sgc3a 


#CHECK 
#need to fill cells for SGC3B 

df_a <- sgc3a %>% filter(subject %nin% df_3b$subject)
table(df_a$condition , df_a$mode)

table(sgc3a$condition, sgc3a$mode)

table(sgc3a$term)





```

```{r}

#LOOK AT DUPLICATES
n_total <- nrow(df_subjects)
paste(n_total, "subjects total")
n_dist <- nrow(df_subjects %>% distinct(subject))
paste(n_dist, "unique subjects")
subs <- df_subjects$subject
dups <- subs[duplicated(subs)] 
n_dups <- length(dups)
paste(n_dups, "duplicated subjects")

#DATAFRAME of DUPLICATES
df_dups <- df_subjects %>% filter(subject %in% dups)
#what studies?
unique(df_dups$condition)
table(df_dups$condition, df_dups$study)

#NEED 62 for SGC4A

```

```{r}

df_3b <- df_3b %>% mutate(
  IMPASSE = recode_factor(condition, 
                           "111" = "none",
                           "121" = "impasse",
                           "211" = "none",
                           "221" = "impasse",
                           "311" = "none",
                           "321" = "impasse"),
   EXPLICIT = recode_factor(condition, 
                           "111" = "none",
                           "121" = "none",
                           "211" = "img",
                           "221" = "img",
                           "311" = "ixn",
                           "321" = "ixn")
)


gf_dhistogram(~item_test_NABS, data = df_3b) %>% 
  gf_facet_grid(IMPASSE ~ EXPLICIT)

gf_boxplot(item_test_NABS ~ condition, data = df_3b) %>% 
  gf_facet_grid(IMPASSE ~ EXPLICIT)



m <- lm(item_test_NABS ~ IMPASSE*EXPLICIT, data = df_3b)
summ(m)
check_model(m)



#KRUSKAL WALLIS RANK SUM TEST  [non parametric anova]
kruskal.test(item_test_NABS ~ pretty_condition, data = df_3b)

#POSTHOC 
pairwise.wilcox.test(df_3b$item_test_NABS, df_3b$pretty_condition,
  p.adjust.method = "holm"
)

```

```{r}
# 
# #MERGE OSPAN DATA WITH SGC DATA 
# df_ospan <- df_subjects %>% filter(
#   subject %in% ospan$subject
# )
# 
# df_ospan <- merge(df_ospan, ospan)
# 
# #RENAME for reuse
# df_all <- df_subjects
# # df_subjects <- df_ospan
# 
# #RECODE other study control conditions to CONTROL
# df <- df_all %>% mutate(
#   pretty_condition = recode_factor(pretty_condition, "Orth-Full" = "control",
#                                    "point" = "control")
# )
# 
# #CLEANUP
# # rm(ospan, sgc3a, sgc4a, sgc4b, sgc5, df_ospan)
# 
# temp <- df_all %>% dplyr::select(s_NABS, condition, study, subject) %>% rename(score = s_NABS)
# temp2 <- sgc4c %>% dplyr::select(absolute_score, condition, study, subject) %>% rename(score = "absolute_score")
# df <- rbind(temp, temp2)
```

```{r}
# table(df$study, df$condition)
# 
# paste("Subjects: ", nrow(df))
# paste("unique subjects ",nrow(unique(df$subject)))
# 
# #get list of unique subjects 
# u <- df %>% dplyr::select(subject) %>% distinct() 
# d <- df$subject %nin% u$subject
# 
# #unique subjects
# u <- unique(df$subject)
# #duplicated subjects
# d <- df$subject[duplicated(df$subject)]
# 
# length(d) 
# length(u) 
# length(d) + length(u) == nrow(df)
# 
# dups <- df %>% filter(subject %in% d)
# 
# table(dups$study, dups$condition)
# 
# #TODO ... NEED TO RUN 4A CONTROL 62
```

```{r}
#| label: IMPORT-ITEMS
#| warning : false
#| message : false

# HACK WD FOR LOCAL RUNNING?
imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
setwd(mbp)

#IMPORT SUBJECTS 
items_sgc3a <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(
  study = "SGC3A"
) #%>% dplyr::select(-absolute_score)

items_sgc3b <- read_rds("analysis/SGC3B/data/2-scored-data/sgc3b_scored_items.rds") %>% mutate(
  study = "SGC3B"
) 

items_sgc4a <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_items.rds") %>% mutate(
  study = "SGC4A",
  pretty_mode = "online-replication"
)

items_sgc4b <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_items.rds") %>% mutate(
  pretty_mode = "online-replication"
) %>% dplyr::select(-answer)

##TEMPORARY 
items_sgc4c <- read_rds("analysis/SGC4C/data/0-session-level/sgc4c_items.rds") %>% mutate(
  pretty_mode = "online-replication"
) %>% dplyr::select(-answer)


items_sgc5 <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5_scored_items.rds") %>% mutate(
  study = "SGC5A",
  pretty_mode = "online-replication"
) %>% dplyr::select(-answer)
# 
# #COMPARE DF COLUMNS
# # janitor::compare_df_cols(sgc3a, sgc3b, sgc4a, sgc4b, sgc5)
# #SGC5 HAS ABSOLUTE SCORE BUT THEY SHOULDN'T 
# 
# #MERGE 
df_items <- rbind(items_sgc3a, items_sgc3b, items_sgc4a, items_sgc4b, items_sgc5)

#RECODE other study control conditions to CONTROL
df_items <- df_items %>% mutate(
  pretty_condition = recode_factor(pretty_condition, "Orth-Full" = "control",
                                   "point" = "control")
)


# #TODO FIGURE OUT DUPLICATES
# #IMPORT SCORING FUNCTIONS
# # source("analysis/utils/scoring.R")
# 
# progress_Absolute <- function(items){
#   
#   #filter for valid items
#   x <- items %>% filter(q %nin% c(6,9)) %>% dplyr::select(subject, pretty_condition, q,score_niceABS) 
#   
#   #pivot wider
#   wide <- x %>% pivot_wider(names_from=q, names_glue = "q_{q}", values_from = score_niceABS)
#   
#   #calc stepwise cumulative score
#   wide$c1 = wide$q_1
#   wide$c2 = wide$c1 + wide$q_2
#   wide$c3 = wide$c2 + wide$q_3
#   wide$c4 = wide$c3 + wide$q_4
#   wide$c5 = wide$c4 + wide$q_5
#   wide$c6 = wide$c5 + wide$q_7
#   wide$c7 = wide$c6 + wide$q_8
#   wide$c8 = wide$c7 + wide$q_10
#   wide$c9 = wide$c8 + wide$q_11
#   wide$c10 = wide$c9 + wide$q_12
#   wide$c11 = wide$c10 + wide$q_13
#   wide$c12 = wide$c11 + wide$q_14
#   wide$c13 = wide$c12 + wide$q_15
#   wide <- wide %>% dplyr::select(subject, pretty_condition,c1,c2,c3,c4,c5,c6, c7,c8,c9, c10,c11,c12,c13)
#   
#   #lengthen 
#   df_absolute_progress <- wide %>% pivot_longer(cols= c1:c13, names_to = "question", names_pattern = "c(.*)", values_to = "score")
#   df_absolute_progress$question <- as.integer(df_absolute_progress$question)
#   
#   #cleanup 
#   rm(x,wide)
#   
#   return(df_absolute_progress)
# }
# 
# 
# #GET ABSOLUTE PROGRESS 
# df_absolute_progress <- progress_Absolute(df_items %>% filter(pretty_condition=="control"))
# 
# #GET SCALED PROGRESS
# df_scaled_progress <- progress_Scaled(df_items)


```

```{r}

# gf_boxplot(s_NABS ~ pretty_condition, data = df_all)
# gf_boxplot(absolute_score ~ pretty_condition, data=sgc4c )

df <- df_items %>% dplyr::select(score_niceABS, pretty_condition) %>% rename(correct=score_niceABS)
d2 <- items_sgc4c %>% dplyr::select(correct, pretty_condition) 

df <- rbind(df, d2)

gf_props(~correct, fill = ~pretty_condition, data = df) %>% 
  gf_facet_wrap(pretty_condition ~ .)


```

### COMBINED

```{r}

library(ggdist)


gf_boxplot( s_NABS ~ pretty_condition, color = ~pretty_condition, data = df_all) +
  labs(title = "Distribution of ABSOLUTE Score by Condition")


gf_boxplot(s_SCALED~ pretty_condition, color = ~pretty_condition, data = df_all) +
  labs(title = "Distribution of SCALED Score by Condition")


#GGDIST HALFEYE
ggplot(df_all, aes(y = pretty_condition, x = s_NABS,
                        fill = pretty_condition) ) + 
  ggdist::stat_halfeye(
    point_interval = "mean_qi"
    # side = "left",
    # justification = 1.1,
    # width = 1, 
    # point_colour = NA
   ) + labs (
     title = "Distribution of Absolute Score by Condition",
     x = "Total Absolute Score"
   ) + theme(legend.position = "blank")


#DENSITY RIDGES BINLINE
ggplot(df_all, aes(x = DV_percent_NABS, y = pretty_condition, height=stat(density))) +
  # geom_density_ridges( stat = "density",alpha = 0.7)
  geom_density_ridges( alpha = 0.7, stat = "binline", bins = 20,   scale = 0.75, draw_baseline = FALSE) +
  labs(title = "Distribution of ABSOLUTE Score by Condition")
  
  

#DENSITY RIDGES SMOOTH
ggplot(df_all, aes(x = DV_percent_NABS, y = pretty_condition, fill=pretty_condition)) +
  xlim(0,1) + 
  stat_density_ridges( scale = 0.75, stat = "density", jittered_points = TRUE, 
        position = position_points_jitter(width = 0.05, height = 0),
        point_shape = '|', point_size = 2, point_alpha = 1, alpha = 0.7) +
  labs(title = "Distribution of ABSOLUTE Score by Condition")
```

## RESPONSE ACCURACY

### Subject Level Scores

Subject level scores summarize the the response accuracy by a particular participant across all discriminant items in the graph comprehension task.

#### Absolute Score

```{r}
#| label: VIS-SUBJ-ABS-TEST

#GGFORMULA | DENSITY HISTOGRAM SUBJECT TOTAL ABSOLUTE
  gf_props(~s_NABS, data = df_all) + 
  labs(x = "number of correct responses ",
       y = "% of subjects",
       title = "Distribution of Absolute Score ",
       subtitle = "Modes at high and low ends of scale suggest concentration of high (vs) low understanding") 


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_all, x = "s_NABS", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of Absolute Score",
        subtitle ="Pattern of response is similar across data collection modes but differs by condition",
        x = "Total Absolute Score", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")

##VERTICAL RAINCLOUD USING GGDISTR
ggplot(df_all, aes(x = pretty_condition, y = s_NABS,
                        fill = pretty_condition) ) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.1,
    width = 1, 
    point_colour = NA
   ) + 
  geom_boxplot(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = s_NABS),
    width = .15, 
    outlier.shape = NA
  ) + 
  geom_point(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = s_NABS, color = pretty_condition),
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .05
  )) + labs( 
    title = "Distribution of Absolute Score ",
    x = "Condition", y = "Total Absolute Score") +
  theme(legend.position = "blank") + 
  coord_cartesian(xlim = c(0.5, NA), clip = "off")

#PLOT EMPIRICIAL CUMULATIVE DISTRIBUTION FUNCTION
ggplot(data = df_all, aes(item_test_NABS)) + 
  stat_ecdf(geom = "step") + 
  facet_wrap(pretty_condition~.) + 
  labs( title = "Empirical Cumulative Density Function — Absolute Score ",
        x = "Total Absolute Score [0,13]", 
        y = "Cumulative Probability")

  
```

#### Scaled Scores

```{r}
#| label: VIS-SUBJ-SCALED-TEST

#GGFORMULA | DENSITY HISTOGRAM SUBJECT TOTAL SCALED
gf_props(~s_SCALED, data = df_all) +
  labs(x = "total scaled score",
       y = "% of subjects",
       title = "Distribution of Scaled Score ",
       subtitle = "Modes at high and low ends of scale suggest concentration of high (vs) low understanding") 


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_all, x = "s_SCALED",binwidth=1,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE) 
facet(p, facet.by=c("pretty_condition")) + 
  labs( title = "Distribution of Scaled Score",
        subtitle ="Pattern of response is similar across data collection modes but differs by condition",
        x = "total scaled score (test phase)", y = "number of participants") + 
 theme_minimal() + theme(legend.position = "blank") 


##VERTICAL RAINCLOUD USING GGDISTR
ggplot(df_all, aes(x = pretty_condition, y = s_SCALED,
                        fill = pretty_condition) ) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.1,
    width = 1, 
    point_colour = NA
   ) + 
  geom_boxplot(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = s_SCALED),
    width = .15, 
    outlier.shape = NA
  ) + 
  geom_point(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = s_SCALED, color = pretty_condition),
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .05
  )) + labs( 
    title = "Distribution of Scaled Score ",
    x = "Condition", y = "Total Absolute Score (Test Phase)") +
  theme(legend.position = "blank") + 
  coord_cartesian(xlim = c(0.5, NA), clip = "off")


#PLOT EMPIRICIAL CUMULATIVE DISTRIBUTION FUNCTION
ggplot(data = df_all, aes(s_SCALED)) + 
  stat_ecdf(geom = "step") + 
  facet_wrap(~pretty_condition) + 
  labs( title = "Empirical Cumulative Density Function — Scaled Score",
        x = "Test Phase Scaled Score [-8,8]", 
        y = "Cumulative Probability") 

```

### First Item Scores

Next we consider the response accuracy on *just* the first question of the graph comprehension task: a subject's first exposure to the TM graph.

#### First Item Absolute Score

```{r}
#| label: VIS-FIRST-ABSOLUTE

#PROPORTIONAL BAR CHART
gf_props(~item_q1_NABS, data = df_all) +
  labs(x = "response accuracy",
       y = "% subjects",
       title = "Proportion of Correct Responses on First Item",
       subtitle="")+
  theme(legend.position = "none")+theme_ggdist()

#PROPORTIONAL BAR CHART
gf_props(~item_q1_NABS, data = df_all, fill = ~pretty_condition) %>% 
  gf_facet_wrap(~pretty_condition) +
  labs(x = "response accuracy",
       title = "Proportion of Correct Responses on First Item (by Modality and Condition)",
       subtitle="")+
  theme(legend.position = "none") 


#MOSAIC PLOT
# vcd::mosaic(main="Proportion of Correct Responses on First Item",
#             data = df_all, pretty_condition ~ item_q1_NABS, rot_labels=c(0,90,0,0), 
#             offset_varnames = c(left = 4.5), offset_labels = c(left = -0.5),just_labels = "right",
#             spacing = spacing_dimequal(unit(1:2, "lines"))) 


```

#### First Item Scaled Score

At the item level, the scaled score gives us a numeric measure of correctness of interpretation, ranging from -1 to 1. (note: we evaluate scaled_score on the first item rather than interpretation, because no orthogonal interpretation is available in the impasse condition)

```{r}
#| label: VIS-FIRST-SCALED

#GGFORMULA | PROPORTIONAL HISTOGRAM SUBJECT FIRST SCALED
gf_props(~item_q1_SCALED, data = df_all) +
  labs(x = "scaled score (first item)",
       y = "% of subjects",
       title = "Distribution of First Item Scaled Score",
       subtitle = "") 


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_all, x = "item_q1_SCALED", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE) 
facet(p, facet.by=c("pretty_condition")) + 
  labs( title = "Distribution of First Item Scaled Score (by Mode and Condition)",
        subtitle ="Impasse condition yields more intermediate scores (indicating uncertainty)",
        x = "scaled score (firt item) ", y = "number of participants") + 
  theme_minimal() + theme(legend.position = "blank") 


##GGFORMULA | HIST+DENSITY SCORE BY CONDITION/MODE
# stats = df_subjects %>% group_by(pretty_condition, mode) %>% dplyr::summarise(mean = mean(item_q1_SCALED))
# gf_density(~item_q1_SCALED, data = df_subjects) %>%
#   gf_facet_grid(pretty_condition~mode, labeller = label_both) %>%
#   gf_lims(x = c(-1, 1)) %>%
#   gf_vline(data = stats, xintercept = ~mean, color = "red") +
# labs( title = "Distribution of First Item Scaled Score (by Mode and Condition)",
#         subtitle ="Pattern of response is the same across data collection modes but differs by condition",
#         x = "scaled score (firt item) ", y = "number of participants") + 
#   theme_minimal()
  
```

### Interpretation Scores

TODO IMPORT ITEMS

```{r}
#| label: VIS-ITEM-INTERPRETATION


# #PROPORTIONAL BAR CHART
# gf_propsh(~interpretation, data = df_items, fill = ~pretty_condition) %>% 
#   gf_facet_grid(pretty_condition~pretty_mode) +
#   labs(x = "% of items",
#        title = "Proportion of Interpretations Across Items",
#        subtitle="Impasse Condition yields shift from Orthogonal to alternative interpretations")+
#   theme(legend.position = "none")


#MOSAIC PLOT
# vcd::mosaic(main="Proportion of Interpretations across Conditions",
#             data = df_items, pretty_condition ~ interpretation, rot_labels=c(0,90,0,0), 
#             offset_varnames = c(left = 4.5), offset_labels = c(left = -0.5),just_labels = "right",
#             spacing = spacing_dimequal(unit(1:2, "lines"))) 


```

### Cumulative Task Performance

**TODO CALC CUMULATIVE PERFORAMNCE OVER ALL ITEMS**

```{r}
#| label: VIZ-PROGRESS

# #VISUALIZE progress over time ABSOLUTE score 
# ggplot(data = df_absolute_progress, aes(x = question, y = score, group = subject, alpha = 0.01, color = pretty_condition)) + 
#  geom_line(position=position_jitter(w=0.15, h=0.15), size=0.1) +
#  facet_wrap(~pretty_condition) + 
#  labs (title = "Cumulative Absolute Score over sequence of task", x = "Question" , y = "Cumulative Absolute Score") + 
#  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13)) +
#  theme_minimal() + theme(legend.position = "blank")
# 
# #VISUALIZE progress over time SCALED score 
# ggplot(data = df_scaled_progress, aes(x = question, y = score, group = subject, alpha = 0.01, color = pretty_condition)) + 
#  geom_line(position=position_jitter(w=0.15, h=0.15), size=0.1) +
#  facet_wrap(~pretty_condition) + 
#  labs (title = "Cumulative Scaled Score over sequence of task", x = "Question" , y = "Cumulative Scaled Score") + 
#  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13)) +
#  theme_minimal() + theme(legend.position = "blank")

```

## RESPONSE LATENCY

### Time on First Item

```{r}
#| label : VIS-FIRSTTIME
#| message : false
#| warning : false


#HISTOGRAM
gf_dhistogram(~item_q1_rt, data = df_all) %>%
  # gf_vline(xintercept = ~time.stats["lab",]$mean, color = "black") %>% 
  gf_fitdistr(dist="gamma", color="red")+
  labs(title="Distribution of First Item Response Time (seconds)", subtitle = "fit by gamma distribution", x = "First Item Response Time (seconds)", y = "% items")


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_all, x = "item_q1_rt", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of First Item Response Time (seconds)",
        subtitle ="",
        x = "First Item Response Time (seconds)", y = "number of items") +
  theme_minimal() + theme(legend.position = "blank")



#recode as boolean correct
df_subjects <- df_subjects %>% mutate(
  item_q1_NABS = as.logical(item_q1_NABS)
)

##RAINCLOUD USING GGDISTR
ggplot(df_all, aes(x = pretty_condition, y = item_q1_rt, color = item_q1_NABS) ) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.2, 
    adjust = .5, 
    width = .6, 
    .width = 0, 
    point_colour = NA) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .1
  )) + 
  labs( title = "Distribution of First Item Response Time (seconds)",
        subtitle ="",
        y = "First Item Response Time (s)", x = "Condition") +
  theme_ggdist() 
# + theme(legend.position = "blank")
# + coord_cartesian(xlim = c(1.2, NA), clip = "off")
                
```

### Time on Item

**TODO**

```{r}
#| label : VIS-ITEMTIME
#| message : false
#| warning : false

# 
# #HISTOGRAM
# gf_dhistogram(~rt_s, data = df_items) %>%
#   gf_vline(xintercept = ~time.stats["lab",]$mean, color = "black") %>% 
#   gf_fitdistr(dist="gamma", color="red")+
#   labs(title="Distribution of Item Response Time (seconds)", 
#        subtitle = "fit by gamma distribution", x = "Item Response Time (seconds)", y = "% items") 
# 
# 
# ##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
# p <- gghistogram(df_items, x = "rt_s", binwidth = 0.5,
#    add = "mean", rug = TRUE,
#    fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
#    add_density = TRUE)
# facet(p, facet.by=c("pretty_condition","pretty_mode")) +
#   labs( title = "Distribution of Item Response Time (seconds)",
#         subtitle ="",
#         x = "Item Response Time (seconds)", y = "number of items") +
#   theme_minimal() + theme(legend.position = "blank")
# 
# 
# #recode as boolean correct
# df_items <- df_items %>% mutate(
#   score_niceABS = as.logical(score_niceABS)
# )
# 
# ##RAINCLOUD USING GGDISTR
# ggplot(df_items, aes(x = pretty_condition, y = rt_s, color = score_niceABS) ) + 
#   ggdist::stat_halfeye(
#     side = "left",
#     # position = position_dodgejust(),
#     justification = 1.5, 
#     # adjust = .5, 
#     width = .5, 
#     .width = 0, 
#     point_colour = NA) + 
#   geom_boxplot(
#     width = .15, 
#     outlier.shape = NA,
#     position = position_dodge2()
#   ) +
#   geom_point(
#     size = 1.3,
#     alpha = .3,
#     position = position_jitterdodge(
#       # seed = 1,
#       dodge.width = 0.5,
#       jitter.width = 0.075
#   )) +
#   labs( title = "Distribution of Item Response Time (seconds)",
#         subtitle ="",
#         y = "Item Response Time (s)", x = "Condition") +
#   theme_ggdist() 
# # + theme(legend.position = "blank")
# # + coord_cartesian(xlim = c(1.2, NA), clip = "off")
                
```

### Time on SCAFFOLD Phase

Here we consider *just* the time spent on the first five items of the task (the scaffold phase).

```{r}
#| label : VIS-SCAFFOLDTIME
#| message : false
#| warning : false


#HISTOGRAM
gf_dhistogram(~item_scaffold_rt, data = df_all) %>%
  # gf_vline(xintercept = ~time.stats["lab",]$mean, color = "black") %>% 
  gf_fitdistr(dist="gamma", color="red")+
  labs(title="Distribution of SCAFFOLD Phase Response Time (minutes)", subtitle = "fit by gamma distribution", x = "Scaffold Phase Time (minutes)", y = "% subjects") 


##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_all, x = "item_scaffold_rt", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of SCAFFOLD Phase Response Time (minutes)",
        subtitle ="",
        x = "Scaffold Phase Time (minutes)", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")

##VERTICAL RAINCLOUD USING GGDISTR
ggplot(df_all, aes(x = pretty_condition, y = item_scaffold_rt, fill = pretty_condition)) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.2, 
    adjust = .5, 
    width = .6, 
    .width = 0, 
    point_colour = NA) + 
  geom_boxplot(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = item_scaffold_rt),
    width = .15, 
    outlier.shape = NA
  ) + 
  geom_point(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=pretty_condition, y = item_scaffold_rt, color = pretty_condition),
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .05
  ))+ labs( title = "Distribution of SCAFFOLD Phase Response Time (minutes)",
        subtitle ="",
        y = "Total Study Time (minutes)", x = "Condition") +
  theme_ggdist() + theme(legend.position = "blank") #+
  # coord_cartesian(xlim = c(0.5, NA), clip = "off")
                
```

## EXPLORING RELATIONSHIPS

### ACCURACY (VS) LATENCY

```{r}

#SCATTERPLOT [SCORE X ITEM RESPONSE TIME]
gf_jitter( s_SCALED ~ rt_m, data = df_all, alpha = 0.5, color=~pretty_condition) %>% 
  gf_facet_wrap(~pretty_condition) + labs(
    title = "Total (Scaled) Score by Total Item Response Time",
    subtitle = "", 
    x = "Total Item Response Time (minutes)", y = "Total Scaled Score "
  ) + theme(legend.position = "blank")


#SCATTERPLOT [SCORE X ITEM RESPONSE TIME]
gf_jitter( s_SCALED ~ item_avg_rt, data = df_all, alpha = 0.5, color=~pretty_condition) %>% 
  gf_facet_wrap(~pretty_condition) + labs(
    title = "Total (Scaled) Score by Average Item Response Time",
    subtitle = "", 
    x = "Average Item Response Time (seconds)", y = "Total Scaled Score"
  ) + theme(legend.position = "blank")

#SCATTERPLOT [SCORE X ITEM RESPONSE TIME]
gf_jitter( s_SCALED ~ item_scaffold_rt, data = df_all, alpha = 0.5, color=~pretty_condition) %>% 
  gf_facet_wrap(~pretty_condition) + labs(
    title = "Total (Scaled) Score by SCAFFOLD PHASE Item Response Time",
    subtitle = "", 
    x = "SCAFFOLD PHASE Item Response Time (minutes)", y = "Total Scaled Score "
  ) + theme(legend.position = "blank")


#SCATTERPLOT [SCORE X ITEM RESPONSE TIME]
gf_jitter( s_SCALED ~ item_max_rt, data = df_all %>% filter(item_max_rt < 400), alpha = 0.5, color=~pretty_condition) %>% 
  gf_facet_wrap(~pretty_condition) + labs(
    title = "Total (Scaled) Score by MAX Item Response Time",
    subtitle = "", 
    x = "MAX Item Response Time (s)", y = "Total Scaled Score "
  ) + theme(legend.position = "blank")



```

```{r}
# 
# q.stats <- df_items %>% filter(q != 6) %>% dplyr::group_by(q, pretty_condition, score_niceABS) %>% summarise(
#   m = mean(rt_s),
#   sd = sd(rt_s),
#   group = paste(pretty_condition,"-",score_niceABS)
# )
# 
# gf_line( m ~ as.factor(q), group = ~group,  color = ~score_niceABS,data = q.stats) %>% 
#   gf_point() %>% 
#   gf_facet_wrap(~pretty_condition) + scale_color_manual(values=c("red","green")) + 
#   labs(title = "Average ITEM response time by condition",
#        subtitle = "Correct responses are generally faster [computational efficiency] except on Q1 [learning]",
#        x = "Question", y = "Averate Item Response Time", color="Correct Response")

```

```{r}
# 
# q.stats <- df_items %>% filter(q != 6) %>% dplyr::group_by(q, pretty_condition, interpretation) %>% summarise(
#   m = mean(rt_s),
#   sd = sd(rt_s),
#   group = paste(pretty_condition,"-",score_SCALED)
# )
# 
# gf_line( m ~ as.factor(q), group = ~group,  color = ~interpretation,data = q.stats) %>% 
#   gf_point() %>% 
#   gf_facet_wrap(~pretty_condition) + #+ scale_color_manual(values=c("red","green")) + 
#   labs(title = "Average ITEM response time by condition",
#        subtitle = "Correct responses are generally faster [computational efficiency] except on Q1 [learning]",
#        x = "Question", y = "Averate Item Response Time", color="Interpretation")

```

```{r}
# 
# x <- df_items %>% filter(q %nin% c(6,9)) %>% dplyr::group_by(subject) %>% mutate(
#                             total = sum(score_niceABS),
#                             perf = as.factor(total < 5),
#                             perf = recode(perf, "FALSE"="high-scoring", "TRUE"="low-scoring")) %>% ungroup()
# 
# gf_line( rt_s ~ as.factor(q), group = ~subject, color = ~interpretation, data = x) %>% 
#   gf_facet_grid(pretty_condition ~ perf)+ scale_x_discrete(labels = c(1 , 2,  3,  4,  5,  7,  8, 10, 11, 12, 13, 14, 15))+ labs(
#     title = "Response Time (s) by Question [facet by score <5]"
#   )

```

```{r}

gf_boxplot( item_avg_rt ~ pretty_condition,  data = df_all) +
  labs(title = "Average item response time by mode and condition")


gf_boxplot( item_scaffold_rt ~ pretty_condition,  data = df_all) +
  labs(title = "Average SCAFFOLD response time by mode and condition")

gf_boxplot( item_test_rt ~ pretty_condition,  data = df_all) +
  labs(title = "Average TEST response time by mode and condition")

gf_boxplot( totaltime_m ~ pretty_condition,  data = df_all) +
  labs(title = "Average TOTAL response time by mode and condition")

gf_boxplot( item_q1_rt ~ pretty_condition,  data = df_all) %>% 
  gf_jitter(width=0.2, alpha = 0.5, size = 0.75, color = ~item_q1_NABS) +
  labs(title = "Average FIRST ITEM response time by mode and condition")

```

## MODELS

```{r}

library(report)
library(see)
library(generics)
library(modelr)
library(distributional)

## A GIANT MODEL
m <- lm(DV_percent_NABS ~ pretty_condition, data = df_all)
summary(m)
report(m)
check_model(m)

##UNCERTAINTY ESTIMATES OF THE MODEL 
df_all %>%
  data_grid(pretty_condition) %>%
  augment(m, newdata = ., se_fit = TRUE) %>%
  ggplot(aes(y = pretty_condition)) +
  stat_halfeye(
    aes(xdist = dist_student_t(df = df.residual(m), mu = .fitted, sigma = .se.fit)), 
    scale = .5
  ) +
  # we'll add the data back in too (scale = .5 above adjusts the halfeye height so
  # that the data fit in as well)
  geom_point(aes(x = DV_percent_NABS), data = df_all, pch = "|", size = 2, position = position_nudge(y = -.15)) +
  labs(
    title = "Regression Model | Predicted Conditional Means",
    caption = " % Correct ~ Condition; halfeye estimated using student's t centered at model predicted mean with SE"
  )

```

### OSPAN

```{r}
#| label: DESC-PARTICIPANTS

title = "OSPAN Participants by Condition"
cols = c("Condition","n")
cont <- table(df_ospan$pretty_condition)
cont %>%  addmargins() %>% kbl(caption = title, col.names = cols) %>% kable_classic()

```

**TODO OSPAN description**

```{r}
#| label: DESC-SUBJ-OSPAN
 
title = "Descriptive Statistics of OSPAN Task Accuracy"
ospan.stats <- rbind(
  "MATH" = df_ospan %>% dplyr::select(math_acc) %>% unlist() %>% favstats(),
  "ORDER" = df_ospan %>%  dplyr::select(order_acc) %>% unlist() %>% favstats(),
  "WEIGHTED" = df_ospan %>% dplyr::select(weighted) %>% unlist() %>% favstats()
  
) 
ospan.stats %>% kbl (caption = title) %>% kable_classic() %>% 
  footnote(general = "MATH = %correct of all math questions; 
           ORDER = % correct of OSPAN ordering (out of 30); WEIGHTED = math*ospan number correct", general_title = "Note: ",footnote_as_chunk = T) 

```

For (online) OSPAN weighted scores (n = `r ospan.stats["WEIGHTED",]$n`) range from `r round(ospan.stats["WEIGHTED",]$min,2)` to `r round(ospan.stats["WEIGHTED",]$max,2)` with a mean score of (M = `r round(ospan.stats["WEIGHTED",]$mean,2)`, SD = `r round(ospan.stats["WEIGHTED",]$sd,2)`).

```{r}
#| label: VIS-SUBJ-OSPAN

#GGFORMULA | DENSITY HISTOGRAM 
  gf_dhistogram(~weighted, data = df_ospan) + 
  labs(x = "OSPAN (weighted) score",
       y = "% of subjects",
       title = "Distribution of OSPAN SCORE",
       subtitle = "") 

##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_ospan, x = "weighted", binwidth = 1,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of OSPAN Score",
        subtitle ="The Distribution of OSPAN scores is similar across conditions",
        x = "OSPAN (weighted) score", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")


#PLOT EMPIRICIAL CUMULATIVE DISTRIBUTION FUNCTION
ggplot(data = df_ospan, aes(weighted)) + 
  stat_ecdf() + 
  facet_wrap(~pretty_condition) + 
  labs( title = "Empirical Cumulative Density Function — OSPAN",
        x = "OSPAN (weighted) score", 
        y = "Cumulative Probability") + theme_minimal()

```

```{r}

#IXN PLOT

# ospan.stats <- favstats(weighted, groups = condition)


```

#### NONPARAMETRIC TEST? nope no factorial anova

```{r}

# HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
setwd(mbp)

#IMPORT SUBJECTS 
df_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(
  study = "SGC3A"
) #%>% dplyr::select(-absolute_score)

#FILTER GET ONLY neeed records

ospan_items <- df_items %>% filter(subject %in% df_ospan$subject)
nrow(ospan_items)
#VERIFY ITEMS ARE FOR SSPAN SUBJECTS
nrow(ospan_items) == nrow(df_ospan) * 15


#STANDARDIZE OSPAN
df_ospan$z_ospan = zscore(df_ospan$weighted)
histogram(df_ospan$z_ospan)

#OSPAN MEDIAN SPLIT
med_ospan = median(df_ospan$weighted)
df_ospan$low_ospan <- as.factor(df_ospan$weighted < med_ospan)


#JOIN TO ospan to items 
df_test <- left_join(df_ospan , ospan_items, by = 'subject') %>% 
  dplyr::select(subject, condition.y, q, score_niceABS,  score_SCALED, low_ospan, z_ospan, weighted, math_acc, order_acc) %>% rename(condition = condition.y)


items <- df_test
q1 <- items %>% filter(q==1)
test <- items %>% filter(q %nin% c(1,2,3,4,5,6,9))
all <- items 

paste("Logistic regression of q1 score")
m <- glm( score_niceABS ~ condition, data = q1, family = 'binomial')
summ(m)

paste("Logistic regression of test phase score")
m <- glmer( score_niceABS ~ condition*low_ospan+ (1|subject), data = test, family='binomial')
summ(m)

paste("Logistic regression of all items score")
m <- glmer( score_niceABS ~ condition*low_ospan+ (1|subject), data = all, family = "binomial")
summ(m)

```

```{r}



```

#### MODELS

```{r}

#CONDITION ONLY MODEL
m1 <- lm( item_test_SCALED ~ condition, data = df_ospan)
summary(m1)

#OSPAN ONLY MODEL
m2 <- lm( item_test_SCALED ~ weighted, data = df_ospan)
summary(m2)

#STANDARDIZE OSPAN
df_ospan$z_ospan = zscore(df_ospan$weighted)
histogram(df_ospan$z_ospan)

#OSPAN MEDIAN SPLIT
med_ospan = median(df_ospan$weighted)
df_ospan$low_ospan <- as.factor(df_ospan$weighted < med_ospan)

#STANDARDIZED OSPAN ONLY MODEL
m3 <- lm( item_test_SCALED ~ low_ospan, data = df_ospan)
summ(m3)

#MULTIPLE WITH STANDARDIZED OSPAN
mr <- lm(item_test_SCALED ~ condition + low_ospan, data = df_ospan)
summ(mr)
anova(mr)

library(interactions)
cat_plot(mr, item_test_SCALED, pred = condition, modx = low_ospan)
df <- df_ospan %>% mutate(
  condition = recode_factor(condition,
                            "111"="111","121"="121")
)
library(ggstatsplot)
grouped_ggbetweenstats(data = df,
                       y = item_test_SCALED,
                       x = condition,
                       grouping.var = low_ospan)

#SINGLE WITH STANDARDIZED OSPAN
mr2 <- lm(item_test_SCALED ~ condition + low_ospan, data = df_ospan)
summary(mr2)

gf_histogram(~item_test_SCALED, data = df_ospan) %>% 
  gf_facet_grid(low_ospan ~ condition, labeller = label_both)


gf_jitter(item_test_SCALED ~ weighted, color = ~pretty_condition, data = df_ospan) %>% 
  gf_facet_wrap(~pretty_condition)

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = df_ospan) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

```

**OPSAN score does NOT predict performance in the control condition.**

```{r}

control <- df_ospan %>% filter(condition == "111")

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = control) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

#test performance predicted by OSPAN score
m <- lm(item_test_NABS ~ weighted, data = control)
summary(m)

```

**OPSAN score does NOT predict performance in the control condition.**

```{r}

impasse <- df_ospan %>% filter(condition == "121")

gf_dhistogram(~item_test_NABS , color = ~pretty_condition, data = impasse) %>% 
  gf_facet_wrap(low_ospan ~ pretty_condition, labeller = label_both)

#test performance predicted by OSPAN score
m <- lm(item_test_NABS ~ low_ospan, data = impasse)
summary(m)

```

**EXPLORE**

```{r}

test <- df_ospan %>% filter(condition == "121")
table(test$condition)

m <- lm(item_test_NABS ~ low_ospan, data = test)
summ(m)

gf_boxplot(item_test_NABS ~ low_ospan, data = test)
```

```{r}

#330 total 
# —————————
#126 lab 
#204 online  
  #  71 no ospan 
  # 133 ospan 
# sgc3a_lab <- df_all %>% filter(mode=="lab-synch") %>% filter(study=="SGC3A")
# sgc3a_online <- df_all %>% filter(mode=="asynch") %>% filter(study=="SGC3A")
# sgc3a_online_noospan <- sgc3a_online %>% filter(subject %nin% ospan$subject)
# sgc3a_onspan <- df_subjects
# 
# m <- lm( item_test_NABS ~ condition, data = sgc3a_online_noospan)
# summary(m)
# 
# m2 <- lm(item_test_NABS ~ condition, data = df_subjects)
# summary(m2)
# 
# 
# m3 <- lm(item_test_NABS ~ low_ospan, data = df_subjects)
# summary(m3)
# 


```

```{r}
# library(lpme)
# m <- modereg(Y = df_ospan$DV_percent_NABS, W= df_ospan$condition, 
#         bw = 2, nstart = 2, PLOT = TRUE)
# 
# summary(m)
```

## RESOURCES

```{r}
#| label: session
sessionInfo()
```

```{r}
gf_histogram(~s_NABS, data = sgc3a) %>% gf_facet_wrap(condition~.)
```

## ITEM

```{r}
# #| label: IMPORT-DATA-ITEMS
# #| warning : false
# #| message : false
# 
# # HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
# # mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
# setwd(imac)
# 
# #IMPORT SUBJECTS 
# sgc3a_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") %>% mutate(
#   study = "SGC3A"
# ) 
# sgc4a_items <- read_rds("analysis/SGC4A/data/2-scored-data/sgc4a_scored_items.rds") %>% mutate(
#   study = "SGC4A",
#   pretty_mode = "online-replication"
# ) %>% dplyr::select(-answer)
# 
# sgc4b_items <- read_rds("analysis/SGC4B/data/2-scored-data/sgc4b_scored_items.rds") %>% mutate(
#   pretty_mode = "online-replication"
# ) %>% dplyr::select(-answer)
# sgc5_items <- read_rds("analysis/SGC5A/data/2-scored-data/sgc5_scored_items.rds") %>% mutate(
#   study = "SGC5A",
#   pretty_mode = "online-replication"
# ) %>% dplyr::select(-answer)
# 
# #COMPARE COLUMNS
# janitor::compare_df_cols(sgc3a_items,sgc4a_items, sgc4b_items, sgc5_items)	
# 
# #MERGE 
# df_items <- rbind(sgc3a_items, sgc4a_items, sgc4b_items, sgc5_items)
# 
# #IMPORT OSPAN DATA 
# # ospan <- read_csv("analysis/combined/data/fall21_scored_ospan.csv") %>% mutate(
#   # subject = SUBJECTID
# # )
# 
# #MERGE OSPAN DATA WITH SGC DATA 
# # df_ospan <- df_subjects %>% filter(
# #   subject %in% ospan$subject
# # )
# # 
# # df_ospan <- merge(df_ospan, ospan)
# # 
# # #RENAME for reuse
# # df_all <- df_subjects
# # df_subjects <- df_ospan
# 
# #CLEANUP
# # rm(ospan, sgc3a, sgc4a, sgc4b, sgc5, df_ospan)
# ```
# 
# ```{r}
# library(lme4)
# library(lmerTest)
# library(performance)
# library(report)
# 
# #LINEAR MODEL ON SUBJECT TOTAL SCORE
# m <- lm(s_NABS ~ condition, data = sgc3a)
# summary(m)
# 
# #LINEAR MODEL ON SUBJECT TOTAL SCORE
# m <- glm(score_niceABS~condition, data = sgc3a_items, family = "binomial")
# summary(m)
# performance(m)
# 
# #0 RANDOM INTEREPTS SUBJECT
# m.r0 <- glmer(score_niceABS ~ (1|subject), data = sgc3a_items, family = "binomial")
# summary(m.r0)
# anova(m.r0)
# performance(m.r0)
# 
# #0 RANDOM SLOPE CONDITION/SUBJECT
# m.r1 <- glmer(score_niceABS ~ (0 + condition|subject), data = sgc3a_items, family = "binomial")
# summary(m.r1)
# anova(m.r1)
# performance(m.r1)
# 
# #0 RANDOM SLOPES and intertercepts CONDITION/SUBJECT
# m.r2 <- glmer(score_niceABS ~ (1 + condition|subject), data = sgc3a_items, family = "binomial")
# summary(m.r2)
# anova(m.r2)
# performance(m.r2)
# 
# ####0 RANDOM INTERCEPTS SUBJECT RANDOM INTERCEPTS ITEM
# m.r3 <- glmer(score_niceABS ~ (1 |subject) + (1|q), data = sgc3a_items, family = "binomial")
# summary(m.r3)
# anova(m.r3)
# performance(m.r3)
# 
# r3m <- glmer(score_niceABS ~ condition + (1 |subject) + (1|q), data = sgc3a_items, family = "binomial")
# summary(r3m)
# anova(r3m)
# performance(r3m)
# 
# plot(r3m)
# check_model(r3m)
# report(r3m)
# 
# 
# 
# #BAD FIT
# r4m <- glmer(score_niceABS ~ condition +  q + (1 |subject) , data = sgc3a_items, family = "binomial")
# summary(r4m)
# 
# #???
# r5m <- glmer(score_niceABS ~ condition +  (1|q) + (1|subject) , data = sgc3a_items, family = "binomial")
# summary(r5m)
# 
# coef(r5m)
# 
# # #1. RANDOM SUBJECT + FIXED CONDITION
# # # X + (1 | SUBJECT)
# # m1 <- glmer(score_niceABS ~ condition + (1|subject), data = sgc3a_items, family = "binomial")
# # summary(m1)
# # anova(m1)
# # performance(m1)
# # 
# # #1. RANDOM condition by SUBJECT + FIXED CONDITION
# # # X + (1 + X | SUBJECT)
# # m2 <- glmer(score_niceABS ~ condition + (1+condition|subject), data = sgc3a_items, family = "binomial")
# # summary(m2)
# # lmerTest::anova(m2)
# # performance(m2)
# # 
# # #2. RANDOM SUBJECT + RANDOM Q + FIXED CONDITION 
# # m2 <- glmer(score_niceABS ~ condition + (1|subject) + (1|q) , data = sgc3a_items, family = "binomial")
# # summary(m2)
# # anova(m2)
# # performance(m2)
# 
# ##R SQUARED marginal is fixed effects only
# ##R SQUARED condition is with random effects 
```
