---
subtitle: 'SGCX | Modelling Reference'
---

\newpage

# Modelling Reference {#modelling .unnumbered}

**In this notebook we use data from study SGC3A to explore different modelling techniques and assess their suitability for the bimodal accuracy distributions in the SGC project data.**

```{r}
#| label: SETUP
#| warning : false
#| message : false

library(Hmisc) # %nin% operator
library(jtools) #misc helpers 

library(ggpubr) #arrange plots
library(cowplot) #arrange shift function plots
library(ggformula) #easy graphs
library(vcd) #mosaic plots
library(vcdExtra) #mosaic plots
library(kableExtra) #printing tables 
library(sjPlot) #visualize model coefficients
library(ggeasy) #the way it should be
library(statsExpressions) #expressions to add to plots
library(ggstatsplot) #plots with expressions
library(ggdist) #uncertainty viz 
library(ggeffects) #visualization log regr models
library(modelsummary) #tables
library(broom) #tidy models
library(broom.mixed) #tidy mixed and mixture models

#misc utilities
library(broom)
library(modelr)
library(distributional)
library(report) #easystats reporting
library(see) #easystats visualization
library(performance) #easystats model diagnostics
library(parameters) #viz and compare models
library(equatiomatic) #extract model equation
library(pwr) #power analyses

#modelling
library(rstatix) #helpful basic stats & non parametrics
library(WRS2) #robust and nonparametric tests 
library(lmerTest) #for CIs in glmer 
library(qqplotr) #confint on qq plot
library(gmodels) #contingency table and CHISQR
library(pscl) #zeroinfl / hurdle models 
library(lme4) #mixed effects models 
library(ordinal) #ordinal regression
library(MASS) # polyr ordinal regression
library(VGAM) #censored, zero infl, etc
library(brant) #brant test for ordinal regression
library(nnet) #multinomial logistic regression 
library(mclogit) #multinomial logistic regression
library(bayestestR)

#one package to rule them all
library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

#GRAPH THEMEING
theme_set(theme_minimal()) 

```

```{r}
#| label: IMPORT-DATA
#| warning : false
#| message : false

# HACK WD FOR LOCAL RUNNING?
# imac = "/Users/amyraefox/Code/SGC-Scaffolding_Graph_Comprehension/SGC-X/ANALYSIS/MAIN"
#mbp = "/Users/amyfox/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/ANALYSIS/MAIN"
#setwd(mbp)

#IMPORT DATA 
df_items <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds')
df_subjects <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds')

#PREP DATA 
df_subjects <- df_subjects %>% mutate(
  test_score = item_test_NABS,
  raw_condition = condition,
  condition = pretty_condition
)

df_items <- df_items %>% mutate(
  accuracy = as.factor(score_niceABS),
  scaled = recode_factor(score_SCALED, "-1" = "orth", "-0.5"="unknown","0"="uncertain","0.5"="lines","1"="tri"),
  scaled = as.ordered(scaled),
  raw_condition = condition,
  condition = pretty_condition,
  state = recode_factor(score_SCALED, #for ordinal
                         "-1" = "orth-like",
                         "-0.5" = "unknown",
                         "0" = "unknown",
                         "0.5" = "tri-like",
                         "1" = "tri-like"),
  state = as.ordered(state)
)
# 3 group ordered STATE variable
# 5 group ordered SCALED variable [matches high_interpretation]


```

# INDEPENDENT SAMPLES

-   test-equivalents of GLM models: https://rpubs.com/palday/glm-test
-   models as test :https://lindeloev.github.io/tests-as-linear/

## CONTINUOUS OUTCOME

**Does CONDITION have an effect on TEST PHASE ABSOLUTE SCORE?\
(# questions correct on test phase of task, \[in lab\] participants)**

*(Can also be transformed to proportion or percentage).*

```{r}
#| label: SETUP-IND-TESTSCORE
#| message: false
#| warning: false

#::::::::::::SETUP DATA
df = df_subjects %>% filter(mode == "lab-synch")

#::::::::::::DESCRIPTIVES
mosaic::favstats(test_score ~ condition, data = df)

#::::::::::::VISUALIZE DISTRIBUTIONS

# #GGFORMULA | FACETED HISTOGRAM
# stats = df %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(item_test_NABS))
# gf_props(~item_test_NABS, 
#          fill = ~pretty_condition, data = df) %>% 
#   gf_facet_grid(~pretty_condition) %>% 
#   gf_vline(data = stats, xintercept = ~mean, color = "red") +
#   labs(x = "# Correct",
#        y = "proportion of subjects",
#        title = "Test Phase Absolute Score (# Correct)",
#        subtitle = "") + theme(legend.position = "blank")

##GGPUBR | HIST+DENSITY SCORE 
p <- gghistogram(df, x = "test_score", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("condition")) +
  labs( title = "Distribution of TEST Absolute Score",
        subtitle ="Pattern of response is similar across data collection modes but differs by condition",
        x = "Total Absolute Score (Test Phase)", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")

##VERTICAL RAINCLOUD USING GGDISTR
ggplot(df, aes(x = condition, y = test_score,
                        fill = condition) ) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.1,
    width = 1, 
    point_colour = NA
   ) + 
  geom_boxplot(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=condition, y = test_score),
    width = .15, 
    outlier.shape = NA
  ) + 
  geom_point(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=condition, y = test_score, color = condition),
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .05
  )) + labs( 
    title = "Distribution of TEST Absolute Score ",
    x = "Condition", y = "Total Absolute Score (Test Phase)") +
  theme(legend.position = "blank") + 
  coord_cartesian(xlim = c(0.5, NA), clip = "off")


```

### Independent Samples T-Test (Student's T)

-   Tests null hypothesis that true difference in population mean is == 0
-   Assumes normally distributed variables
-   Assumes equal variance of samples (homogeneity of variance)
-   **SGC accuracy data violate both homogeneity and normality**\*

```{r}
#| label: STUDENTS-T

(t <- t.test( test_score ~ condition,data = df,
              paired = FALSE, var.equal = TRUE, alternative = c("two.sided"))) # less, greater for one sided tests
report(t)

```

```{r}
#STATSPLOT | VIOLIN

#one tailed tests must be done manually by extracting results expression and adding as subtitle
#default is two tailed test
# results <- two_sample_test( data = df, x = pretty_condition, 
#                             y = item_test_NABS,
#                             alternative = "g")

ggbetweenstats(y = test_score, x = condition, data = df,
               type = "parametric", var.equal = TRUE,
               alternative  = "g",
               pairwide.display = "significant", ) 
# + labs(subtitle = results$expression[[1]])

```

### Independent Samples T-Test (Welch's T)

-   Tests null hypothesis that true difference in population mean is == 0
-   Assumes normally distributed variables
-   **Does not** assumes equal variance of samples (homogeneity of variance)
-   ***SGC accuracy data violates normality assumption***

```{r}
#| label: WELCHES-T

(t <- t.test( test_score ~ condition,data = df,
              paired = FALSE, var.equal = FALSE))
report(t)

```

```{r}
#STATSPLOT | VIOLIN
ggbetweenstats(y = test_score, x = condition, data = df,
               type = "parametric", var.equal = FALSE,
               pairwide.display = "significant", )

```

### Yuen's T-Test (Trimmed Means)

-   **Robust alternative to** to t-test is Yuen's t-test which uses trimmed means
-   Trimmed means are not desireable for this research scenario because they trim data from the extremes, and in this study these are true, interesting values
-   <https://garstats.wordpress.com/2017/11/28/trimmed-means/>

```{r}
#| label: YUENS-T

(y <- yuenbt( formula = test_score ~ condition, data = df,
              side = TRUE, EQVAR = FALSE))

```

```{r}
#STATSPLOT | VIOLIN
ggbetweenstats(y = test_score, x = condition, data = df,
               type = "robust", var.equal = FALSE,
               pairwide.display = "significant", )

```

### Wilcoxon Rank Sum (Mann-Whitney Test)

-   **Non parametric alternative** to t-test; compares median rather than mean by ranking data
-   Does not assume normality
-   Does not assume equal variance of samples (homogeneity of variance)
-   **Appropriate for SGC accuracy data**
-   https://www.datanovia.com/en/lessons/wilcoxon-test-in-r/#two-sample-wilcoxon-test
-   https://data.library.virginia.edu/the-wilcoxon-rank-sum-test/

```{r}
#| label: WILCOXON RANK SUM

(w <- wilcox.test(df$test_score ~ df$condition,
                 paired = FALSE, alternative = "two.sided")) #less, greater
report(w)

#effect size
df %>% wilcox_effsize(test_score ~ condition)

```

```{r}
#STATSPLOT | VIOLIN
ggbetweenstats(y = test_score, x = condition, data = df,
               type = "nonparametric", var.equal = FALSE,
               pairwide.display = "significant", )

```

### Floor/Ceiling Corrections + Wilcoxon Rank Sum

-   The bimodal distribution of the subject-level score data do not meet the requirements for t-tests.
-   However, a non-parametric alternative is available (Wilcoxon rank sum test / Man-Whitney test)
-   Additional corrections are available for data with 'floor' and/or 'ceiling' effects via the 'DACR' package
-   https://link.springer.com/article/10.3758/s13428-020-01407-2#Sec14 see also https://qmliu.shinyapps.io/DACFE/

For comparison we run a standard followed by a Wilcoxon rank-sum (Mann-Whitney) test that is a nonparametric alternative for non-normally distributed data.

```{r}

(t <- wilcox.test(df$test_score ~ df$condition))

```

Next, we calculate the t-test and ANOVA (F-test) based on a series of corrections provided for data with floor and/or ceiling effects.

-   <https://link.springer.com/article/10.3758/s13428-020-01407-2#Sec14>

Using properties from truncated normal distributions, the authors propose an easy-to-use method for the *t*-test and ANOVA with ceiling/floor data.

*The proposed method calculates the degrees of freedom based on the after-truncation sample sizes (where l = number of floor observations, and r = number of ceiling observations). The rationale was that the proposed method utilizes full information only from data points of n − r − l participants and partial information from data points of r + l participants of a group for the mean and variance estimation. Specifically, the corrected mean and variance estimates (Eqs. 14 and 15) are functions of mean and variance estimates using after-truncation data (n − r − l participants) and the standardized floor and ceiling threshold estimates. The thresholds are estimated using the ceiling and floor percentage estimates based on data points of n − r and n − l participants, respectively. This is a relatively conservative approach for calculating the degrees of freedom, which can help control the type I error rate. This feature can be beneficial, especially given the "replication crisis" in psychological and behavioral research.*

```{r}

#FLOOR-CEILING ADJUSTED T TESTS
library(DACF) #tests for data with floor and ceiling 
# https://www.rdocumentation.org/packages/DACF/versions/1.0.0

#prepare data [vector of scores for each group]
score_111 <- df %>% filter(condition == "control") %>% dplyr::select(test_score) %>% pull()
score_121 <- df %>% filter(condition == "impasse") %>% dplyr::select(test_score) %>% pull()

# recover the mean and variance for ceiling/floor data
a <- rec.mean.var(score_111) %>% unlist()
# recover the mean and variance for ceiling/floor data
b <- rec.mean.var(score_121) %>% unlist()
r <- as.data.frame(rbind("control"=a ,"impasse"=b))
r

# method "a" uses original sample size
# method "b" uses after-truncation sample size

# perform adjusted t test
lw.t.test(score_111,score_121,"b")

#FLOOR-CEILING ADJUSTED F* TEST ANOVA
lw.f.star(df,test_score~condition,"b")
# method "a" uses original sample size
# method "b" uses after-truncation sample size


```

*The control condition has 11% of data at ceiling and 71% at floor, with corrected mean of -7 and variance of 207. The impasse condition has 17% at ceiling, and only 42% at floor, with with corrected mean of 2.39 and variance respectively as of 58.*

*A corrected t-test t statistic is -4.94, p = 0.05. The estimated Cohen's d is -0.89 with a confidence interval 0f \[-14.22, -5.97\].*

*A correct F-test (ANOVA) has a corrected F-statstic of 5.85, p \< 0.05, and Fsquared effect size of 0.195*

### Linear Regression

-   Assumes homogeneity of variance
-   Assumes normally distributed residuals
-   when default dummy coding is used;
    -   intercept = predicted mean of first group,
    -   predictor coefficient = difference to mean of second group

```{r}
#SCORE predicted by CONDITION
lm.1 <- lm(test_score ~ condition, data = df)
paste("Model")
summary(lm.1)
paste("Partition Variance")
anova(lm.1)
paste("Confidence Interval on Parameter Estimates")
confint(lm.1)
report(lm.1) #sanity check
check_model(lm.1)
```

```{r}
#| label: PLOT-MODEL-COEFSS-GGDIST

#MODEL ESTIMATES WITH UNCERTAINTY

#setup references
m <- lm.1
df <- df
call <- m$call %>% as.character()

# uncertainty model visualization
df  %>%
  modelr::data_grid(condition) %>%
  augment(lm.1, newdata = ., se_fit = TRUE) %>%
  ggplot(aes(y = condition, color = condition)) +
  stat_halfeye( scale = .5,
      aes(
        xdist = dist_student_t(df = df.residual(m), mu = .fitted, sigma = .se.fit),
        fill = stat(cut_cdf_qi(cdf,
                .width = c(.90, .95),
                labels = scales::percent_format())))) +
  scale_fill_brewer(direction = -1) +
  labs (title = "(LAB) Test Phase Accuracy ~ Condition",
        x = "model predicted mean (% correct)", y = "Condition", fill = "Interval",
        subtitle = paste("lm(",call[2],")")
  ) + theme(legend.position = "blank")

```

```{r}
#| label: PLOT-MODEL-PRED-SJPLOT

#sjPlot
p1 <- plot_model(lm.1,  type = "eff", 
           show.data = TRUE, jitter = TRUE,
           show.p = TRUE) 

#BS TO MANUALLY ADD REGRESSION FORMULA AS SUBTITLE 
# library(equatiomatic)
# library(latex2exp)
# (x <- extract_eq(lm.1, use_coefs = TRUE, ital_vars=TRUE, coef_digits = 1, raw_tex = FALSE))
# b = TeX(x)
# p1[["condition"]][["labels"]][["subtitle"]]= expression( paste(widehat(accuracy), " = ", 1.7, " + ", "1.6", "*", condition[impasse]) )
p1

```

### Censored (Tobit) Regression

<https://stats.oarc.ucla.edu/r/dae/tobit-models/>

**For censored data (i.e. truncated axis).** The tobit model, also called a censored regression model, is designed to estimate linear relationships between variables when there is either left- or right-censoring in the dependent variable (also known as censoring from below and above, respectively). Censoring from above takes place when cases with a value at or above some threshold, all take on the value of that threshold, so that the true value might be equal to the threshold, but it might also be higher. In the case of censoring from below, values those that fall at or below some threshold are censored.

-   censored vs truncated : There is sometimes confusion about the difference between truncated data and censored data.
-   With censored variables, all of the observations are in the dataset, but we don't know the "true" values of some of them.
-   With truncation some of the observations are not included in the analysis because of the value of the variable.
-   When a variable is censored, regression models for truncated data provide inconsistent estimates of the parameters. See Long (1997, chapter 7) for a more detailed discussion of problems of using regression models for truncated data to analyze censored data.

```{r}

#set censoring values 
lo = 0
hi = 8 
range(df$test_score)
print("Lo and Hi should equate to upper and lower bounds of the # Qs ")

library(VGAM)

#FIT MODEL
m1<- vglm(test_score ~ condition, tobit(Lower = lo, Upper =hi ), data = df)
summary(m1)

#CONFIDENCE INTERVALS
b <- coef(m1)
se <- sqrt(diag(vcov(m1)))
cbind(LL = b - qnorm(0.975) * se, UL = b + qnorm(0.975) * se)

#TEST FIT
#We can test the significance of condition overall by fitting an empty model and using a likelihood ratio test.
m0 <- vglm(test_score ~ 1, tobit(Lower = lo, Upper = hi), data = df)
(p <- pchisq(2 * (logLik(m1) - logLik(m0)), df = 2, lower.tail = FALSE))
paste("P value of likelihood ratio test less than alpha = 0.05? ", p <0.05)
compare_performance(m0,m1)

#DIAGNOSTICS
plot(m1)

df$yhat <- fitted(m1)[,1]
df$rr <- resid(m1, type = "response")
df$rp <- resid(m1, type = "pearson")[,1]

par(mfcol = c(2, 3))

with(df, {
  plot(yhat, rr, main = "Fitted vs Residuals")
  qqnorm(rr)
  plot(yhat, rp, main = "Fitted vs Pearson Residuals")
  qqnorm(rp)
  plot(test_score, rp, main = "Actual vs Pearson Residuals")
  plot(test_score, yhat, main = "Actual vs Fitted")
})

#VARIANCE ACCOUNTED FOR
print("VARIANCE ACCOUNTED FOR")
# correlation
(r <- with(df, cor(yhat, test_score)))
# variance accounted for
r^2
performance(m1)


#NOTE: censReg package also does Tobit regression [including mixed models]
```

-   The coefficient labeled "(Intercept):1" is the intercept or constant for the model.
-   The coefficient labeled "(Intercept):2" is an ancillary statistic. If we exponentiate this value, we get a statistic that is analogous to the square root of the residual variance in OLS regression. logSigma is the variance of the model (logarithmized) (same as the second intercept in the VGAM output)
-   The predicted value of test_phase_score is 5.75 points *higher* for students in the impasse condition than for students in the control condition. (72% improvement in score!)

**Using censReg package** - https://cran.r-project.org/web/packages/censReg/vignettes/censReg.pdf

```{r}
library(censReg) #censored regression

#FIT MODEL
c1 <- censReg( test_score ~ condition, left=lo, right=hi, data = df )
summary(c1)

#CONFIDENCE INTERVALS
b <- coef(c1)
se <- sqrt(diag(vcov(c1)))
cbind(LL = b - qnorm(0.975) * se, UL = b + qnorm(0.975) * se)

#TEST FIT
#We can test the significance of condition overall by fitting an empty model and using a likelihood ratio test.

c0 <- censReg( test_score ~ 1, left=lo, right=hi, data = df )
(p <- pchisq(2 * (logLik(c1) - logLik(c0)), df = 2, lower.tail = FALSE))
paste("P value of likelihood ratio test less than alpha = 0.05? ", p <0.05)


performance(c1)
```

-   logSigma is the variance of the model (logarithmized) (same as the second intercept in the VGAM output)
-   output should match that of VGAM

## COUNT OUTCOME

**Does CONDITION have an effect on TEST PHASE ABSOLUTE SCORE?\
(# questions correct on test phase of task, \[in lab\] participants)**

The outcome variable absolute score is clearly not normal. As it represents the cumulative number of items a participant has answered correctly, we can try considering it a type of *count*, (i.e. count of the number of questions the participant got correct) and attempt to model it using a General Linear Model with the Poisson distribution (and the default log-link function). Note that the process of answering questions on a test do not seem to strictly match the assumptions of a Poisson process.

```{r}
#| label: SETUP-COUNT-TESTSCORE
#| message: false
#| warning: false

#::::::::::::SETUP DATA
df = df_subjects %>% filter(mode == "lab-synch")

#::::::::::::DESCRIPTIVES
mosaic::favstats(test_score ~ condition, data = df)

#::::::::::::VISUALIZE DISTRIBUTIONS

# #GGFORMULA | FACETED HISTOGRAM
# stats = df %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(item_test_NABS))
# gf_props(~item_test_NABS, 
#          fill = ~pretty_condition, data = df) %>% 
#   gf_facet_grid(~pretty_condition) %>% 
#   gf_vline(data = stats, xintercept = ~mean, color = "red") +
#   labs(x = "# Correct",
#        y = "proportion of subjects",
#        title = "Test Phase Absolute Score (# Correct)",
#        subtitle = "") + theme(legend.position = "blank")

##GGPUBR | HIST+DENSITY SCORE 
p <- gghistogram(df, x = "test_score", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("condition")) +
  labs( title = "Distribution of TEST Absolute Score",
        subtitle ="Pattern of response is similar across data collection modes but differs by condition",
        x = "Total Absolute Score (Test Phase)", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")

##VERTICAL RAINCLOUD USING GGDISTR
ggplot(df, aes(x = condition, y = test_score,
                        fill = condition) ) + 
  ggdist::stat_halfeye(
    side = "left",
    justification = 1.1,
    width = 1, 
    point_colour = NA
   ) + 
  geom_boxplot(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=condition, y = test_score),
    width = .15, 
    outlier.shape = NA
  ) + 
  geom_point(
    inherit.aes = FALSE, #supress fill
    mapping = aes(x=condition, y = test_score, color = condition),
    size = 1.3,
    alpha = .3,
    position = position_jitter( 
      seed = 1, width = .05
  )) + labs( 
    title = "Distribution of TEST Absolute Score ",
    x = "Condition", y = "Total Absolute Score (Test Phase)") +
  theme(legend.position = "blank") + 
  coord_cartesian(xlim = c(0.5, NA), clip = "off")

```

### Poisson Regression

<https://stats.oarc.ucla.edu/r/dae/poisson-regression/>\
*General Linear model using the Poisson distribution*

```{r}

#SCORE predicted by CONDITION --> POISSON DISTRIBUTION
p.1 <- glm(test_score ~ condition, data = df, family = "poisson")
paste("Model")
summ(p.1)
paste("Partition Variance")
anova(p.1)
paste("Confidence Interval on Parameter Estimates")
confint(p.1)
report(p.1) #sanity check
plot_model(p.1)
check_model(p.1)

```

*The variable condition has a coefficient of 0.67, (p \< 0.005). This means that for the impasse condition, the expected log count \# of questions increases by 0.67. By exponentiating the estimate we see that \# question correct rate for the impasse condition is 1.95x that of the control condition. However, model diagnostics suggest the residuals are not normally distributed.*

### Negative Binomial Regression

<https://stats.oarc.ucla.edu/r/dae/negative-binomial-regression/> -*overdispersed count data (variance much greater than mean)*

-   similar to Poisson regression, but using the negative binomial distribution, which can better account for 'overdispersed' data where variance is much greater than the mean

```{r}
#NEGATIVE BIONOMIAL REGRESSION

library(MASS)

#fit model 
nb.1 <- glm.nb(test_score ~ condition, data = df)

#check overdispersion need 
#assumes conditional means are not equal to conditional variances
#conduct likelihood ration test to compare and test [need poisson]
m.t <- glm(test_score ~ condition, family = "poisson", data = df)
#pchisq(2 * (logLik(nb.1) - logLik(m.t)), df = 1, lower.tail = FALSE)
#A large (+) log likelihood suggests that the negative binomial is more appropriate than the Poisson model
test_lrt(m.t, nb.1)


#EXPONENTIATE PARAMETER ESTIMATES
est <- cbind(Estimate = coef(nb.1), confint(nb.1))
#exponentiate parameter estimates
print("Exponentiated Estimates")
exp(est)

summary(nb.1)
report(nb.1)
plot_model(nb.1)
check_model(nb.1)
```

*The variable condition has a coefficient of 0.67, (p \< 0.005). This means that for the impasse condition, the expected log count \# of questions increases by 0.67. By exponentiating the estimate we see that \# question correct rate for the impasse condition is nearly 1.95x that of the control condition. However, model diagnostics suggest the residuals are not normally distributed.*

```{r}
#COMPARE POISSON AND NEGATIVE BINOMIAL
compare_performance(p.1, nb.1)
test_lrt(p.1, nb.1)
```

*AIC, Pseudo-R2 and a likelihood ratio test indicate that the negative binomial regression model are a better fit for the distribution of test-phase scores.*

## MIXTURE MODELS

### Zero Inflated Poisson

<https://stats.oarc.ucla.edu/r/dae/zip/>\
*A Poisson count process with excess zeros*

-   The Zero-Inflated model allows specification of two models (a mixture) where *some* of the zeros are included in the DGP for the Poisson model, while only the 'excess' zeros are included in the DGP for the Binomial model \[the zero-inflated part\]
-   The model includes:
    -   A logistic model to model which of the two processes the zero outcome is associated with
    -   A poisson model to model the count process
-   Can specify different predictors for each part of the model
-   predictors after the \| are for the binomial zero-inflated part of the model, while those infront are for the poisson process

```{r}
#ZERO INFLATED POISSON

zinfp.1 <- zeroinfl(test_score ~  condition| condition , data = df)
summary(zinfp.1)
report(zinfp.1)
performance(zinfp.1)
#check_model(zinfp.1)

```

*In the count model, the coefficient for the condition is not significant.*

*In the zero-inflation model, the coefficient for the condition variable is -1.23 and statistically significant. This suggests that the log odds of being an excessive zero decrease by 1.23 if you are in the impasse condition*

### Zero Inflated Negative Binomial Regression

<https://stats.oarc.ucla.edu/r/dae/zinb/>\
*count data that are overdispersed and have excess zeros*

Zero-inflated negative binomial regression is for modelling count variables with excessive zeros, and especially when the count data are overdispersed (mean is much larger than variance). It can help account for situations where theory suggests that excess zeros are generated by 2 separate processes, one that includes the other count values, and the other that is just the zeros, and thus that the *excess* zeros can be modelled independently.

Total Absolute Score (# items correct) may fit this situation, as the data are overdispersed (variance much greater than the mean) and there are are very large number of zeros. It is theoretically plausible that these excess zeros (no answers correct) are the result of a different 'process' ... (i.e) little understanding and/or resistance to restructuring understanding of the coordinate system. However, I am not certain if it is plausible to suggest that the zeros themselves are the result of two different processes: (ie. perhaps trying to understand, and not trying to understand?) \<- this could maybe be disentangled by first question latency?

The model includes:

-   A logistic model to model which of the two processes the zero outcome is associated with
-   A negative binomial model to model the count process

```{r}
#| label: MODEL-TESTABS-ZINFNEGBINOM

library(pscl) #  for zeroinfl negbinomial

#ZERO INFLATED NEGATIVE BINOMIAL
zinb.1 <- zeroinfl(test_score ~ condition | condition , 
                   data = df, dist = "negbin")
#before the | is the count part, after the | is the logit model
paste("Model")
summary(zinb.1)
report(zinb.1)
performance(zinb.1)

#   rootogram(zinb.1)



# #EXPONENTIATE PARAMETER ESTIMATES
# est <- cbind(Estimate = coef(zinb.1), confint(zinb.1))
# #exponentiate parameter estimates
# print("Exponentiated Estimates")
# exp(est)

```

*In the count model, the coefficient for the condition is very small, and not significant (suggesting it does not contribute to the count yielding process?).*

*In the zero-inflation model, the coefficient for the condition variable is -1.056 and statistically significant. This suggests that the log odds of being an excessive zero decrease by 1.06 if you are in the impasse condition*

```{r}
compare_performance(zinfp.1, zinb.1)
test_likelihoodratio(zinfp.1, zinb.1)
```

**TODO come back to this and discuss further**\

### Hurdle Model

-   <https://data.library.virginia.edu/getting-started-with-hurdle-models/>
-   <https://en.wikipedia.org/wiki/Hurdle_model#:~:text=A%20hurdle%20model%20is%20a,of%20the%20non%2Dzero%20values.>

class of models for count data with both overdispersion and excess zeros;\
different from zero-inflated models where the excess zeros are theorized to arise from two *different* processes; in the hurdle model, there is a separate model for P(x=0) and a separate model for P(x!=0)

The model includes:

-   A binary logit model to model whether the observation takes a positive count or not. (1) Does the student get *any* questions right?

-   a truncated Poisson or Negative binomial model that only fits positive counts (2) How many questions does the student get right?

```{r}
library(pscl) #zero-inf and hurdle models 
library(countreg) #rootogram
#install.packages("countreg", repos="http://R-Forge.R-project.org")

#SYNTAX OUTCOME ~ count model predictor | hurdle predictor

h.1 <- pscl::hurdle(test_score ~ condition | condition , data = df,
              zero.dist = "binomial", dist = "poisson", size = 8)

h.2 <- pscl::hurdle(test_score ~ condition | condition , data = df,
              zero.dist = "binomial", dist = "negbin", size = 8)

summary(h.1)
summary(h.2)


rootogram(h.1)
rootogram(h.2)
plot(compare_performance(h.1,h.2))
test_lrt(h.1, h.2)

```

## BETA REGRESSION

### BETA Distribution

Beta regression on % correct (with standard transformation for including \[0,1\])

<https://stats.stackexchange.com/questions/63350/how-to-interpret-the-coefficients-from-a-beta-regression>

https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/#2-fractional-logistic-regression

If a test score is conceptualized as correct (a) and b (incorrect) score = correct / (correct + incorrect) = a/ (a + b)

In the beta distribution, a and b are the shape parameters!

```{r}
theme_clean <- function() {
  theme_minimal(base_family = "Barlow Semi Condensed") +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
          axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
          strip.text = element_text(family = "BarlowSemiCondensed-Bold",
                                    size = rel(1), hjust = 0),
          strip.background = element_rect(fill = "grey80", color = NA))
}


ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 6, shape2 = 4),
                aes(color = "Beta(shape1 = 6, shape2 = 4)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom")+ labs(
    title = "Beta a = 6, b = 4"
  )


ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 8, shape2 = 0.001),
                aes(color = "Beta(shape1 = 8, shape2 = 0.001)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "ALL CORRECT"
  )

ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 0.001, shape2 = 8),
                aes(color = "Beta(shape1 = 0.001, shape2 = 8)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "ALL WRONG"
  )

ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 4, shape2 = 4),
                aes(color = "Beta(shape1 = 4, shape2 = 4)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "FIGURED IT OUT HALFWAY "
  )

```

ALTERNATIVE conceptualization of a and b is as mean MU and precision PHI (same idea as variance).

a = mu \* phi b = (1 - mu) \* phi

MU = a / (a+b) \[the score\] PHI = a + b \[ the total \# items \]

The `extraDistr::dprop` function can also handle mu and phi parameters, named as mean and size

```{r}
library(extraDistr)
ggplot() +
  geom_function(fun = dprop, args = list(mean = 0.999, size = 8),
                aes(color = "Beta(mean = 0.99, size = 8)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "ALL CORRECT"
  )

ggplot() +
  geom_function(fun = dprop, args = list(mean = 0.001, size = 8),
                aes(color = "Beta(mean = 0.001, size = 8)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "ALL WRONG"
  )

ggplot() +
  geom_function(fun = dprop, args = list(mean = 0.5, size = 8),
                aes(color = "Beta(mean = 0.5, size = 8)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom") + labs(
    "FIGURED IT OUT HALFWAY "
  )


ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 0.001, shape2 = 1),
                aes(color = "Beta(shape1 = 6, shape2 = 4)"),
                size = 1) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_clean() +
  theme(legend.position = "bottom")


```

**Note that in all cases, we have to slightly adjust the parameters because the distribution can't handle values at 0 or 1. So we scrunch the data in just a bit.**

### BETA Regression

https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/#2-fractional-logistic-regression

The syntax is just like all other formula-based regression functions, but with an added bit: the first part of the equation (x \~ y) models the mean, or MU, while anything that comes after a \| in the formula will explain variation in the precision, or PHI.

The link parameter determines the link function for the mu estimate?.

#### Setup

```{r}
# 
library(betareg)

#RESCLAE VARIABLE
#beta reg can't handle 0s and 1s 
sub <- df_subjects %>% dplyr::select(condition, DV_percent_NABS)
n = nrow(sub) %>% unlist()
sub$dv_transformed = (sub$DV_percent_NABS * (n-1) + 0.5)/n
 
#VISUALIZE VARIABLES
paste("Slightly squished data")
histogram(sub$dv_transformed)


#FIT MODEL
mb1 <- betareg(dv_transformed ~ condition | condition, 
              data = sub)

summary(mb1)
plot(mb1)


```

-   Which link function should be used for each? We now have two sets of coefficients, one set for each parameter (the mean and precision). The parameters for the mean are measured on the logit scale, just like with logistic regression previously, and we can calculate the marginal effect of having a quota by using plogis() and piecing together the coefficient and the intercept:

#### Interpret

```{r}

beta_mu_intercept <- mb1 %>% 
  tidy() %>% 
  filter(component == "mean", term == "(Intercept)") %>% 
  pull(estimate)

beta_mu_condition <- mb1 %>% 
  tidy() %>% 
  filter(component == "mean", term == "conditionimpasse") %>% 
  pull(estimate)


print("Coefficients —- PROBABILITIES")
print("[these should math pred plots]")
print("Probability at reference level")
plogis(beta_mu_intercept)
print("Probability at coefficient level")
plogis(beta_mu_intercept + beta_mu_condition)
print("CHANGE in probability by coefficient")
plogis(beta_mu_intercept + beta_mu_condition) - plogis(beta_mu_intercept)

plot_model(mb1, type = "pred")
```

Impasse increases the average of the distribution of accuracy by 12% percentage points, on average. This should match the value we get with fractional regression. (logistic regression with count variable instead of binomial)

-   The intercept of MU indicates the average of the distribution of score proportion in the control condition
-   The coefficient of MU indicates the change in the average distribution of score proportion in the impasse condition

The PHI parameter estimates are not measured on a logit scale. Instead, they're log values. We can invert them by exponentiating them with exp(). - The phi intercept indicates the precision of the distribution of score proportion in the control condition - The phi coefficient indicates the change in the precision of the distribution of score proportion in the impasse condition

-   CONDITION is a significant predictor of MEAN (location) but not PHI (precision). Put another way, the shape of both distributions is similar, but the impasse condition distribution is shift to the right (higher) than the control distribution.



#### Compare

```{r}
# COMPARE MODELS


#PREP DATA 
df <- df_subjects %>% filter(mode == "lab-synch")
n = nrow(df) %>% unlist()
df <- df %>%  mutate(
    accuracy = item_test_NABS/8,
    test_trans = (accuracy * (n-1) + 0.5)/n,
    all_trans = (DV_percent_NABS * (n-1) + 0.5)/n
  )

# LINEAR REGRESSION
lm1 <- lm( test_trans ~ pretty_condition    , data = df)
# summary(lm1)

# LINEAR REGRESSION on LOGTRANSFORMED
qlm1 <- lm( qlogis(test_trans) ~ pretty_condition    , data = df)
# summary(qlm1)

# FRACTIONAL REGRESSION
#uses quasibinomial [like binomial] on proportion 
#instead of binomial data. coeffs on logit scale
flm1 <- glm( test_trans ~ pretty_condition, data = df, family = quasibinomial())
summary(flm1)

# BETA REGRESSION predictor phi
bm1 <- betareg(test_trans ~ pretty_condition | pretty_condition, 
              data = df, link = "logit")
summary(bm1)


models = list("LM" = lm1, 
              "LM log transform" = qlm1, 
              "GLM Quasibinomial" = flm1, 
              "beta" = bm1)
modelsummary(models)

```


```{r}

#GET ESTIMATES
model_beta <- bm1

beta_mu_intercept <- model_beta %>% 
  tidy() %>% 
  filter(component == "mean", term == "(Intercept)") %>% 
  pull(estimate)

beta_mu_condition <- model_beta %>% 
  tidy() %>% 
  filter(component == "mean", term == "pretty_conditionimpasse") %>% 
  pull(estimate)

beta_phi_intercept <- model_beta %>% 
  tidy() %>% 
  filter(component == "precision", term == "(Intercept)") %>% 
  pull(estimate)

beta_phi_condition <- model_beta %>% 
  tidy() %>% 
  filter(component == "precision", term == "pretty_conditionimpasse") %>% 
  pull(estimate)


mu_control = plogis(beta_mu_intercept)
phi_control = plogis(beta_phi_intercept)
#when you don’t explicitly model the precision, the resulting coefficient in the table isn’t on the log scale—it’s a regular non-logged number, so there’s no need to exponentiate

#PLOT ESTIMATES

control_title <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                         "), size = exp(", round(beta_phi_intercept, 2), "))")

impasse_title <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                      " + ", round(beta_mu_condition, 2), 
                      "), size = exp(", round(beta_phi_intercept, 2),
                      " + ", round(beta_phi_condition, 2), "))")

ggplot(data = tibble(x = 0:1), aes(x = x)) +
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept), 
                            mean = plogis(beta_mu_intercept)),
                aes(color = control_title)) + 
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept + beta_phi_condition), 
                            mean = plogis(beta_mu_intercept + beta_mu_condition)),
                aes(color = impasse_title)) +
  geom_density(data = df, 
               aes(x = test_trans, fill = pretty_condition), 
               alpha = 0.5, color = NA) +
  labs(x = "Beta Regression Predictions") +
  theme_clean() +
  theme(legend.position = "right")

```

We can also convert the parameter estimates back to A and B. for the impasse condition, MU = 0.397 SIZE = 0.666

a = mu \* phi = 0.397 \* 0.66 = 0.264 b = (1 - mu) \* phi = (1 - 0.397) \* 0.666 = 0.402

we should expect it to look more like ... (for 50% ) a = correct = 3 b = incorrect = 10 a + b = total = 13

MU = a / (a+b) \[the score\] = 3 / 13 = 0.23 PHI = a + b \[ the total \# items \] = 13

mu = 0.277 = a / (a + b) phi = 0.395 = a + b let b = 13

```{r}
#RECOVER A AND B FROM MU AND PHI
muphi_to_shapes <- function(mu, phi) {
  shape1 <- mu * phi
  shape2 <- (1 - mu) * phi
  return(list(shape1 = shape1, shape2 = shape2))
}

shapes_to_muphi <- function(shape1, shape2) {
  mu <- shape1 / (shape1 + shape2)
  phi <- shape1 + shape2
  return(list(mu = mu, phi = phi))
}

paste("model predicts control A and B ")
muphi_to_shapes(mu_control, phi_control)
paste("should be something like ")
shapes_to_muphi(5,13)
```

It seems like the although the (shrinkled) beta regression model is working, it is really messing up the estimate of the PHI parameter, such that we can't even recover the A and B. I expect whole numbers, and instead I'm geting values less than 1. I think this might be because the phi value is so low (note that phi = precision, so like inverse of variance. Lower values are more less precise (more variance)). Since I have so many values near 0 and near 1 (even though I've shrunk the estimates to be in the open interval (0,1) rather than \[0,1\], I think it causing problems with the estimation. )

#### bayesian version
https://www.bayesrulesbook.com/

```{r}
library(brms)
model_beta_bayes <- brm(
  bf(test_trans ~ pretty_condition,
     phi ~ pretty_condition),
  data = df,
  family = Beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  # Use the cmdstanr backend for Stan because it's faster and more modern than
  # the default rstan You need to install the cmdstanr package first
  # (https://mc-stan.org/cmdstanr/) and then run cmdstanr::install_cmdstan() to
  # install cmdstan on your computer.
  backend = "cmdstanr"
)
```

```{r}
summary(model_beta_bayes)
paste("Cool! parameters are almost identical to betareg")
```

```{r}
library(tidybayes)
posterior_beta <- model_beta_bayes %>% 
  gather_draws(`b_.*`, regex = TRUE) %>% 
  mutate(component = ifelse(str_detect(.variable, "phi_"), "Precision", "Mean"),
         intercept = str_detect(.variable, "Intercept"))

ggplot(posterior_beta, aes(x = .value, y = fct_rev(.variable), fill = component)) +
  geom_vline(xintercept = 0) +
  stat_halfeye(aes(slab_alpha = intercept), 
               .width = c(0.8, 0.95), point_interval = "median_hdi") +
  scale_fill_viridis_d(option = "viridis", end = 0.6) +
  scale_slab_alpha_discrete(range = c(1, 0.4)) +
  guides(fill = "none", slab_alpha = "none") +
  labs(x = "Coefficient", y = "Variable",
       caption = "80% and 95% credible intervals shown in black; RAW COEFF") +
  facet_wrap(vars(component), ncol = 1, scales = "free_y") +
  theme_clean()



##GET USEFUL ESTIMATES
paste("TRANSFORMED ESTIMATES")
model_beta_bayes %>% 
  spread_draws(`b_.*`, regex = TRUE) %>% 
  mutate(across(starts_with("b_phi"), ~exp(.))) %>%
  mutate(across((!starts_with(".") & !starts_with("b_phi")), ~plogis(.))) %>%
  gather_variables() %>% 
  median_hdi()
# now phi is a raw value indicating precision
# now mu is an odds ratio for median? of the distribution
```

#### Bayesian predictions

```{r}

# Plug a dataset where quota is FALSE and TRUE into the model
beta_bayes_pred <- model_beta_bayes %>% 
  epred_draws(newdata = tibble(pretty_condition = c("control", "impasse")))

ggplot(beta_bayes_pred, 
       aes(x = .epred, y = pretty_condition, fill = pretty_condition)) +
  stat_halfeye(.width = c(0.8, 0.95), point_interval = "median_hdi") +
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  guides(fill = "none") +
  labs(x = "Predicted Comprehension Task Score Percentage Correct", y = NULL,
       caption = "80% and 95% credible intervals shown in black") +
  theme_clean()

```
```{r}
#TODO POSTERIOR PREDICTIVE CHECKS 

# describe_posterior(model_beta_bayes, test = c("p_direction", "rope", "bayesfactor"))

# library(rstanarm)
# pp_check(model_beta_bayes, plotfun = "boxplot", nreps = 10, notch = FALSE)
# pp_check(model_beta_bayes, plotfun = "stat_grouped", stat = "median", group = "pretty_condition")

```
### ZERO INFLATED BETA 

1. logistic regression for zero inflated $\alpha$ (logit scale)
2. beta regression for proportion $\mu$ (logit scale) and $\phi$ (log scale)

```{r}
df <- df_subjects %>% dplyr::select(pretty_condition, DV_percent_NABS) %>% 
  mutate(
    condition = pretty_condition,
    accuracy = DV_percent_NABS
  )
#shrink accuracy just a tiny bit from 1 for testing purposes 
library(tidyfst)
df <- df %>% mutate_when(accuracy == 1, accuracy = 0.99)

#INTERCEPT ONLY model for alpha
zinb <- brm(
  bf(accuracy ~ condition, #mean of (0,1) values, mu
     phi ~ condition, #precision of (0,1) values, phi
     zi ~ condition), #zero inflated is extreme? alpha
  data = df,
  family = zero_inflated_beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file ="analysis/utils/models/zinb.rds"
)
```
```{r}
summary(zinb)
performance(zinb)
# plot(rope(zinb))
```


```{r}

tidy(zinb, effects = "fixed")


paste("INTERPRET ALPHA")
zi_intercept <- tidy(zinb, effects = "fixed") %>% 
  filter(component == "zi", term == "(Intercept)") %>% 
  pull(estimate)

# Logit scale intercept
zi_intercept
## b_zi_Intercept 
##          0.525

# Transformed to a probability/proportion
plogis(zi_intercept)
## b_zi_Intercept 
##         0.628

nrow(df %>% filter(accuracy == 0 & condition == "control"))/nrow(df %>% filter(condition == "control"))

```
The value of the z_intercept should be the same as the proportion of 0s in the control condition. 

VISUALIZE ZERO INFLATED COMPONENT
```{r}
beta_zi_pred_int <- zinb %>% 
  predicted_draws(newdata = tibble(condition = c("control", "impasse"))) %>% 
  mutate(is_zero = .prediction == 0,
         .prediction = ifelse(is_zero, .prediction - 0.01, .prediction))

ggplot(beta_zi_pred_int, aes(x = .prediction)) +
  geom_histogram(aes(fill = is_zero), binwidth = 0.025, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0) +
  scale_fill_viridis_d(option = "plasma", end = 0.5,
                       guide = guide_legend(reverse = TRUE)) +
  labs(x = "Predicted proportion of Correct Responses", 
       y = "Count", fill = "Is zero?") +
  facet_wrap(vars(condition), ncol = 2,
             labeller = labeller(condition = c(`control` = "Control", 
                                           `impasse` = "Impasse"))) + 
  theme_clean() +
  theme(legend.position = "bottom")
```


### ZERO-ONE INFLATED BETA Regression

https://sometimesir.com/posts/2019-02-18-analyze-analog-scale-ratings-with-zero-one-inflated-beta-models/#zoib-regression
https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/#2-fractional-logistic-regression

**TODO** 
- explore priors
- explore random effects
- explore non-beta distribution; once the 1s and zeros are gone, do we need the beta? 

A zero-inflated beta regression allows us to model a separate data generating process for the 0s and 1s and the rest.

A ZOINB is a mixture of:

1.  A logistic regression model that predicts if an outcome is either 0 or 1 or not 0 or 1, defined by (or alternatively, a model that predicts if outcomes are extreme (0 or 1) or not (between 0 and 1); "IS EXTREME" model; defined by $\alpha$

2.  A logistic regression model that predicts if any of the 0 or 1 outcomes are actually 1s, defined by (or alternatively, a model that predicts if the extreme values are 1) "WHICH EXTREME" model defined by $\gamma$

3.  A beta regression model that predicts if an outcome is between 0 and 1 if it's not zero or not one, defined by and (or alternatively, a model that predicts the non-extreme (0 or 1) values) "THE REST" model $\mu$ and $\phi$

Can also do mixed! (but no reason do this here, as the item level data is not beta reg )

```{r}

#DATA SETUP
# sub <- df_subjects %>% dplyr::select(condition, DV_percent_NABS) 
# don't need to transform anymore
df <- df_subjects %>% filter(mode == "lab-synch") %>% 
  mutate(
    condition = pretty_condition,
    scaled = (s_SCALED + 13)/26,  #transform scaled onto 0-1 scale
    accuracy = item_test_NABS/8
  ) %>% dplyr::select(condition, scaled, accuracy)


#INTERCEPT ONLY model for alpha
zoinb <- brm(
  bf(accuracy ~ condition, #mean of (0,1) values, mu
     phi ~ condition, #precision of (0,1) values, phi
     zoi ~ condition, #is extreme? alpha
     coi ~ condition), #is 1? gamma
  data = df,
  family = zero_one_inflated_beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file ="analysis/utils/models/zoinb.rds"
)


zoinb_scaled <- brm(
  bf(scaled ~ condition, #mean of (0,1) values, mu
     phi ~ condition, #precision of (0,1) values, phi
     zoi ~ condition, #is extreme? alpha
     coi ~ condition), #is 1? gamma
  data = df,
  family = zero_one_inflated_beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file ="analysis/utils/models/zoinb_scaled.rds"
)

```
```{r}
summary(zoinb)

plot(rope(zoinb))
plot_model(zoinb)
plot(rope(zoinb_scaled))
plot_model(zoinb_scaled)
```

More plots 

```{r}

tidy(zoinb, effects = "fixed")

paste("INTERPRET ALPHA")
zoi_intercept <- tidy(zoinb, effects = "fixed") %>% 
  filter(component == "cond", term == "zoi_(Intercept)") %>% 
  pull(estimate)

# Logit scale intercept
zoi_intercept

# Transformed to a probability/proportion
plogis(zoi_intercept)

nrow(df %>% filter(accuracy %in% c(0,1) & condition == "control"))/nrow(df %>% filter(condition == "control"))

```
The value of the zoi_intercept should be the same as the proportion of 0s and 1s (vs non 0s and 1s) in the control condition... and it does!

VISUALIZE ZERO INFLATED COMPONENT
```{r}
beta_zoi_pred_int <- zoinb %>% 
  predicted_draws(newdata = tibble(condition = c("control", "impasse"))) %>% 
  mutate(is_extreme = .prediction %in% c(0,1),
         is_one = .prediction ==1, 
         .prediction = ifelse(is_extreme & !is_one, .prediction - 0.01, .prediction),
         .prediction = ifelse(is_one, .prediction + 0.01, .prediction))

ggplot(beta_zoi_pred_int, aes(x = .prediction)) +
  geom_histogram(aes(fill = is_extreme), binwidth = 0.05, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 1) +
  scale_fill_viridis_d(option = "plasma", end = 0.5,
                       guide = guide_legend(reverse = TRUE)) +
  labs(x = "Predicted proportion of Correct Responses", 
       y = "Count", fill = "Is extreme?") +
  facet_wrap(vars(condition), ncol = 2,
             labeller = labeller(condition = c(`control` = "Control", 
                                           `impasse` = "Impasse"))) + 
  theme_clean() +
  theme(legend.position = "bottom")
```
```{r}
beta_zoi_pred_int <- zoinb_scaled %>% 
  predicted_draws(newdata = tibble(condition = c("control", "impasse"))) %>% 
  mutate(is_extreme = .prediction %in% c(0,1),
         is_one = .prediction ==1, 
         .prediction = ifelse(is_extreme & !is_one, .prediction - 0.01, .prediction),
         .prediction = ifelse(is_one, .prediction + 0.01, .prediction))

ggplot(beta_zoi_pred_int, aes(x = .prediction)) +
  geom_histogram(aes(fill = is_extreme), binwidth = 0.05, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 1) +
  scale_fill_viridis_d(option = "plasma", end = 0.5,
                       guide = guide_legend(reverse = TRUE)) +
  labs(x = "Predicted proportion of Correct Responses", 
       y = "Count", fill = "Is extreme?") +
  facet_wrap(vars(condition), ncol = 2,
             labeller = labeller(condition = c(`control` = "Control", 
                                           `impasse` = "Impasse"))) + 
  theme_clean() +
  theme(legend.position = "bottom")

```


```{r}
pp_check(zoinb, draws = 8000 )

```

### HURDLE + NEG BINOM (bayesian)

Rather than transform score to a proportion correct, keep it in native count form, and model as negative binomial. 
hurdle model : 

(1) A logistic regression model that predicts if an outcome is 0 or not (this is the hurdle part)
(2) A lognormal (hurdle_lognormal()), gamma (hurdle_gamma()), Poisson (hurdle_poisson()), or negative binomial (hurdle_negbinomial()) model for outcomes that are not zero


```{r}

#SETUP DATA 
df <- df_subjects %>% filter(mode == "lab-synch") %>% 
  mutate(
    score = item_test_NABS,
    condition = pretty_condition
  )

#PLOT
gf_histogram(~score, fill = ~ condition, data = df) %>% gf_facet_wrap(~condition)

#FIT MODEL
hln <- brm(
  bf(score ~ condition,
     hu ~ condition),
  data = df,
  family = hurdle_lognormal(),
  chains = 4, iter = 2000, warmup = 1000, seed = 1234,
  silent = 2, cores = 4, 
  backend = "cmdstanr",
  file ="analysis/utils/models/hln.rds"
)

hnb <- brm(
  bf(score ~ condition,
     hu ~ condition),
  data = df,
  family = hurdle_negbinomial(),
  chains = 4, iter = 2000, warmup = 1000, seed = 1234,
  silent = 2, cores = 4, 
  backend = "cmdstanr",
  file ="analysis/utils/models/hnb.rds"
)

hpoi <- brm(
  bf(score ~ condition,
     hu ~ condition),
  data = df,
  family = hurdle_poisson(),
  chains = 4, iter = 2000, warmup = 1000, seed = 1234,
  silent = 2, cores = 4, 
  backend = "cmdstanr",
  file ="analysis/utils/models/hpoi.rds"
)


summary(hln)
summary(hnb)
summary(hpoi)
plot_model(hln)
plot_model(hnb)
plot_model(hpoi)


compare_performance(hln, hnb, hpoi)

# Exponential
p1 <- pp_check(hpoi) + labs(title="Poisson Hurdle")
p2 <- pp_check(hnb) + labs(title="NegBinom Hurdle")
p3 <- pp_check(hln) + labs(title="LogNorm Hurdle")
cowplot::plot_grid( p1 + p2 + p3)

# Logged
pred <- posterior_predict(hpoi)
bayesplot::ppc_dens_overlay(y = log1p(df$score), 
                            yrep = log1p(pred[1:10,])) + 
  labs(title = "Posterior Predictive Checks")
```
```{r}
pred_gdp_hurdle <- hpoi |> 
  predicted_draws(newdata = tibble(condition = c("control","impasse"))) |>
  mutate(is_zero = .prediction == 0,
         .prediction = ifelse(is_zero, .prediction - 0.1, .prediction))

ggplot(pred_gdp_hurdle, aes(x = .prediction)) +
  geom_histogram(aes(fill = is_zero), binwidth = 1, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0) +
  labs(x = "test phase correct", y = "Count", fill = "Is zero?",
       title = "Predicted number correct from hurdle model") +
  # coord_cartesian(xlim = c(-2500, 75000)) +
  theme_nice() +
  theme(legend.position = "bottom")

```
This model is making predictions outside the bounds of the data (beyond 8)
**TODO**
https://jsdajournal.springeropen.com/articles/10.1186/s40488-021-00121-4
consider motivation of hurdle vs zero inflated 

### MIXED ZERO-ONE-INFLATED BETA Regression

Next we'll explore doing a beta regression on a proportion transformed scaled score at the item level, as a zero one inflated mixed beta reg 
```{r}
df <- df_items %>% filter(mode == "lab-synch") %>% 
  filter(q %nin% c(6,9)) %>% 
  mutate(scaled = (score_SCALED + 1)/2,
         condition = pretty_condition) %>% 
  dplyr::select(subject, scaled, condition)


zoinb_scaled <- brm(
  bf(scaled ~ condition + (1|subject) , #mean of (0,1) values, mu
     phi ~ condition , #precision of (0,1) values, phi
     zoi ~ condition , #is extreme? alpha
     coi ~ condition ), #is 1? gamma
  data = df,
  family = zero_one_inflated_beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file ="analysis/utils/models/zoinb_scaled_2.rds"
)

summary(zoinb_scaled)
plot_model(zoinb_scaled)
```


### BETA HURDLE Regression (Frequentist)

<https://github.com/markhwhiteii/beta_hurdle/blob/master/manuscript/beta_hurdle.pdf>

-   MU tells if mean is different by condition
-   SIGMA tells if variance is different by condition
-   NU coefficient tells if condition yields different probability at floor
-   TAU coefficient tells if condition yields different probability at ceiling

```{r}
#BETA HURDLE REGRESSION EXAMPLE
library(gamlss)

#CREATE SAMPLE DATA 
n <- 5000 
mu <- 0.40 
sigma <- 0.60 
p0 <- 0.13 
p1 <- 0.17 
p2 <- 1- p0- p1
a <- mu * (1- sigma ^ 2) / (sigma ^ 2) 
b <- a * (1- mu) / mu

#CREATE DIST
set.seed(1839) 
y <- rbeta(n, a, b) 
cat <- sample(1:3, n, prob = c(p0, p2, p1), replace = TRUE) 
y[cat == 1] <- 0 
y[cat == 3] <- 1

#VISUALIZE DISTRIBUTION
x <- as.data.frame(y)
gf_histogram(~x$y)
#this looks not unlike my distribution! 

#CREATE AN EMPTY MODEL
fit <- gamlss( formula = y ~ 1, # formula for mu 
               formula.sigma = ~ 1, # formula for sigma 
               formula.nu = ~ 1, # formula for nu 
               formula.tau = ~ 1, # formula for tau 
               family = BEINF() )

summary(fit)
plot(fit)

#TRANSFORM PARAMETRS BACK 
inv_logit <- function(x) exp(x) / (1 + exp(x)) # inverse of link function
fit_mu <- inv_logit(fit$mu.coefficients) 
paste("MU: ",fit_mu)
fit_sigma <- inv_logit(fit$sigma.coefficients) 
paste("SIGMA: ",fit_sigma)
fit_nu <- exp(fit$nu.coefficients) 
fit_tau <- exp(fit$tau.coefficients) 
fit_p0 <- fit_nu / (1 + fit_nu + fit_tau) 
paste("P0: ",fit_p0)
fit_p1 <- fit_tau / (1 + fit_nu + fit_tau)
paste("P1: ",fit_p1)

```

**BETA HURDLE INTERPRETATION** - beta component\
- MU "location" (mean)\
- SIGMA "scale" (positively related to variance; variance = sigma.squared *mean* (1-mean)\
- Rigby, Stasinopoulos, Heller, and De Bastiani (2017) "reparameterized" the beta distribution so that the two parameters determining the shape of the distribution would be more useful in a regression framework (see Ferrari & Cribari-Neto, 2004 for a different parameterization)

**ZERO-ONE HURDLE COMPONENT**\
- The two additional parameters, ν NU and τTAU , are related to p0 and p1, respectively.\
- p0 is the probability that a case equals 0,\
- p1 is the probability that a case equals 1,\
- p2 (i.e., 1 −p0 −p1) is the probability that the case comes from the beta distribution

```{r}

#MY DATA
#SETUP DATA 

min = 0 #min possible value of scale
max = 13 #max possible value of scale

library(mosaic) #for shuffling
#1. Rescale accuracy using 
# recommended adjustment 
#rescaled = value-min/(max-min)
df <- df_subjects %>% mutate(
  accuracy = s_NABS,
  R_acc = (accuracy-min)/(max-min), #as %
  T_acc = (accuracy * (nrow(df)-1) + 0.5)/nrow(df)/max, #transform for no 0 and 1
  # perm = shuffle(condition),
  scaffold_rt = item_scaffold_rt
) %>% dplyr::select(accuracy,R_acc, T_acc, condition,scaffold_rt)

#VISUALIZE DISTRIBUTION
gf_histogram(~R_acc, fill = ~condition, data = df) %>% gf_facet_wrap(~condition) + labs(title = "Histogram of accuracy")

#VISUALIZE DISTRIBUTION
gf_histogram(~T_acc, fill = ~condition, data = df) %>% gf_facet_wrap(~condition) + labs(title = "Histogram of [rescaled] accuracy")

# gf_histogram(~R_acc, fill = ~perm, data = df) %>% gf_facet_wrap(~perm) + labs(title = "Histogram of shuffled accuracy")

#SUMMARIZE SAMPLE
paste("Grand mean", mean(df$R_acc))

library(mosaic)
stats = favstats(df$R_acc ~ df$condition)
stats$mean <- mean(df$R_acc ~ df$condition)
stats$var <- var(df$R_acc ~ df$condition)
print("Grand stats")
stats 
print("P0")
nrow(df %>% filter(R_acc ==0))/nrow(df)
print("P1")
nrow(df %>% filter(R_acc ==1))/nrow(df)

#CREATE MODEL

#CREATE AN EMPTY MODEL
m0 <- gamlss( formula = R_acc ~ 1, # formula for mu 
              formula.sigma =  ~ 1, # formula for sigma 
              formula.nu =  ~ 1, # formula for nu 
              formula.tau =  ~ 1, # formula for tau 
              family = BEINF(), data = df )

m0 <- gamlss(R_acc ~ 1, ~ 1, ~ 1, ~ 1, 
            data = df, family = BEINF())
paste("THE EMPTY MODEL")
summary(m0)
plot(m0)

#TRANSFORM PARAMETRS BACK 
inv_logit <- function(x) exp(x) / (1 + exp(x)) # inverse of link function
m0_mu <- inv_logit(m0$mu.coefficients) 
paste("MU: ",m0_mu)
m0_sigma <- inv_logit(m0$sigma.coefficients) 
paste("SIGMA: ",m0_sigma)
m0_nu <- exp(m0$nu.coefficients) 
paste("NU: ",m0_nu)
m0_tau <- exp(m0$tau.coefficients) 
paste("TAU: ",m0_tau)
m0_p0 <- fit_nu / (1 + fit_nu + fit_tau) 
paste("P0: ",m0_p0)
m0_p1 <- fit_tau / (1 + fit_nu + fit_tau)
paste("P1: ",m0_p1)



#CREATE PREDICTOR MODEL
m1 <- gamlss(R_acc ~ condition, 
             ~ condition, 
             ~ condition, 
             ~ condition, 
            data = df, family = BEINF())
paste("THE PREDICTOR MODEL")
summary(m1)

#LOOKING PREDICTOR MODEL
m <- gamlss(R_acc ~ condition , 
            ~ condition , 
            ~ condition , 
            ~ condition , 
            data = df, family = BEINF())
paste("THE PREDICTOR MODEL")
summary(m)


#CREATE PREDICTOR MODEL ON SHUFFLED [PERMUTATION TEST]
# mperm <- gamlss(R_acc ~ perm, ~ perm, ~ perm, ~ perm, 
#             data = df, family = BEINF())
# summary(mperm)

#sanity check with scaled outcome, no zeros ones
# m3 <- gamlss(T_acc ~ condition, ~ condition, ~ condition, ~ condition, 
#             data = df, family = BEINF())
# summary(m3)
#m3 shouldn't show condition as significant for nu and tau, because T_acc was scaled to not include any 0s and 1s

#investigate beta negative binomial distribution
#https://en.wikipedia.org/wiki/Beta_negative_binomial_distribution

#TRANSFORM PARAMETRS BACK 
inv_logit <- function(x) exp(x) / (1 + exp(x)) # inverse of link function
m1_mu <- inv_logit(m1$mu.coefficients) 
paste("MU: ",m1_mu)
m1_sigma <- inv_logit(m0$sigma.coefficients) 
paste("SIGMA: ",m1_sigma)
m1_nu <- exp(m1$nu.coefficients) 
paste("NU: ",m1_nu)
m1_tau <- exp(m1$tau.coefficients) 
paste("TAU: ",m1_tau)

summary(m)
plot(m)

```

-   MU tells if mean is different by condition
-   SIGMA tells if variance is different by condition
-   NU coefficient tells if condition yields different probability at floor
-   TAU coefficient tells if condition yields different probability at ceiling

## BINOMIAL OUTCOME

```{r}
#| label: SETUP-Q1-ACCURACY

#PREPARE DATA 
df <- df_items %>% filter(q ==1) %>% filter(mode == "lab-synch")
# %>% mutate(
#   accuracy = recode_factor(score_niceABS, "0" ="incorrect","1"="correct"),
#   scaled = as.ordered(score_SCALED),
#   q = as.factor(q),
#   high_interpretation = as.factor(high_interpretation)
# )

#GROUPED PROPORTIONAL BAR CHART
gf_props(~accuracy, fill = ~pretty_condition, x =~pretty_condition,
       position = position_dodge(), data = df) %>%
  gf_facet_grid(~pretty_mode) +
   labs(x = "Question 1 Accuracy",
       title = "Accuracy on Q1",
       subtitle="")

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = accuracy)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
   labs(#y = "",
       title = "Accuracy on Test Phase",
       x = "Condition",
       fill = "",
       subtitle="Impasse Condition yields a greater proportion of correct responses")

```

#### CHI SQUARE

```{r}
#| label : CHISQR-Q1

#::::::::::::CROSSTABLE
# CrossTable( x = df$condition, y = df$accuracy, 
#              fisher = TRUE, chisq=TRUE, expected = TRUE, sresid = TRUE)

#::::::::::::MOSAIC PLOT
# note: blue indicates cell count higher than expected, 
# red indicates cell count less than expected; under null hypothesis
# mosaicplot(main="Accuracy on First Question by Condition",
#             data = df, pretty_condition ~ accuracy, 
#             shade = T)

#::::::::::::TABLE
df %>% sjtab( fun = "xtab", var.labels=c("accuracy", "pretty_condition"),
        show.row.prc=F, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T,
        statistics = c("auto"))


#::::::::::::BAR PLOT
ggbarstats(data = df, x = accuracy, y = condition,
           type = "nonparametric")

#::::::::::::CHISQR TEST
(x <- stats::chisq.test(x = df$accuracy, y = df$condition, simulate.p.value = T))


#::::::::::::POWER ANALYSIS
(po <- pwr.chisq.test( w = 0.1, df=(2-1), N = nrow(df), sig.level = 0.05))
```

**A Chi-Square test of independence of Q1 accuracy \[correct,incorrect\] by condition indicates the question accuracy is not dependent on condition. *However, this test may be underpowered, as with the given sample size it has only 20% power to detect a small effect (w = 0.1)***

### LOGISTIC REGRESSION

-   http://post8000.svmiller.com/lab-scripts/logistic-regression-lab.html

*Fit a logistic regression (at the subject-item level), predicting Q1 accuracy (absolute score) by condition.*\
*note: this example uses the combined dataset rather than lab-only, as learning notes were done with the combined and I don't want to recalcualte all the marginal probabilities by hand for learning purposes.*

-   Fit a logistic regression predicting accuracy (absolute score) (n = r nrow(df)) by condition. (k = 2).
-   Parameter estimate: $\beta_{0}$ = Log Odds of (correct) responses in CONTROL condition
    -   $e^{\beta_{0}}$ = ODDS of correct response in CONTROL condition
-   Parameter estimate: $\beta_{1}$ = $\beta_{1impasse}$ Log Odds (Log OR; change in odds for correct response in impasse (vs) control \[log scale\])
    -   $e^{\beta_{1}}$ = ODDS RATIO of correct response in IMPASSE (vs) CONTROL
-   Null hypothesis:$\beta_{impasse} \le 0$ the odds for a correct response does not change, or decreases
-   Alternative hypothesis: $\beta_{impasse} \gt 0$ the odds of a correct response increases

#### Fit Model

*First, we fit a logistic regression with condition as predictor, and compare its fit to an empty (intercept-only) model.*

```{r}
#| label: FIT-Q1ACC-LOG-combined
#| warning: false
#| message: false

#combined dataset, not lab only
df <- df_items %>% filter(q==1) %>% mutate(
  accuracy = as.factor(score_niceABS)
)

# FREQUENCY TABLE
# my.table <- table(df$accuracy, df$pretty_condition)
# addmargins(my.table) #counts
# addmargins(prop.table(my.table)) #props

# MODEL FITTING:::::::::::::::::::::::::::::::::::::

#: 1 EMPTY MODEL baseline glm model intercept only
m0 = glm(accuracy ~ 1, data = df, family = "binomial")
print("EMPTY MODEL")
summary(m0)

#: 2 CONDITION model
m1 <- glm( accuracy ~ pretty_condition, data = df, family = "binomial")
print("PREDICTOR MODEL")
summary(m1)

#: 3 TEST SUPERIOR FIT
paste("AIC wth predictor is lower than empty model?", m0$aic > m1$aic)
test_lrt(m0,m1) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(m0,m1))$p[2])
```

*The Condition predictor significantly improves model fit.*

#### Learning Notes

```{r}
#| label: MODEL-Q1ACC-LOG-combined

# DESCRIBE MODEL ::::::::::::::::::::::::::::::::::::: 

print("PREDICTOR MODEL")
summary(m1)

#: INTERPRET COEFFICIENTS
print("Coefficients —- LOG ODDS")
confint(m1)
print("Coefficients —- ODDS RATIOS")
(e <- cbind( exp(coef(m1)), exp(confint(m1)))) #exponentiate)

print("Coefficients —- PROBABILITIES")
print("[these should math pred plots]")
print("Probability at reference level")
plogis(m1$coefficients[[1]])
print("Probability at coefficient level")
plogis(m1$coefficients[[1]] + m1$coefficients[[2]])
print("CHANGE in probability by coefficient")
plogis(m1$coefficients[[1]] + m1$coefficients[[2]]) - plogis(m1$coefficients[[1]])

print("MODEL PREDICTIONS")
# Retrieve predictions as probabilities 
# (for each level of the predictor)
pred.control <- predict(m1,data.frame(pretty_condition="control"),type="response")
#this should match : plogis(intercept coefficient)
paste("Probability of success in control,", pred.control)
pred.impasse <- predict(m1,data.frame(pretty_condition="impasse"),type="response")
#this should match : plogis(intercept coefficient + predictor coeff)
paste("Probability of success in impasse,", pred.impasse)


# MANUAL ONE-SIDED SIGTEST ::::::::::::::::::::::::::: 

# one-sided (right tail) z test for B COEFFICIENT
#https://stats.stackexchange.com/questions/330655/strategy-for-a-one-sided-test-of-glms-coefficients
#https://www.ibm.com/support/pages/can-one-get-one-tailed-tests-logistic-regression-dividing-significance-levels-half

#SANITY CHECK 2-tailed test should match the model output
tt <- 2*pnorm(summary(m1)$coefficients[2,3], lower.tail = F)
paste("p value for two-tailed test, null B = 0 : ",round(tt,3))
ot <- pnorm(summary(m1)$coefficients[2,3], lower.tail = F)
paste("BUT we want a one tailed directional, null: B <= 0: ",round(ot,3))
paste("adjusted confint for directional hypothesis")
(dcint <- confint(m1, level = 0.90)) # get 90% for right side))
# https://stats.stackexchange.com/questions/20734/is-a-1-sided-90-prediction-interval-equivalent-to-a-2-sided-95-prediction-inte


#:::::::: INTERPRET COEFFICIENTS [directional]

# print("Confidence Interval —- LOG ODDS")
# confint(m1) #not adjusted for 1-tailed
# print("Coefficients —- ODDS RATIOS")
# (e <- cbind( exp(coef(m1)), exp(confint(m1)))) #exponentiated, not adjusted
# (e <- cbind( exp(coef(m1)), exp(dcint))) #exponentiated, adjusted
```

**Understanding the logistic regression model**

```{r}
#::::::::::::DESCRIPTIVES

table(df$accuracy, df$condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns


table(df$accuracy, df$condition) %>% addmargins(2) %>% #display sum for row
      addmargins(1) #sanity check sum of columns
```

*The logistic regression intercept gives the log odds of the outcome for the reference level of the predictor variable*

*The logistic regression coefficients give the change in the log odds of the outcome for a one unit increase in the predictor variable.*

**\[the empty model**

-   The intercept of an empty model (glm(accuracy \~ 1) is equal to log(p/(1-p)), where p = the overall probability of a correct response (df\$accuracy ==1 ).
-   In SGC3A Q1 accuracy this = 71 correct / 330 = 0.215 -\> log(0.215 / (1-0.215)) = -1.29.
-   In other words, the intercept from the model with no predictor variables is the estimated log odds of a correct response for the whole sample.
-   We can also transform the log of the odds back to a probability: p = ODDS/ (1+ODDS) = exp(-1.29)/(1+exp(-1.29)) = 0.215. This should matched the prediction of the empty model

**\[a dichotomous predictor\]**

natural log (odds of +) = -1.822 + 0.901(x1) ; x1 = 0 for control, 1 for impasse - b0 intercept is ODDS OF CORRECT RESPONSE IN REFERENCE (control) - b1 intercept is ODDS RATIO (difference in odds) FOR CORRECT RESPONSE IN IMPASSE

-   INTERCEPT: log odds of (+ response) in control condition
    -   log odds of (+) in control : -1.822 + 0.9(0) = -1.822
    -   convert to odds by exponentiating the coefficients\
        log odds of (+) in control = exp(-1.822) = 0.162 odds
    -   convert to probability by formula =\>\
        p(+) = odds / (1+odds) = 0.162 / (1 + 0.162) = 0.139\
        probability of (+) in control = \~14%
-   B1 COEFFICIENT: DIFFERENCE in log odds of (+) in impasse vs. control
    -   log odds of (+) in impasse: -1.822 + 0.901 = -0.921
    -   convert to odds by exponentiating log odds\
        log odds (+) in impasse = exp(-0.921) = 0.398
    -   convert to probability by formula =\>\
        p(+) = odds / (1 + odds) = 0.398 / (1+0.398) = 0.285\
        probaility of (+) in impasse = \~ 29%
-   ODDS RATIO : exponentiated B1 COEFFICIENT
    -   B1 = (slope of logit model = difference in log odds = log odds ratio
    -   B1 = 0.901 is log odds ratio of (+) in impasse vs control
    -   exp(b1) = exp(0.901) = 2.46
    -   Ratio of odds in impasse are 2.46 times higher than in control. Bein in the impasse condition yields odds athat are 2.46 X higher than in control.

+:----------------------------------------------------------------------+
| MARGINAL\                                                             |
| total = 330 success : 71, failure : 259\                              |
| p(+) = 71 / 330 = 0.215 = 22%\                                        |
| odds(+) = 71 / 259 = 0.274                                            |
+-----------------------------------------------------------------------+
| CONTROL total = 158 success = 22; failure = 136\                      |
| p(+) = 22/158 = 0.139 = 14%\                                          |
| odds(+) = 22/136 = 0.162                                              |
+-----------------------------------------------------------------------+
| IMPASSE total = 172 success = 49; failure = 123\                      |
| p(+) = 49/172 = 0.285 = 29%\                                          |
| odds(+) = 49/123 = 0.398                                              |
+-----------------------------------------------------------------------+

#### Visualize

```{r}
#:::::::: PLOT

#GGSTATS | MODEL | LOG ODDS 
ggcoefstats(m1, output = "plot", 
              conf.level = 0.90) + 
  labs(x = "Log Odds Estimate", 
       subtitle = "p is for two tailed test")

#SJPLOT | MODEL | ODDS RATIO
#library(sjPlot)
plot_model(m1, vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.1, #manually adjust to account for directional test
           ci.lvl = 0.90 ) + #manually adjusted for directional test   
  labs(title = "Model Predicted Odds Ratio",
       subtitle = "",
       x = "Condition")

#SJPLOT | MODEL | PROBABILITIES
plot_model(m1, type="eff",
           show.intercept = TRUE,
           show.values = TRUE,
           title = "Model Predicted Probability of Accuracy",
           axis.title = c("Condition","Probability of Accurate Response"))


#SJPLOT | MODEL | TABLE
# tab_model(m1)

#MODELSUMMARY | MODEL | PROBABILITIES
# modelsummary(m1, exponentiate = TRUE)

#MARGINALEFFECTS | MODEL | PROBABILITIES
library(marginaleffects)
#plot adjusted predictions, conditional on the CONDITION variable using the plot_cap
#interactions plot_cap(mod, condition = c("length", "style"))
plot_cap(m1, condition = "pretty_condition")
# The plot_cme function can be used to draw “Conditional Marginal Effects.” 
#use for ixns 
# plot_cme(m1, effect = "var1", condition = "pretty_condition")

```

#### Diagnostics

```{r}
print("SANITY CHECK REPORTING")
report(m1)

print("MODEL PERFORMANCE")
performance(m1)

print("MODEL DIAGNOSTICS")
check_model(m1)
binned_residuals(m1)

# logitgof(df$accuracy, m1$fitted.values, ord=FALSE)
# hoslem.test(x = df$accuracy, y = fitted(m1), g =  2)
# p should be non significant

```

#### Inference

We fit a logistic regression model to analyze the effect of experimental condition on probability of a correct answer on the first question. In this model, the effect of condition is statistically significant (z = 3.16, p = 0.0016). The model predicts that the odds of a correct response on the first question in the *impasse condition* increase by 146% ($e^{beta_1}$ = 2.46, 95% CI \[1.42, 4.37\]) over the *control condition*.

*Equivalent statements:*

-   being in impasse condition increases log odds of correct response by 0.901 (over control)
-   being in impasse increases odds of correct response in impasse over control increases by factor of 2.46
-   probability of correct response in control predicted as 28.5%, vs only 14% in control condition

```{r}
#PRETTY TABLE SJPLOT
tab_model(m1)
```

## ORDINAL OUTCOME

**Does CONDITION affect the Q1 \[ordered\] type of response given?**

```{r}
#| label : SETUP-INDEPENDENT-ORDINAL

#SETUP DATA
df <- df_items %>% filter(q==1)  %>% filter(mode == "lab-synch") 
#scaled has already been set as an ordered factor of score_SCALED

#::::::::::::DESCRIPTIVES
prop.table(table(df$scaled, df$condition))

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_props(~scaled, fill= ~condition, data = df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       fill = scaled)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~pretty_mode) + 
   labs(#y = "",
       title = "Q1 Interpretation",
       x = "Condition",
       fill = "",
       subtitle="")

```

**Does CONDITION affect the Q1 \[ordered\] state of understanding?**

```{r}
#| label : SETUP-INDEPENDENT-ORDINAL-STATE

#SETUP DATA
df <- df_items %>% filter(q==1)  %>% filter(mode == "lab-synch") 
#scaled has already been set as an ordered factor of score_SCALED

#::::::::::::DESCRIPTIVES
prop.table(table(df$state, df$condition))

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_props(~state, fill= ~condition, data = df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       fill = state)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~pretty_mode) + 
   labs(#y = "",
       title = "Q1 State",
       x = "Condition",
       fill = "",
       subtitle="")
```

### ORDINAL REGRESSION --- Cumulative Link; Proportional Odds

*Fit an ordinal logistic regression (at the subject level), predicting Q1 interpretation by condition.*

-   <https://stats.oarc.ucla.edu/r/faq/ologit-coefficients/>
-   <https://journals.sagepub.com/doi/full/10.1177/2515245918823199>
-   todo see ordinal regression video: <https://www.youtube.com/watch?v=rPcMcW25PPA&ab_channel=NCRMUK>
-   <https://peopleanalytics-regression-book.org/ord-reg.html>
-   <https://medium.com/evangelinelee/brant-test-for-proportional-odds-in-r-b0b373a93aa2>
-   <https://github.com/runehaubo/ordinal/blob/master/old_vignettes/clm_tutorial.pdf>

**Learning Notes**

-   proportional odds regression models effectively act as a series of stratified binomial models under the assumption that the 'slope' of the logistic function of each stratified model is the same.
-   thus need to verify proportional odds assumption - ordinal regression requires an **proportional odds assumption** (the same slope holds for each equation)
-   this is required because the model simultaneously estimates k-1 equations, but each equation has the same *slope*, with different intercepts.
-   conversely, a multinomial (categorical) model will have different slopes as well as intercepts - the intercepts are always ordered in size alpha 1 \< alpha 2 \< alpha k-1...
-   **TODO** - see difference between the three types of ordinal models (adjacent category (vs) cumulative proportions; check Agresti book

#### Fit Model

```{r}

#::::::::::::ORDINAL REGRESSION MODELS

#EMPTY MODEL
paste("EMPTY Ordinal regression of q1 SCALED score (ordered interpretation)")
om.0 <- clm(state ~ 1 , data = df)
# summary(om.0)

#PREDICTOR MODEL
paste("Ordinal regression of q1 SCALED score (ordered interpretation)")
om <- clm(state ~ condition, data = df)
# summary(om)

#COMPARE EMPTY AND PREDICTOR
test_lrt(om.0, om)


#::::::::: EQUIVALENT APPROACH USING POLYR 

# #MODEL
m <- polr(state ~ condition , data = df, Hess=TRUE)
# summary(m)

#exponentiate coefficients and CIs
# (ctable <- coef(summary(m)))
# (p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2)
# (ctable <- cbind(ctable, "p value" = p))
# (ci <- confint(m))
# (e <- coef(m))

```

*Likelihood ratio test suggests the predictor model is a better fit than the empty (intercept only) model.*

#### Inference

```{r}

paste("SUMMARY")
summary(om)

#LOG ODDS
paste("IN LOG ODDS")
(ctable <- coef(summary(om)))
(ci <- confint(om)) 

paste("IN ODDS RATIOS")
#ODDS RATIOS
exp(coef(om))
exp(ci)

```

**Overall, participants in the impasse condition had higher odds (6.68 X as likely) to offer *more correct* interpretations than those in the control condition (z = 4.74, p \< 0.001).**

-   we see the estimates for the 2 intercepts, which are sometimes called cutpoints.
-   The intercepts indicate where the latent variable is cut to make the three groups that we observe in our data.
-   Note that this latent variable is continuous. In general, these are not used in the interpretation of the results.
-   The cutpoints are closely related to thresholds, which are reported by other statistical packages.
-   for k groups there will be k-1 intercepts (cutpoints)
-   confirm that the CI does not include 0 (the units are ordered logits \[ordered log odds\])
-   as with logistic regression we exponentiate the coefficients and confints to get odds ratio

#### Visualize Model

```{r}
# sjPlot::tab_model(om)
sjPlot::plot_model(om)
sjPlot::plot_model(om, type = "eff")
                   # show.data = TRUE, jitter = TRUE)

```

#### Diagnostics

```{r}

#:: ASSESS FIT
performance(om)

# #test proporitional odds assumption 
brant(m) #only works for polyr type model not clm type model
# # A p-value of less than 0.05 on this test—particularly on the Omnibus plus at least one of the variables—should be interpreted as a failure of the proportional odds assumption.

#test proportional odds assumption
library(pomcheckr)
# https://cran.r-project.org/web/packages/pomcheckr/pomcheckr.pdf
(p <- pomcheck( scaled ~ condition , data = df))
plot(p)

```

**The output of the graphical test for proportional odds assumption suggests that the proportional odds assumption may be unreasonable for this dataset. Also, inspecting the output table, we see the coefficients for each level of the scaled variable are quite different.**

Thus, an alternative approach may be more appropriate:

-   Baseline logistic model. This model is the same as the multinomial regression model covered in the previous chapter, using the lowest ordinal value as the reference.

-   Adjacent-category logistic model. This model compares each level of the ordinal variable to the next highest level, and it is a constrained version of the baseline logistic model. The brglm2 package in R offers a function bracl() for calculating an adjacent category logistic model.

-   Continuation-ratio logistic model. This model compares each level of the ordinal variable to all lower levels. This can be modeled using binary logistic regression techniques, but new variables need to be constructed from the data set to allow this. The R package rms has a function cr.setup() which is a utility for preparing an outcome variable for a continuation ratio model.

\*Note: for multiple regression, the ordinal package offers a parameter (nominal = \~predictors) that allow you to designate some predictors as nominal rather than ordinal. But this is not appropriate for this use case.\_

## CATEGORICAL OUTCOME

**Does CONDITION affect the Q1 \[ordered\] type of response given?**

```{r}

#SETUP DATA
df <- df_items %>% filter(q==1)  %>% filter(mode == "lab-synch") 
#scaled has already been set as an ordered factor of score_SCALED

#::::::::::::DESCRIPTIVES
prop.table(table(df$scaled, df$condition))

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_props(~scaled, fill= ~condition, data = df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       fill = scaled)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~pretty_mode) + 
   labs(#y = "",
       title = "Q1 Interpretation",
       x = "Condition",
       fill = "",
       subtitle="")

```

**Does CONDITION affect the Q1 \[ordered\] state of understanding?**

```{r}

#SETUP DATA
df <- df_items %>% filter(q==1)  %>% filter(mode == "lab-synch") 
#scaled has already been set as an ordered factor of score_SCALED

#::::::::::::DESCRIPTIVES
table(df$state, df$condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns

#::::::::::::VISUALIZE DISTRIBUTIONS
gf_props(~state, fill= ~condition, data = df) %>% 
  gf_facet_grid(condition ~ .) + easy_remove_legend()

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       fill = state)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~pretty_mode) + 
   labs(#y = "",
       title = "Q1 State",
       x = "Condition",
       fill = "",
       subtitle="")

#GGSTATSPLOT
ggbarstats(df, state, condition)

```

### CHI SQUARE

**Does CONDITION affect the Q1 \[categorical\] type of response given?**

```{r}

CrossTable( x = df$condition, y = df$scaled, 
             fisher = TRUE, chisq=TRUE, expected = TRUE, sresid = TRUE)
```

```{r}

#MOSAIC PLOT
#note: blue indicates cell count higher than expected, red indicates cell count less than expected; under null hypothesis
mosaicplot(main="Interpretation on First Question by Condition",
            data = df, condition ~ scaled, 
            shade = T)

#::::::::::::WITH STATISTICS
ggbarstats(data = df, x = condition, y = scaled,
           type = "nonparametric") 
```

```{r}
#| label : CHISQR-interpreation

#MOSAIC PLOT
#note: blue indicates cell count higher than expected, red indicates cell count less than expected; under null hypothesis
mosaicplot(main="Response Type on First Question by Condition",
            data = df, condition ~ scaled, 
            shade = T)

CrossTable( x = df$condition, y = df$scaled, 
            fisher = TRUE, chisq=TRUE, expected = TRUE, sresid = TRUE)

df %>%
  sjtab(fun = "xtab", var.labels=c("scaled", "pretty_condition"),
        show.row.prc=T, show.col.prc=T, show.summary=T, show.exp=T, show.legend=T,
        statistics = "fisher")
```

**TODO INFERENCE**

### MULTINOMIAL REGRESSION

on interpreting coefficients (marginal probability)
- https://data.princeton.edu/wws509/stata/mlogit

**TODO** RECONCILE actual predictions with coefficients and visualization of predictions

*Does condition affect the response state of Q1?*

-   <https://peopleanalytics-regression-book.org/multinomial-logistic-regression-for-nominal-category-outcomes.html#running-a-multinomial-regression-model>
-   <https://bookdown.org/chua/ber642_advanced_regression/multinomial-logistic-regression.html>
-   <https://www.youtube.com/watch?v=JcCBIPqcwFo&list=PLzv58M2GAfm50X_Twskr1aXaV5qMuIszx&ab_channel=NCRMUK>
-   Can use nnet package multinom() or mclogit package mblogit() \["baseline logit model"\] or brms with family = "categorical"

*Fit a logistic regression predicting interpretation (k=#response categories) by condition. (k = 2).*

-   2 equations will be estimated (# categories - 1); each representing the odds of answering in that particular interpretation (vs) the reference category (orthogonal answer)

-   (from experience, seems) Each cell must have at least one observation (if one cell is blank, then it seems to be incorrectly estimated, see predicting high_interpretation vs. state. suggests nonsignificant OR for 'unknown' category, when infact that difference drives the effect

-   For *each* equation:

    -   $\beta_{0}$ *= Log Odds of \[this category type vs. reference category type) response in CONTROL condition*

        -   $e^{\beta_{0}}$ *= ODDS of \[this category type vs. reference category type\] response in CONTROL condition*

    -   $\beta_{1}$ *=* $\beta_{1impasse}$ *Log Odds (Log OR; change in odds for \[this category\] type response in impasse (vs) control \[log scale\])*

    -   $e^{\beta_{1}}$ *= ODDS RATIO of \[this. vs reference category type\] response in IMPASSE (vs) CONTROL*

    -   *Null hypothesis:* $\beta_{impasse} \le 0$ *the odds for \[this category of response vs. reference\] does not change, or decreases*

    -   *Alternative hypothesis:* $\beta_{impasse} \gt 0$ *the odds of \[this category of response vs. reference\] increases*

    -   

```{r}
#::::::::::::DESCRIPTIVES RESPONSE STATE

table(df$state, df$condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns


(t <- table(df$state, df$condition) %>% addmargins(2) %>% #display sum for row
      addmargins(1)) #sanity check sum of columns

```

#### Fit Model

```{r}

library(nnet)

#check reference level 
levels(df$state)

#FIT EMPTY MODEL
catm.0 <- multinom(state ~ 1, data = df)
# summary(catm.0)

#FIT PREDICTOR MODEL
catm <- multinom(formula = state ~ condition, data = df, model = TRUE)
# summary(catm)

#COMPARE MODEL FIT
test_lrt(catm.0, catm)

##compare bayesian version
#library(brms)
# bm1 <- brm( state ~ pretty_condition, data = df, family = "categorical")
# summary(bm1)
# plot_model(bm1)
# report(bm1)
# coefficient estimates are very simliar to catm. super cool!

##compare mclogit version
#"baseline-category logit model
# https://www.elff.eu/software/mclogit/manual/mblogit/
# blm1 <- mblogit(state ~ pretty_condition , data = df)
# summary(blm1)
#identical to catm. super cool!

```

*Likelihood ratio test indicates predictor model is significantly better fit to the sample data than the empty (intercept only) model.*

#### Interpretation

```{r}

#::::::::INTERPRETATION
paste("MODEL SUMMARY")
summary(catm)

# calculate z-statistics of coefficients
(z_stats <- summary(catm)$coefficients/summary(catm)$standard.errors)
# convert to p-values
p_values <- (1 - pnorm(abs(z_stats)))*2
# display p-values in transposed data frame
p_values <- data.frame(p = (p_values))
# display odds ratios in transposed data frame

paste("ODDS RATIOS")
odds_ratios <- data.frame(OR = exp(summary(catm)$coefficients))

# options(scipen = 3)
(results <- cbind(odds_ratios, p_values))
```

**Learning Notes**

-   Model estimates encompass two equations:
-   effect of predictor on log odds of being in \[unknown\] instead of reference category \[orth-like\]
-   effect of predictor on log odds of being in \[tri-like\] instead of reference category \[orth-like\]
-   \[need to to double check interpretation, but I *think* that the OR intercepts converted to probabilities equate to the marginal probability of being in each state in the \[reference\] control condition. which makes sense. I think\]
-   IF I change reference category for condition... then the intercepts should no longer be significant. The b1 coefficients should still be significant, but with changed sign (much less likely) \[Yup! this works!\]

#### Inference

-   Being in the IMPASSE condition increases the odds of giving 'unknown/uncertain' response rather than an orthogonal (or satisficing) response by a factor of 50 (z = 3.59, p \< 0.001 ).
-   Being in the IMPASSE condition increases the odds of giving an 'triangular or line-driven' response rather than an orthogonal (or satisficing) response by a factor of 6 (z = 3.98, p \<0.001 )

#### Visualize

```{r}
plot_model(catm, vline.color = 'red')
plot_model(catm, type = "eff")

```

#### Diagnostics

```{r}

#EXAMINE PREDICTIONS
#create sample data frame
test <- data.frame(condition = c("control", "impasse"))
pred <- predict(catm, newdata = test, "probs")
paste("Predicted Probability of Being in Each State")
(cbind(test, pred))

#performance
performance(catm)
DescTools::PseudoR2(catm, which = c("McFadden", "CoxSnell", "Nagelkerke"))

#General Goodness of Fit
# library(generalhoslem)
# logitgof(df$state, catm$fitted.values, g = 3)
# hoslem.test(x = df$state, y = catm$fitted.values, g =  10)
#A non-significant p value indicates that there is no evidence that the observed and expected frequencies differ (i.e., evidence of good fit).
#don't fret! this version of the hoslem-lem test is problematic with fewer than 10 input variables
```

# MIXED (Repeated Measures)

```{r}
#| label: SETUP-TEST-ACC-ITEM

#PREPARE DATA 
n_items = 8 #number of items in test

#item level
df = df_items %>% filter(q %nin% c(1,2,3,4,5,6,9)) %>% mutate(
  accuracy = recode_factor(score_niceABS, "0" ="incorrect","1"="correct"),
  q = as.factor(q)
)

#FACETED HISTOGRAM
stats = df_subjects %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(item_test_NABS))
gf_props(~item_test_NABS, 
         fill = ~pretty_condition, data = df_subjects) %>% 
  gf_facet_grid(pretty_condition ~ pretty_mode) %>% 
  gf_vline(data = stats, xintercept = ~mean, color = "red") +
  labs(x = "# Correct",
       y = "proportion of subjects",
       title = "Test Phase Absolute Score (# Correct)",
       subtitle = "") + theme(legend.position = "blank")

#GROUPED PROPORTIONAL BAR CHART
gf_props(~accuracy, fill = ~pretty_condition, x =~pretty_condition,
       position = position_dodge(), data = df) %>%
  gf_facet_grid(~pretty_mode) +
   labs(x = "Correct Responses in Test Phase",
       title = "Accuracy on Task by Condition",
       subtitle="Impasse Condition yields a greater proportion of correct responses")

#STACKED PROPORTIONAL BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = accuracy)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~pretty_mode) + 
   labs(#y = "",
       title = "Accuracy on Test Phase",
       x = "Condition",
       fill = "",
       subtitle="Impasse Condition yields a greater proportion of correct responses")

```

## MIXED --- CONTINUOUS

## MIXED --- BINOMIAL

#### Mixed Logistic Regression

*Fit a mixed logistic regression (at the item level), predicting accuracy (absolute score) on test phase questions by condition; accounting for random effects of subject.*

##### Fit Model

```{r}

#SETUP DATA 
n_items = 8 #number of items in test

#item level
df = df_items %>% filter(q %nin% c(1,2,3,4,5,6,9)) %>% mutate(
  accuracy = as.factor(score_niceABS),
  q = as.factor(q)
) %>% filter(mode =="lab-synch")

library(lmerTest) #for CIs in glmer 

## 1 | SETUP RANDOM EFFECT

#:: EMPTY MODEL (baseline, no random effect)
m0 = glm(accuracy ~ 1, family = "binomial", data = df) 

#:: RANDOM INTERCEPT SUBJECT
mm.rS <- glmer(accuracy ~ (1|subject), data = df,family = "binomial")

#:: RANDOM INTERCEPT SUBJECT ITEM
mm.rS <- glmer(accuracy ~ (1|subject) + (1|q), data = df,family = "binomial")

# :: TEST random effect
paste("AIC with random effect is lower than glm empty model?", m0$aic > AIC(logLik(mm.rS)))
test_lrt(m0,mm.rS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(m0,mm.rS))$p[2])


## 2 | ADD FIXED EFFECT

# SUBJECT INTERCEPT | FIXED CONDITION 
mm.CrS <- glmer(accuracy ~ pretty_condition + (1|subject), 
                data = df,family = "binomial")

# SUBJECT INTERCEPT | FIXED CONDITION 
mm.CrSQ <- glmer(accuracy ~ pretty_condition + (1|subject) + (1|q), 
                data = df,family = "binomial")

# :: TEST fixed factor 
paste("AIC with fixed effect is lower than random intercept only model?", AIC(logLik(mm.rS)) > AIC(logLik(mm.CrS)) )
test_lrt(mm.rS,mm.CrS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rS,mm.CrS))$p[2])
```

##### Visualize

```{r}

#: PRINT MODEL 
print("PREDICTOR MODEL")
summ(mm.CrS)

#: INTERPRET COEFFICIENTS

print("MODEL PERFORMANCE")
performance(mm.CrS)
print("SANITY CHECK REPORTING")
report(mm.CrS)

# se <- sqrt(diag(stats::vcov(m1)))
# # table of estimates with 95% CI
# (tab <- cbind(Est = fixef(m1), LL = fixef(m1) - 1.96 * se, UL = fixef(m1) + 1.96 *
#     se))
# (e <- exp(tab))

#: PLOT

#SJPLOT | MODEL | ODDS RATIO
#library(sjPlot)
plot_model(mm.CrS, type="std2", vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE) +  
  labs(title = "Model Predicted Odds Ratio",
       subtitle = "",
       x = "Condition")

#SJPLOT | MODEL | PROBABILITIES
plot_model(mm.CrS, type="pred",
           show.intercept = TRUE,
           show.values = TRUE,
           title = "Model Predicted Probability of Accuracy",
           axis.title = c("Condition","Probability of Accurate Response"))

#GGEFFECTS | MODEL | PROBABILITIES
# library(ggeffects)
# ggeffect(model = mm.CrS) %>% plot()

#SANITY CHECK SJPLOT
# library(effects)
# plot(allEffects(mm.CrS))

```

##### Diagnostics

```{r}
check_model(mm.CrS)
binned_residuals(mm.CrS)
```

##### Inference

We fit a mixed-effect binomial logistic regression model with random intercepts for subjects to investigate the effect of condition on test phase item accuracy. The model including a fixed effect of condition performed significantly better than an intercept-only baseline model (χ2(3): 4.98, p \< 0.05). Consistent with the pattern of results for the first question only, across all test-phase items, being in the impasse condition increases the odds of a correct response by a factor of 5 over the control condition $e^{\beta_1}$ = 5.11, 95% CI \[1.17,22,36\], p \< 0.05.

```{r}
# PRETTY TABLE SJPLOT
tab_model(mm.CrS)
```


#### BAYESIAN Mixed Lostistic Regression

**Sanity Check** Compare frequentist model with bayesian alternative

```{r}
#| label: MODEL-BAYES-ACC

df_i = df_items %>% filter(q %nin% c(6,9)) 


##:::::::: FLAT PRIORS
#fit matching bayesian model [flat priors]
FLAT_B.mm.CrSQ <-  brm(
  bf(accuracy ~ pretty_condition + (1|subject) + (1|q)),
  data = df_i,
  family = bernoulli(link = "logit"),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file = "analysis/utils/models/FLAT_B.mm.CrSQ.rds"
)

#describe model
summary(FLAT_B.mm.CrSQ)

#compare to frequentist model
paste("COMPARE TO FREQUENTIST")
compare_models(mm.CrSQ, FLAT_B.mm.CrSQ)


#_Here we see that the model with default priors yields estimates comparable to the frequentist mixed model, though slightly smaller for fixed effects condition, with tighter credible intervals._


##:::::::: INFORMATIVE PRIORS

#describe priors that were used
paste("DEFAULT PRIORS")
prior_summary(FLAT_B.mm.CrSQ)

#set informative priors for fixed effects
#parameters normally distributed around the mean of the expected log coefficient
#we expect low probability of accuracy in control [intercept] and in impasse, but higher variance in impasse
priors <- c(
  prior(normal(-1, 2), class = b),
  prior(normal(-1, 4), class = b, coef="pretty_conditionimpasse")
)

#fit model with informative priors
B.mm.CrSQ <-  brm(
  accuracy ~ pretty_condition + (1|subject) + (1|q),
  prior = priors,
  data = df_i,
  family = bernoulli(link = "logit"),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234,
  backend = "cmdstanr",
  file = "analysis/utils/models/B.mm.CrSQ.rds"
)

#describe model
summary(B.mm.CrSQ)

#compare to frequentist model
paste("COMPARE TO UNFINFORMATIVE PRIOR")
compare_models(FLAT_B.mm.CrSQ, B.mm.CrSQ)

```
_Here we see tha the model with the informative prior makes predictions similar to the uninformative prior model_. 

```{r}
#| label: MODEL-DESC-ACC


#DIAGNOSTICS

#set model 
m <- B.mm.CrSQ

summary(m)

#POSTERIOR CHECKS
library(bayestestR)
diagnostic_posterior(m)
pp_check(m, ndraws = 500)

plot(m)


#INTERPRET
describe_posterior(m, test = c("p_direction", "rope", "bayesfactor"))


#set working model
m <- B.mm.CrSQ


#equivalence test 
result <- equivalence_test(m, rule = "cet")
plot(result)

#rope 
result <- rope(m)
plot(result)

#direction of effect 
result <- pd(m)
plot(result)


# all parameters
m %>% 
  posterior_samples() %>% 
  select(-c(lp__, contains("["))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(x = value)) +
  stat_halfeye(point_interval = mode_hdi,
               fill = "lightblue") + 
  facet_wrap(~ variable,
             ncol = 2,
             scales = "free") +
  theme(text = element_text(size = 12))

#plot parameters
stanplot(m, 
         type = "areas",
         prob = 0.95) 


#plot model
plot_model(m)


#PLOT PARAMETERS
result <- model_parameters(m, exponentiate = TRUE, 
                           component = "all")
plot(result) + 
  labs(
  title = "Accuracy ~ Condition + (1 | subject) + (1 | question)",
  subtitle = "Logistic Mixed Effects Model"
)  + theme_clean() +  theme(legend.position = "blank") 

#PLOT MODEL PREDICTION
plot_model(m, type = "pred")[[1]] + 
  ylim(0,1) + labs(
    title = "Model Prediction | Probability of Accurate Response",
    subtitle = "Impasse increases Probability of Correct Response"
  )
```
```{r}
#report(m)
```




## MIXED --- ORDINAL

## MIXED --- CATEGORICAL 
### Mixed Multinomial Regression


*Does condition affect the response state of of items across the task?*

*Fit a MIXED logistic regression predicting interpretation state (k=3) by condition(k = 2).*

-   Can use mclogit mblogit() with random effect *or* bayesian brms package b/c nlme, lme4 don't support random effects on multinomial (ie no categorical family on glmer())

-   Alternative would be to manually run [k-1] X binomial mixed models \[should compare outcomes\]

-   [k-1] equations will be estimated (# categories - 1); each representing the odds of answering in that particular interpretation (vs) the reference category (orthogonal answer) \[essentially a series of binary logistic regressions, but instead of incorrect/correct, comparing \[reference category\] vs \[this category\])

-   For *each* equation:

    -   $\beta_{0}$ *= Log Odds of \[this category type vs. reference category type) response in CONTROL condition*
    -   $e^{\beta_{0}}$ *= ODDS of \[this category type vs. reference category type\] response in CONTROL condition*
    -   $\beta_{1}$ *=* $\beta_{1impasse}$ *Log Odds (Log OR; change in odds for \[this category\] type response in impasse (vs) control \[log scale\])*
    -   $e^{\beta_{1}}$ *= ODDS RATIO of \[this. vs reference category type\] response in IMPASSE (vs) CONTROL*
    -   Two-tailed NHST *Null hypothesis:* $\beta_{impasse} = 0$ *the odds for \[this category of response vs. reference\] are not different for IMPASSE condition*
    -   *Alternative hypothesis:* $\beta_{impasse} \ne 0$ *the odds of \[this category of response vs. reference\] increases or decreases for IMPASSE condition*

##### Fit Model \[MANUAL INDIVIDUAL BINOMIALS\]

```{r}

#VERIFY RESULTS BELOW VIA MULTIPLE INDIVIDUAL MODELS

paste("ORTH vs. UNKNOWN")

d1 <- df_i %>% filter(state %nin% c("tri-like")) %>% droplevels()
table(d1$pretty_condition, d1$state)
plot(d1$pretty_condition, d1$state)

m.unknown <- glmer(state ~ pretty_condition + ( 1 | subject) + (1 | q), family = "binomial", data = d1)
summary(m.unknown)
```

*Being in the IMPASSE condition increases* the odds of giving an unknown (potentially nonsense) or blank/uncertain response rather than an orthogonal (or satisficing) response by a factor of 19 ( logodds = 2.92, z = 6.89, p \< 0.001) . Participants in the impasse condition were 19x as likely to give an unknown/uncertain response rather than an orthogonal response compared to participants in control. (Conditional R2 = 0.67, marginal R2 = 0.21; SD(subject) = 1.55, SD(question) = 2.22)

```{r}

paste("ORTH vs. TRIANGULAR")

d2 <- df_i %>% filter(state %nin% c("unknown")) %>% droplevels()
table(d2$pretty_condition, d2$state)
plot(d2$pretty_condition, d2$state)

m.tri <- glmer(state ~ pretty_condition + ( 1 | subject) + (1 | q), family = "binomial", data = d2)
summary(m.tri)

```

*Being in the IMPASSE condition increases* the odds of giving an triangle-like response by a factor of 305 ( logodds = 5.72, z = 4.14, p \< 0.001) . Participants in the impasse condition were 305 times as likely to give an triangle-like response rather than an orthogonal response compared to participants in control. (Conditional R2 = 0.91, marginal R2 = 0.21; SD(subject) = 5.05, SD(q) = 1.09)

##### Fit Model \[mblogit\]

```{r}

#https://www.elff.eu/software/mclogit/manual/mblogit/
#"baseline category logit" model matches multinom()

#check reference level 
print("Categories (first is reference)")
levels(df_i$state)

#FIT EMPTY MODEL
# print("EMPTY MODEL")
m.mbl0 <- mblogit(state ~ pretty_condition ,  #no random intercepts; fixed only model 
                  data = df_i)
#summary(m.mbl0)

#FIT PREDICTOR MODEL
# print("PREDICTOR MODEL")
m.mbl1 <- mblogit(state ~ pretty_condition , 
                  random = list( ~ 1|subject, ~1|q), 
                  data = df_i)
# summary(m.mbl1)

#COMPARE MODEL FIT
paste("AIC wth predictor is lower than empty model?", AIC(m.mbl0) > AIC(m.mbl1))
test_lrt(m.mbl0, m.mbl1)

#DESCRIBE MODEL
summary(m.mbl1)

#INTERPRET COEFFICIENTS
cint <- confint(m.mbl1, level = 0.95)
print("ODDS RATIO")
(e <- cbind( exp(coef(m.mbl1)), exp(cint))) #exponentiated, adjusted

#PERFORMANCE
performance(m.mbl1)

#TABLE
tab_model(m.mbl1, transform = "exp", title = "Model Predicted Odds Ratio")

```


##### Inference

-   Being in the IMPASSE condition *increases* the odds of giving an unknown (potentially nonsense) or blank/uncertain response rather than an orthogonal (or satisficing) response by a factor of 12 (z = 6.48, p \< 0.001) . **Participants in the impasse condition were 12x as likely to give an unknown/uncertain response rather than an orthogonal response compared to participants in control**.

-   Being in the IMPASSE condition *increases* the odds of giving 'triangle-like' response rather than an orthogonal (or satisficing) response by a factor of 29 (z = 4.63, p \< 0.001 ). **Participants in the impasse condition were more than 29x as likely to give an a triangular response rather than an orthogonal response compared to participants in control**.

-   As with the (binary) logistic regression on accuracy \~ condition, significant model intercepts indicate that the odds of being in any particular response state (vs) orthogonal are significantly different than 1 *in the control condition.* (i.e. not 1:1 odds or 50% chance of being in that response state. Orthogonal is a much more probable response state in control)

##### Visualize

```{r}


#set model 
m <- m.mbl1

#:::::::: PLOT

#SJPLOT | MODEL | ODDS RATIO
#library(sjPlot)
plot_model(m, vline.color = "red", 
           transform = "exp", #for some reason have to manually add for mixed?
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.05, #can manually adjust to account for directional test
           ci.lvl = 0.95) +  #can manually adjusted for directional test   
  labs(title = "ODDS RATIO | State ~ condition",
       subtitle = "(p for  two-tailed test)")

library(parameters)

#PLOT PARAMETER ESTIMATES
result <- simulate_parameters(m)
plot(result, stack = FALSE)

result <- equivalence_test(m, rule = "cet")
plot(result)

#raw coefficients
plot_model(m)


#PLOT PARAMETERS
result <- model_parameters(m, exponentiate = TRUE, 
                           component = "all")
plot(result) + 
  labs(
  title = "Accuracy ~ Condition + (1 | subject) + (1 | question)",
  subtitle = "Logistic Mixed Effects Model"
)  + theme_clean() +  theme(legend.position = "blank") 

s
#TODO CAN'T SEEM TO COMPUTE MARGINAL EFFECTS PLOT FOR THE MIXED MODEL
# pr <- ggpredict(m, "pretty_condition", type = "fixed")
# plot(pr)
#PLOT MODEL PREDICTION
# plot_model(m, type = "pred")
# plot_model(m, type = "eff")

# TERNARY PLOT
# library(plot3logit)
# field3logit(ggpredict(m))
#https://cran.r-project.org/web/packages/plot3logit/vignettes/plot3logit-overview.html


#SJPLOT | MODEL | TABLE
tab_model(m)

```

##### Diagnostics

**COMPARE TO MANUAL MODELS**
```{r}
compare_models(m, m.unknown, m.tri)
```

##### Fit Model \[brms\]

```{r}

#BAYESIAN MIXED VERSION
mixcat.1 <- brm( state ~ pretty_condition + (1|subject) + (1|q), 
                 data = df_i, 
                 family = "categorical",
                 chains = 4, iter = 2000, warmup = 1000,
                 cores = 4, seed = 1234,
                 backend = "cmdstanr",
                 file ="analysis/utils/models/mixcat1.rds")
```


```{r}

m <- mixcat.1

#DESCRIBE MODEL
summary(m)
# report(mixcat.1)

#TABLE
# tab_model(m,
          # show.r2 = FALSE) #, transform = "exp", title = "Model Predicted Odds Ratio")


#TODO OUTPUT TABLE 
#https://arelbundock.com/posts/modelsummary_multinomial_logit/
# modelsummary(m)

```

##### Inference

-   Being in the IMPASSE condition *increases* the odds of giving an unknown (potentially nonsense) or blank/uncertain response rather than an orthogonal (or satisficing) response by a factor of 43.86.

-   Being in the IMPASSE condition *increases* the odds of giving 'triangle-like' response rather than an orthogonal (or satisficing) response by a factor of 6.64.


##### Visualize

```{r}
# 
# #:::::::: PLOT
# 
# #SJPLOT | MODEL | ODDS RATIO
# #library(sjPlot)
# plot_model(m, vline.color = "red", 
#            show.intercept = TRUE, 
#            show.values = TRUE,
#            p.threshold = 0.05, #can manually adjust to account for directional test
#            ci.lvl = 0.95 ) +  #can manually adjusted for directional test   
#   labs(title = "ODDS RATIO | State ~ condition",
#        subtitle = "(p for one two test)")
# 
# 
# result <- estimate_density(m)
# plot(result, stack = FALSE, priors = TRUE)
# 
# result <- describe_posterior(m)
# plot(result, stack = FALSE, priors = TRUE)
# 
# result <- p_direction(m)
# plot(result, stack = FALSE)
# 
# #ROPE
# result <- rope(m, ci = 0.89)
# plot(result) + theme_clean()
# 
# result <- equivalence_test(m)
# plot(result)
# 
# #ERROR incompatible arguments to calculate multivariate normal distribution
# # result <- simulate_parameters(m)
# # plot(result)
# 
# #check posterior
# pp_check(m, ndraws=1000)
# 
# #PERFORMANCE
# # performance(m)
# 
# #Plot parameters
# plot(model_parameters(m, exponentiate = TRUE))
# 
# #plot marginal predictions of model
# plot_model(m, type = "pred")[[1]] + 
#   labs(
#     title = "Model Predicted Marginal Probabilities",
#     x = "Condition", 
#     y = "probability of being in each state"
#   ) + theme_clean() 
```

##### COMPARE

```{r}
# compare_models(m.unknown, m.tri, m.mbl1, mixcat.1)
```
The predictions of the manual, frequentist mixed multinomial and bayesian mixed multinomial models are comparable. 

DECISION: Report bayesian model, because for some reason the visualization of the marginal predictions is actually working?

##### TODO

-   priors? used default flat priors... ok?
-   posterior predictive checks
-   diagnostics on random effects
-   reconcilliation of mblogit() vs brms versions of the model; seems like they should yield similar estimates




```{r}
#BAYESIAN MIXED VERSION
#df <- df_items %>% filter(q %nin% c(1,2,3,4,5,6,9)) %>% filter(term=="spring18")
# mm1 <- brm( state ~ condition + (1|subject), data = df, family = "categorical", 
# file = "analysis/models/sgc3a_brms_mixedcat_teststate.rds" # cache model (can be removed)))
# summary(mm1)
# performance(mm1)
# plot(mm1)
# #report(mm1)
# #check_posterior_predictions(mm1, draws=100)
# # library(bayesplot)
# library(bayestestR)
# plot(rope(mm1))
```

## WIP UNKNOWN

### Cummulative Ordinal (Bayesian) --- Equal Variance

https://journals.sagepub.com/doi/full/10.1177/2515245918823199

```{r}
library(brms)


#DEFINE DATA 
df <- df_items %>% filter(q==1) %>% filter(mode =="lab-synch")
  
#TODO: why is this probit instead of logit?
cumord <- brm( formula = state ~ condition,
               data = df,
               family = cumulative("probit"),
               file = "analysis/utils/models/sgc3a_brms_cumord_q1state.rds" # cache model (can be removed)
)

summary(cumord)
conditional_effects(cumord, "condition", categorical = TRUE)

#SJPLOT
plot_model(cumord)
plot(cumord)
plot(rope(cumord))

#REPORT
report(cumord)

# ord_cum %>%
#   spread_draws(b_Intercept, r_condition[condition,]) %>%
#   mutate(condition_mean = b_Intercept + r_condition) %>%
#   ggplot(aes(y = condition, x = condition_mean)) +
#   stat_halfeye()

# performance(cumord)
# plot(cumord)
```

### Cummulative Ordinal (Bayesian) --- Unequal Variance

https://journals.sagepub.com/doi/full/10.1177/2515245918823199

```{r}
library(brms)


#DEFINE DATA 
df <- df_items %>% filter(q==1) %>% filter(mode =="lab-synch")
  
#TODO: why is this probit instead of logit?
u.cumord <- brm( 
  formula = bf(state ~ condition) +
               lf(disc ~ 0 + condition, cmc = FALSE),
               data = df,
               family = cumulative("probit"),
               file = "analysis/utils/models/sgc3a_brms_ucumord_q1state.rds" # cache model (can be removed)
)

summary(u.cumord)
conditional_effects(u.cumord, "condition", categorical = TRUE)

#SJPLOT
plot_model(u.cumord)
plot(u.cumord)
plot(rope(u.cumord))

#REPORT
report(u.cumord)

# ord_cum %>%
#   spread_draws(b_Intercept, r_condition[condition,]) %>%
#   mutate(condition_mean = b_Intercept + r_condition) %>%
#   ggplot(aes(y = condition, x = condition_mean)) +
#   stat_halfeye()

# performance(cumord)
# plot(cumord)
```

In brms, the parameter related to latent variances is called disc (short for "discrimination"), following conventions in item response theory. Note that disc is not the variance itself, but the inverse of the standard deviation, s. That is, s = 1/disc. Further, because disc must be strictly positive, it is by default modeled on the log scale.

### Adjacent-Category Ordinal (Bayesian)

```{r}
df <- df_items %>% filter(q==1) %>% filter(mode =="lab-synch")


# # To specify an adjacent-category model, we use family = acat() instead of family = cumulative() as an argument to the brm() function. Then, to model condition with possible category-specific effects, we wrap this variable in cs() in the model’s formula:

adjcat <- brm( formula = state ~ cs(condition),
               data = df,
               family = acat("probit"),
               file = "analysis/utils/models/sgc3a_brms_adjcat_q1state.rds" # cache model (can be removed)
)
 
summary(adjcat)

conditional_effects(cumord, "condition", categorical = TRUE)
conditional_effects(adjcat, "condition", categorical = TRUE)

#WHICH IS BETTER? cumulative or adjacent?
plot(compare_performance(cumord, adjcat))
compare_performance(cumord, adjcat)

# #TIDYBAYES VISUALIZATION
# library(tidybayes)
# adjcat %>%
#   spread_draws(b_Intercept, r_condition[condition,]) %>%
#   mutate(condition_mean = b_Intercept + r_condition) %>%
#   ggplot(aes(y = condition, x = condition_mean)) +
#   stat_halfeye()

plot(cumord)
plot(adjcat)
report(adjcat)

```

*Converges, but estimates are unreliable?*
