---
title: "Fall 2017 Data Cleaning"
author: "Amy Rae Fox"
date: "11/2/2021"
output:
  html_document: default
  pdf_document: default
---
*The purpose of this file is processing the combined data files for Fall 2017 into study-level files that contain only valid data for analysis, excluding invalid sessions and conditions.*

```{r SETUP, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(dplyr)
library(tidyjson)

#set current folder as working directory
# setwd("~/Sites/RESEARCH/SGC—Scaffolding Graph Comprehension/SGC-X/DATA/3_ready_files/fall_2017")
```

Data is imported from 2 files, indicating two levels of analysis: participants and blocks (item-level). **Note: mouse-cursor data contained in final_mouse_blocks.json file is not handled here.**

```{r IMPORT}
#IMPORT DATA
df_participants <- fromJSON("combined_files/final_participants.json")
df_blocks <- fromJSON('combined_files/final_blocks.json')

#add term indicator
df_participants$term <- "fall17"
df_blocks$term <- "fall17"

```

```{r PARTICIPANTS-SETUP}

#create factors in PARTICIPANTS
df_participants <- df_participants %>%  
  select(subject,session,term,condition,  #re-arrange columns
         ts_n, tt_n,triangular_score,
         os_n, ot_n,orthogonal_score,
         explicit,impasse,axis,
         triangular_time, totalTime, ts_t, tt_t,
         attn_check,
         native_language, year, major, country, sex, age
         ) %>% #reorder columns 
  mutate( #create factors and remove extraneous ""
    subject=factor(subject),
    condition=factor(condition),
    session=factor(session),
    term=factor(term),
    explicit=factor(explicit),
    axis=factor(axis),
    impasse=factor(impasse),
    sex = as.factor(gsub('"',"",sex)),
    age = as.double(gsub('"',"",age)),
    country = gsub('"',"",country),
    major = gsub('"',"",major),
    year = gsub('"',"",year),
    native_language = gsub('"',"",native_language),
) 

```

```{r BLOCKS-SETUP}
df_blocks <- df_blocks %>% 
  select( #reorder columns
    subject, session, term, condition,
    q,question,answer,rt,
    correct, orth_correct,
    explicit, impasse, axis) %>% 
  mutate(
    subject=factor(subject),
    condition=factor(condition),
    session=factor(session),
    term=factor(term),
    explicit=factor(explicit),
    axis=factor(axis),
    impasse=factor(impasse),
    q=factor(q),
    question=factor(question)
  )
```

### Sessions
The (string) session code is entered by the participant based on instructions given by the experimenter, and documents the data-collection session (eg. in-person at a particular time). This code is also used by the experimenter to differentiate test or expert data collection runs.  
```{r SESSIONS}

#MANUALLY INSPECT sessions
df_participants %>% group_by(session) %>% 
  summarize(n=n())

#manually recode sessions in participants
df_participants$session <- recode(df_participants$session,
                                  'alpha'='alfa',
                                  '111'="XTRA")

#manually recode sessions in blocks
df_blocks$session <- recode(df_blocks$session,
                                  'alpha'='alfa',
                                  '111'="XTRA")

df_participants %>% group_by(session) %>% 
  arrange(desc(session)) %>% 
  summarize(n=n())
```
In Fall 2017, 14 data collection sessions were used, from ALFA -> PINECONE. Participants who mispelled their sessions have been manually recoded, and one participant erroneously entered their condition code as their session code, and this entry is corrected to 'XTRA'. 

**No data need to be excluded based on SESSION CODE.**

### Conditions
The three digit condition code is entered by the participant based on instructions given by the experimenter, and determines the stimulus that the participant experiences during the study. 

```{r CONDITIONS}
df_participants %>% group_by(condition) %>% 
  summarize(n=n())

#SET CONDITION FACTORS FOR EACH STUDY
#SGC3A is the simple insight study, control (111) vs impasse (121)
f_sgc3a <- c(111,121) 

#SGC3B is the factorial insight study (111 control, 121 insight, 211 static, 221 static-impasse, 311 ixn 321 ixn-impasse)

f_sgc3b <- c(111,121,211,221,311,321)

#SGC4 is the gridlines study 111, 112, 113
f_sgc4 <- c(111,112,113)
```


In Fall 2017, data were gathered for three study designs: SGC3A (simple insight vs. control), SGC3B (partial data collected: full factorial insight vs. explicit) and SGC4(partial data collected:gridlines). 

Finally, data from the master participants and blocks files are segregated into separate files for each individual study, separated by condition.

```{r GENERATE FILES}

#SEPARATE PARTICIPANTS FILES
df_sgc3a <- df_participants %>% filter (condition %in% f_sgc3a)
df_sgc3a %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc3a,"study_files/fall17_sgc3a_participants.csv", row.names = FALSE)

df_sgc3b <- df_participants %>% filter (condition %in% f_sgc3b)
df_sgc3b %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc3b,"study_files/fall17_sgc3b_participants.csv", row.names = FALSE)

df_sgc4 <- df_participants %>% filter (condition %in% f_sgc4)
df_sgc4 %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc4,"study_files/fall17_sgc4_participants.csv", row.names = FALSE)

#SEPARATE BLOCKS FILES
df_sgc3a <- df_blocks %>% filter (condition %in% f_sgc3a)
df_sgc3a %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc3a,"study_files/fall17_sgc3a_blocks.csv", row.names = FALSE)

df_sgc3b <- df_blocks %>% filter (condition %in% f_sgc3b)
df_sgc3b %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc3b,"study_files/fall17_sgc3b_blocks.csv", row.names = FALSE)

df_sgc4 <- df_blocks %>% filter (condition %in% f_sgc4)
df_sgc4 %>% group_by(condition) %>% 
  summarize(n=n())
write.csv(df_sgc4,"study_files/fall17_sgc4_blocks.csv", row.names = FALSE)

```