---
# title: 'Exploration'
subtitle: 'Study SGC3A | 3 Exploration'
# author: 'Amy Rae Fox'
# always_allow_html: true  
# header-includes:
#    - \usepackage{amsmath}
# output:
#   html_document:
#     theme: yeti
#     code_folding: hide
#     fig_caption: yes
#     number_sections: yes
#     toc: yes
#     toc_depth: 4
#     toc_float:
#       collapsed: no
#       smooth_scroll: yes
#   pdf_document: 
#     toc: true
#     toc_depth: 3
#     latex_engine: xelatex
# font-family: "DejaVu Sans"
# mainfont: "DejaVu Sans"
---

\newpage

# Exploration {#sec-SGC3A-exploration}

**THIS NOTEBOOK IS INCOMPLETE**

TODO

-   double check which Qs in task are non-discriminating (6 and 9?)

```{r}
#| label: SETUP
#| echo : true
#| warning : false
#| message : false

library(tidyverse) #ALL THE THINGS
library(kableExtra) #printing tables 
library(mosaic) #simple descriptives [favstats]
library(Hmisc) # %nin% operator
library(ggpubr) #arrange plots

# library(pastecs) #stat.desc
# library(car) #ANOVA, qqplot
# library(effectsize) #effect size

#set some output options
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

```

*The purpose of this notebook is explore the distribution of dependent variables for Study SGC3A.*

+------------------------+--------------+
| Pre-Requisite          | Followed By  |
+========================+==============+
| 1_sgc3A_harmonize.qmd\ |              |
| 2_sgc3A_rescoring.qmd  |              |
+------------------------+--------------+

```{r}
#| label: IMPORT-DATA
#| echo : true
#| warning : false
#| message : false

#IMPORT DATA 
#TODO REPLACE WITH 2-scored-data when its done
df_items <- read_rds('data/2-scored-data/sgc3a_scored_items.rds')
df_subjects <- read_rds('data/1-study-level/sgc3a_participants.rds')

#SEPARATE ITEM DATA BY QUESTION TYPE
df_scaffold <- df_items %>% filter(q < 6)
df_test <- df_items %>% filter(q > 6) %>% filter (q %nin% c(6,9))
```

## Sample

### Data Collection

Data was initially collected (in person, SONA groups in computer lab) in Fall 2017. In Spring 2018, additional data were collected after small modifications were made to the experimental platform to increase the size of multiple-choice input buttons, and to add an additional free-response question following the main task block. In Fall 2021, the study was replicated using asynchronous, online SONA pool, with additional participants collected in Winter 2022.

```{r}
#| label : INSPECT-DATA-COLLECTION

title = "Participants by Condition and Data Collection Period"
cols = c("Control Condition","Impasse Condition","Total for Period")
cont <- table(df_subjects$term, df_subjects$condition)
cont %>% addmargins() %>% kbl(caption = title, col.names = cols) %>%  kable_classic()

```

### Participants

```{r PARTICIPANTS}
#Describe participants
subject.stats <- rbind(
  "lab"= df_subjects %>% filter(mode == 'lab-synch') %>% select(age) %>% unlist() %>% favstats(),
  "online" = df_subjects %>% filter(mode == "asynch") %>% select(age) %>% unlist() %>% favstats()
) 
subject.stats$female <- c(
  (df_subjects %>% filter(mode == 'lab-synch') %>%  filter(gender=="Female") %>% count())$n,
  (df_subjects %>% filter(mode == "asynch") %>% filter(gender=="Female") %>% count())$n
)

title = "Descriptive Statistics of Participant Age and Gender"
subject.stats %>% kbl (caption = title) %>% kable_classic()
```

For **in-person** collection, `r subject.stats["lab",]$n` participants (`r round(subject.stats["lab",]$female/subject.stats["lab",]$n,1) * 100` % female ) undergraduate STEM majors at a public American University participated *in person* in exchange for course credit (age: `r (subject.stats['lab','min'])` - `r (subject.stats['lab','max'])` years). Participants were randomly assigned to one of two experimental groups.

For **online replication** `r subject.stats["online",]$n` participants (`r round(subject.stats["online",]$female/subject.stats["online",]$n,1) * 100` % female ) undergraduate STEM majors at a public American University participated *online, asynchronously* in exchange for course credit (age: `r (subject.stats['online','min'])` - `r (subject.stats['online','max'])` years). Participants were randomly assigned to one of two experimental groups.

## Response Accuracy

### Cumulative Scores

Cumulative scores indicate the response accuracy by participant across all items in the graph comprehension task.

#### Cumulative Absolute Score

Recall from [Section -@sec-absolute-scoring] that the absolute score (following the dichotomous scoring approach) indicates if the subject's response for a particular item was *perfectly* correct: whether they selected all correct answer options and no others (excluding certain allowed exceptions, such as also selecting the data point referenced in the question). The absolute score for an individual item is either 0 or 1. When summarized across the entire set of items, the cumulative absolute score for an individual subject ranges from \[0,15\].

```{r}
#| label: ABS-SUBJECT

title = "Descriptive Statistics of Response Accuracy (Absolute Score)"
abs.stats <- rbind(
  "lab"= df_subjects %>% filter(mode == 'lab-synch') %>% select(absolute_score) %>% unlist() %>% favstats(),
  "online" = df_subjects %>% filter(mode == "asynch") %>% select(absolute_score) %>% unlist() %>% favstats()
) 
abs.stats %>% kbl (caption = title) %>% kable_classic()

```

For *in person* collection, cumulative absolute scores (n = `r abs.stats["lab",]$n`) range from `r round(abs.stats["lab",]$min,2)` to `r round(abs.stats["lab",]$max,2)` with a mean score of (M = `r round(abs.stats["lab",]$mean,2)`, SD = `r round(abs.stats["lab",]$sd,2)`).

For *online replication*, (online) cumulative accuracy scores (n = `r abs.stats["online",]$n`) range from `r round(abs.stats["online",]$min,2)` to `r round(abs.stats["online",]$max,2)` with a mean score of (M = `r round(abs.stats["online",]$mean,2)`, SD = `r round(abs.stats["online",]$sd,2)`).

```{r}
#| label: VIS-SUBJECT-ABS-SCORE
 
#VISUALIZE distribution of response accuracy across SUBJECTS

# plab <- ggplot(data = df_subjects %>% filter(mode == 'lab-synch'), aes(x = absolute_score)) + 
#   geom_bar() +  # = 16, color = "black", data =  %>%
#   facet_wrap(condition) + 
#   geom_vline (aes(xintercept=mean(absolute_score)),color="blue", linetype="dashed", size=0.5) +
#   labs(title="Lab", x = "Cumulative Absolute Score", y="% of subjects") + theme_minimal()
# 
# ponline <- ggplot(data = df_subjects %>% filter(mode == 'asynch'), aes(x = absolute_score)) + 
#   geom_bar() +  # = 16, color = "black", data =  %>%
#   facet_wrap(~condition) + 
#   geom_vline (aes(xintercept=mean(absolute_score)),color="blue", linetype="dashed", size=0.5) +
#   labs(title="Online", x = "Cumulative Absolute Score", y="% of subjects") + theme_minimal() 
# 
# plot <-ggarrange(plab, ponline, common.legend = TRUE, nrow = 1, ncol =2)
# annotate_figure(plot, top = text_grob("Cumulative Absolute Accuracy Score by MODALITY",
#    color = "black", face = "bold", size = 14))
# 
# 
# ggplot(data = df_subjects, aes(x = absolute_score)) + 
#   geom_bar() +  # = 16, color = "black", data =  %>%
#   facet_grid(condition~mode) + 
#   # geom_vline (aes(xintercept=mean(absolute_score)),color="blue", linetype="dashed", size=0.5) +
#   labs(title="Online", x = "Cumulative Absolute Score", y="% of subjects") + theme_minimal() 


stats = df_subjects %>% group_by(condition, mode) %>% summarise(mean = mean(absolute_score))
gf_bar(~absolute_score, data = df_subjects) %>% 
  gf_facet_grid(condition~mode) %>% 
  gf_vline(data = stats, xintercept = ~mean, color = "red") +
  labs(x = "Cumulative Absolute Score for Subject")+
  theme_minimal()
  

#RIDGEPLOT
# ggplot(data = df_subjects, aes(x = absolute_score, y = mode)) +
#   geom_density_ridges() + xlim(0,15)+ 
#   facet_wrap(~condition, labeller = label_both) + 
#   labs( x = "Cumulative Absolute Score for Subject") +
#   theme_minimal()

```

**TODO** double check that the subject-totals match the sum of the item level totals

#### Cumulative Interpretation Scores















### Item Score

Item scores indicate the response accuracy by a participant on each individual question in the graph comprehension task.

#### Item Absolute Score



```{r}
#| label: ABS-ITEM

title = "Descriptive Statistics of Response Accuracy (Absolute Score)"
abs.stats.items <- rbind(
  "lab"= df_items %>% filter(mode == 'lab-synch') %>% select(score_niceABS) %>% unlist() %>% favstats(),
  "online" = df_items %>% filter(mode == "asynch") %>% select(score_niceABS) %>% unlist() %>% favstats()
) 
abs.stats.items %>% kbl (caption = title) %>% kable_classic()

```


For *in person* collection, item absolute scores (n = `r abs.stats.items["lab",]$n`) range from `r round(abs.stats.items["lab",]$min,2)` to `r round(abs.stats.items["lab",]$max,2)` with a mean score of (M = `r round(abs.stats.items["lab",]$mean,2)`, SD = `r round(abs.stats.items["lab",]$sd,2)`).

For *online replication*, (online) item accuracy scores (n = `r abs.stats.items["online",]$n`) range from `r round(abs.stats.items["online",]$min,2)` to `r round(abs.stats.items["online",]$max,2)` with a mean score of (M = `r round(abs.stats.items["online",]$mean,2)`, SD = `r round(abs.stats.items["online",]$sd,2)`).

```{r}
#| label: VIS-ITEMS-ABS-SCORE

#VISUALIZE distribution of response accuracy across ITEMS

ggplot(data = df_items, aes(x = score_niceABS, y = mode)) +
  geom_density_ridges()+ 
  facet_wrap(~condition, labeller = label_both) + 
  labs( x = "Absolute Score for Item") +
  theme_minimal()


```







#### Item Interpretation Scores

## Response Latency

-   [TODO: Investigate super high and super low response times.]{style="color: red;"}.

-   [TODO: Investigate appropriate models for response time data. (see: https://lindeloev.github.io/shiny-rt/)]{style="color: red;"}.

-   Especially see https://lindeloev.github.io/shiny-rt/ for ideas on modelling reaction time data

### Time on Study

```{r DESCRIBE-TOTALTIME}
#DESCRIBE distribution of response time
time.stats <- rbind(
  "lab"= df_subjects %>% filter(mode == 'lab-synch') %>% select(totaltime_m) %>% unlist() %>% favstats(),
  "online"= df_subjects %>% filter(mode == 'asynch') %>% select(totaltime_m) %>% unlist() %>% favstats()
)

title = "Descriptive Statistics of Response Latency (Time on Study)"
time.stats %>% kbl(caption = title) %>% kable_classic()
```

Total time on study for *in person* subjects (n = `r time.stats["lab",]$n`) ranged from `r round(time.stats["lab",]$min,2)` to `r round(time.stats["lab",]$max,2)` minutes with a mean duration of (M = `r round(time.stats["lab",]$mean,2)`, SD = `r round(time.stats["lab",]$sd,2)`).

Total time on study for *online replication* subjects (n = `r time.stats["online",]$n`) ranged from `r round(time.stats["online",]$min,2)` to `r round(time.stats["online",]$max,2)` minutes with a mean duration of (M = `r round(time.stats["online",]$mean,2)`, SD = `r round(time.stats["online",]$sd,2)`).

```{r}
#| label : VIS-TOTALTIME
#| message : false
#| warning : false

#VISUALIZE distribution of response time
plab <- gf_dhistogram(~totaltime_m, data = df_subjects) %>%
  gf_vline(xintercept = ~time.stats["lab",]$mean, color = "black") %>% 
  gf_fitdistr(dist="gamma", color="red")+
  labs(title="Lab", x = "Total Time (mins)", y = "% subjects") +  theme_minimal()

ponline <- gf_dhistogram(~totaltime_m, data = df_subjects) %>%
  gf_vline(xintercept = ~time.stats["online",]$mean, color = "black") %>% 
  gf_fitdistr(dist = "gamma", color="red")+
  labs(title="Online", x = "Total Time (mins)", y = "% subjects") +  theme_minimal()

plot <-ggarrange(plab, ponline, common.legend = TRUE, nrow = 1, ncol =2)

annotate_figure(plot, 
                top = text_grob("Total Time by Study Mode",color = "black", face = "bold", size = 14),
                bottom = text_grob("fit by Gamma distribution", face = "italic", size = 10))
                
```

**TODO consider log transform of response latency data** *see* archive sgc3A_participants.Rmd

### Time on Question

**TODO time on question**

## Resources

-   https://rpkgs.datanovia.com/ggpubr/reference/index.html
