---
subtitle: 'Study SGC3A | (Lab) Hypothesis Testing'
# YAML FOR generating modelsummary tables
# uncomment to run those  cells only 
# \usepackage{booktabs}
# \usepackage{siunitx}
# \newcolumntype{d}{S[input-symbols = ()]}
---

\newpage

# SGC3A (Lab) Hypothesis Testing {#sec-SGC3A-lab-hypotesting}

*The purpose of this notebook is test the hypotheses that determined the design of the SGC3A study.*

```{r}
#| label: SETUP
#| warning : false
#| message : false

#UTILITIES
library(Hmisc) # %nin% operator
library(broom) #tidy model output
library(broom.mixed) #tidy mixed models
library(mosaic) #favstats
library(svglite) #saving plots as svg
library(distributional)

#VISUALIZATION
# library(ggpubr) #arrange plots
# library(cowplot) #arrange shift function plots
library(ggformula) #easy graphs
# # library(vcd) #mosaic plots
# # library(vcdExtra) #mosaic plots
library(kableExtra) #printing tables
library(sjPlot) #visualize model coefficients
library(ggdist) #uncertainty viz
library(gghalves) # plots. in half
library(ggbeeswarm) # violin plot stuffs
library(statsExpressions)
library(ggstatsplot) #plots with stats
library(modelsummary) #latex tables for models!

#MODELLING
# library(rstatix) #helpful testing functions incl wilcoxon, etc
library(report) #easystats reporting
library(see) #easystats visualization
library(performance) #easystats model diagnostics
library(parameters) #easystats model summary and vis
# library(qqplotr) #confint on qq plot
# library(gmodels) #contingency table and CHISQR
# library(equatiomatic) #extract model equation
# library(pscl) #zeroinfl / hurdle models 
library(lme4) #mixed effects models
library(lmerTest) #for CIs in glmer
library(merTools) #prediction intervalus for glmers 
# library(ggeffects) #visualization log regr models
#MULTINOMIAL 
library(nnet) #multinomial logistic regression [not mixed] #no p values
library(mclogit) #frequentist mixed multinomial logistic regression [mblogit] #gives p values
#BAYESIAN
library(cmdstanr) #executing stan
library(brms) #bayesian mixed multinomials [+ other bayesian reg models]
library(bayestestR) 
library(modelr) #helping with tidybayes
library(tidybayes)

library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

#GRAPH THEMEING
# theme_set(theme_minimal()) 

# Custom ggplot theme to make pretty plots
# Get the font at https://fonts.google.com/specimen/Barlow+Semi+Condensed
theme_clean <- function() {
  theme_minimal(base_family = "Barlow Semi Condensed") +
    theme(panel.grid.minor = element_blank(),
          plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
          axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
          strip.text = element_text(family = "BarlowSemiCondensed-Bold",
                                    size = rel(1), hjust = 0),
          strip.background = element_rect(fill = "grey80", color = NA))
}

set_theme(base = theme_clean())

```

**Research Questions**

In SGC3A we set out to answer the following question: Does posing a mental impasse improve performance on the interval graph comprehension task?

**Experimental Hypothesis**\
*Learners posed with scenario designed to evoke a mental impasse will be more likely to correct interpret the graph.*

-   H1A \| Learners in the IMPASSE condition will score higher overall than learners in CONTROL.
-   H1B \| Learners in the IMPASSE condition will be more likely to correctly answer the first question than learners in CONTROL.

**Null Hypothesis**\
*No significant differences in performance will exist between learners in the IMPASSE and CONTROL conditions.*

```{r}
#| label: IMPORT-DATA
#| warning : false
#| message : false

#IMPORT DATA 
df_subjects <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds') %>% 
  filter(mode == "lab-synch") %>% #only get lab-run participants
  mutate(
    task_percent = DV_percent_NABS
  ) %>% droplevels()

df_items <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds') %>% 
  filter(subject %in% df_subjects$subject) %>% #only get lab-run participants
  mutate (
    q = as.factor(q), 
    subject = as.factor(subject),
    accuracy = recode_factor(score_niceABS, "0" ="incorrect","1"="correct"),
    # CODES TVERSKY AS TRI-LIKE
    # state = recode_factor(score_SCALED, #for ordinal
    #                      "-1" = "orth-like",
    #                      "-0.5" = "unknown",
    #                      "0" = "unknown",
    #                      "0.5" = "tri-like",
    #                      "1" = "tri-like"),
    # CODES TVERSKY AS OTHER
    state = recode_factor(score_SCALED, #for ordinal
                         "-1" = "orthogonal",
                         "-0.5" = "other",
                         "0" = "other",
                         "0.5" = "angular",
                         "1" = "triangular"),
    state = as.ordered(state)) 

```

## SAMPLE

### Data Collection


```{r}
#| label : DESC-DATA-COLLECTION

title = "Participants by Condition and Data Collection Period"
cols = c("Control Condition","Impasse Condition","Total for Period")
cont <- table(df_subjects$term, df_subjects$condition)
cont %>% addmargins() %>% kbl(caption = title, col.names = cols) %>%  kable_classic()

```

### Participants

```{r}
#| label: DESC-PARTICIPANTS

#Describe participants
subject.stats <-df_subjects %>% dplyr::select(age) %>% unlist() %>% favstats()
subject.stats$percent.male <- ((df_subjects %>% filter(gender=="Male") %>% count())/count(df_subjects))$n
subject.stats$percent.female <- ((df_subjects %>% filter(gender=="Female") %>% count())/count(df_subjects))$n
subject.stats$percent.other <- ((df_subjects %>% filter(gender %nin% c("Female","Male")) %>% count())/count(df_subjects))$n


title = "Descriptive Statistics of Participant Age and Gender"
subject.stats %>% kbl (caption = title) %>% kable_classic()%>% 
  footnote(general = "Age in Years", 
           general_title = "Note: ",footnote_as_chunk = T) 
```

**Overall** `r subject.stats$n` participants (`r round((subject.stats$percent.male),2) * 100` % male, `r round((subject.stats$percent.female),2) * 100` % female, `r round((subject.stats$percent.other),2) * 100` % other)  undergraduate STEM majors at a public American University participated in exchange for course credit (age: `r (subject.stats$min)` - `r (subject.stats$max)` years).





## H1A \| OVERALL ACCURACY

+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Research Question     | Do Ss in the IMPASSE condition score higher across the entire task than those in the CONTROL group?                                                                                                                                                                                |
+=======================+====================================================================================================================================================================================================================================================================================+
| **Hypothesis**        | (H1) Participants in the IMPASSE condition will be more likely to correctly interpret the graph than those in the CONTROL condition.                                                                                                                                               |
+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Data**              | **data**: `df_items` where `q nin 6,9` (the 13 discriminating Qs ), `df_subjects`                                                                                                                                                                                                  |
|                       |                                                                                                                                                                                                                                                                                    |
|                       | **outcome**:                                                                                                                                                                                                                                                                       |
|                       |                                                                                                                                                                                                                                                                                    |
|                       | -   \[at item level\] : *accuracy* ( factor(incorrect/correct) from `score_niceABS` \[absolute score\]                                                                                                                                                                             |
|                       | -   \[subject level\]: accuracy (number of test phase qs correct from total `s_NABS`)                                                                                                                                                                                              |
|                       |                                                                                                                                                                                                                                                                                    |
|                       | **predictor**: `condition` \[between-subjects factor\]                                                                                                                                                                                                                             |
+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Analysis Strategy** | 1.  Wilcoxon-Rank Sum (Mann-Whitney) test on subject-level total accuracy of test phase (`s_NABS`)                                                                                                                                                                                 |
|                       | 2.  Mixed Logistic Regression\                                                                                                                                                                                                                                                     |
|                       |     `accuracy` \~ `condition` + (1 \| `subject` ) + (1 \| `question`)\                                                                                                                                                                                                             |
|                       |     model effect of condition on probability of correct response \[during test phase\] while accounting for subject (and item-level?) effects                                                                                                                                      |
+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Alternatives**      | -   Ordinal Mixed Logistic Regression on `scaled_score`                                                                                                                                                                                                                            |
|                       | -   OLS Linear Regression: bimodal distribution at tails makes the mean a poor predictor; LMs violate assumptions of normally distributed residuals; both absolute and scaled scores yield non-normal residuals; no transformation of the outcome variables yield normal residuals |
+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Notes**             | **Also exploring:**                                                                                                                                                                                                                                                                |
|                       |                                                                                                                                                                                                                                                                                    |
|                       | -   Hurdle model (mixture model w/ binomial + \[poisson or negbinom count; 0s from 1 DGP)                                                                                                                                                                                          |
|                       | -   Zero Inflated model (mixture model w/ binomial + poisson or negbinom count; 0s from 2 DGPs)                                                                                                                                                                                    |
|                       | -   Beta regression hurdle model? (mixture with location and scale parameters \[mean, variance\] and hurdles for floor and ceiling effects)                                                                                                                                        |
|                       | -   Other way to account for the severe bimodality?                                                                                                                                                                                                                                |
+-----------------------+------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

#### Setup
```{r}
#| label: SETUP-ACC

df_i = df_items %>% filter(q %nin% c(6,9)) %>% 
  dplyr::select(pretty_condition, accuracy, subject,q)

df_s <- df_subjects %>% 
   dplyr::select(pretty_condition, task_percent)

```

#### Visualize
```{r}
#| label: DESC-ACC

#:::::::: STACKED PROPORTIONAL BAR CHART
df_i %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = accuracy)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  # scale_fill_manual(values = paletteer::paletteer_d("ggthemes::calc", 2))+
  scale_fill_brewer(palette = "Set1")  +
  # facet_wrap(~pretty_mode) + 
  # coord_flip() +
  theme(legend.position="bottom")+
   labs(title = "DISTRIBUTION | Question Accuracy",
       x = "Condition",
       y = "Proportion of Questions",
       fill = "",
       subtitle="Impasse Condition yields a greater proportion of correct responses")

#:::::::: STACKED BAR CHART BY QUESTION
df_items %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = accuracy)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  facet_wrap(~q) +
   labs(title = "DISTRIBUTION | Accuracy by Question",
       x = "Condition",
       fill = "",
       subtitle="Q6 and Q9 are non-discriminative")

#:::::::: FACETED HISTOGRAM
stats = df_s %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(task_percent))
gf_props(~task_percent,
         fill = ~pretty_condition, data = df_s) %>%
  # gf_facet_grid(pretty_condition ~ pretty_mode) %>%
  gf_facet_grid(~pretty_condition) %>%
  gf_vline(data = stats, xintercept = ~mean, color = "red") +
  labs(x = "% Correct",
       y = "proportion of subjects",
       title = "DISTRIBUTION | Total Absolute Score (% Correct)",
       subtitle = "") + theme(legend.position = "blank")

#:::::::: RAINCLOUD WITH STATS
  
df <- df_s %>% mutate(task_percent = task_percent*100)

p <-   ggbetweenstats(data = df, x = pretty_condition, y = task_percent,
               plot.type = "box", type = "nonparametric", var.equal = FALSE,
               centrality.type = "parametric",
               # package = "RColorBrewer",
               # palette = "PRGn",
               centrality.point.args = list(color="black", size = 3, shape = 1),
               point.args = list(alpha=0), #suppress points
               ggplot.component = ## modify further with `{ggplot2}` functions
                list(
                  # aes(color = pretty_condition, fill = pretty_condition),
                  # scale_colour_manual(values = paletteer::paletteer_c("viridis::viridis", 3)),
                  # scale_fill_manual(values = paletteer::paletteer_c("viridis::viridis", 3)),
                  theme(axis.text.x = element_text(angle = 90)))
               ) +
  ggdist::stat_halfeye(
    alpha = 0.7, 
    point_colour = NA,
    adjust = .5, 
    width = .5, .width = 0, 
    justification = -.5) +
  geom_boxplot(
    alpha = 0.1,
    width = .2, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 2,
    alpha = .5,
    position = position_jitter(
      seed = 1, width = .08, height = 1.5
    )
  )  +
coord_flip() + theme_clean() + theme(legend.position = "blank")
p$layers[[3]]=NULL #remove default boxplot
e <- statsExpressions::two_sample_test(y = task_percent, x = pretty_condition, data = df,
                                type = "nonparametric", alternative = "less",
                                var.equal = FALSE)
#labels are layer 4
p + labs(title = "STUDY 3A | DISTRIBUTION of Total Score",
         y = "Percentage of correct responses across task", x = "",
         # caption=e$expression[[1]],
         subtitle = "Impasse condition yields greater variance and more high scores")

```


### Describe
```{r}
#| label: DESC2-ACC

title = "Descriptive Statistics of Response Accuracy (Total % Correct)"
tbl1 <- mosaic::favstats(~task_percent, data = df_s) 
tbl1 %>% kbl (caption = title) %>% kable_classic()


title = "Descriptive Statistics of Response Accuracy (Total % Correct) BY CONDITION"
tbl2 <- mosaic::favstats(task_percent ~ pretty_condition, data = df_s) 
tbl2 %>% kbl (caption = title) %>% kable_classic()

```

#### WILCOXON RANK SUM (Mann-Whitney Test) 

-   **Non parametric alternative** to t-test; compares median rather than mean by ranking data
-   Does not assume normality
-   Does not assume equal variance of samples (homogeneity of variance)

##### Test

```{r}
#| label: TEST-ACC

(w <- wilcox.test(df_s$task_percent ~ df_s$pretty_condition,
                 paired = FALSE, alternative = "less")) #less, greater
report(w)

```


##### Inference

**(not reported)**

Because the distribution of the outcome variable is not normally distributed, we evaluate the effect of CONDITION on ACCURACY via a non-parametric test. Consistent with our hypothesis, a Wilcoxon rank sum test (with continuity correction) on ACCURACY by CONDITION indicates that data in each condition likely come from different population distributions (W = 1243, p < 0.001; one-tailed), and that the distribution of the control condition is less (i.e. shifted to the left/ lower scores) than the impasse condition (\^{r} = -0.37, 95% CI \[-1.00, -0.22\]), a large-sized effect.

##### Visualize
```{r}
#| label: VIZ-TEST-ACC

ggbetweenstats( x = pretty_condition, y = task_percent, data = df_s,
                type = "nonparametric", var.equal = FALSE)

```



#### MIXED LOGISTIC REGRESSION

*Fit a mixed logistic regression (at the item level), predicting accuracy (absolute score) on all discriminating questions by condition; accounting for random effects of subject and item.*

##### Fit Model

```{r}
#| label: MODEL-FIT-ACC


## 0 | SETUP
#confirm 13 items [all discriminating items]
nrow(df_i) / nrow(df_s) == 13
#confirm all factors 
is.factor(df_i$q) && is.factor(df_i$subject) && is.factor(df_i$pretty_condition) && is.factor(df_i$accuracy)

## 1 | SETUP RANDOM INTERCEPT SUBJECT

#:: EMPTY MODEL (baseline, no random effect)
print("Empty fixed model")
m0 = glm(accuracy ~ 1, family = "binomial", data = df_i) 
# summary(m0)

#:: RANDOM INTERCEPT SUBJECT
print("Subject intercept random model")
mm.rS <- glmer(accuracy ~ (1|subject), data = df_i, family = "binomial")
# summary(mm.rS)

# :: TEST random effect
paste("AIC decreases w/ new model?", m0$aic > AIC(logLik(mm.rS)))
test_lrt(m0,mm.rS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(m0,mm.rS))$p[2])

#:: RANDOM INTERCEPT SUBJECT + ITEM
print("Subject Intercept + Item intercept random model")
mm.rSQ <- glmer(accuracy ~ (1|subject) + (1|q), data = df_i, family = "binomial")
#summary(mm.rSQ)

# :: TEST random effect
paste("AIC decreases w/ new model?", AIC(logLik(mm.rS)) > AIC(logLik(mm.rSQ)))
test_lrt(mm.rS, mm.rSQ) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rS, mm.rSQ))$p[2])


## 2 | ADD FIXED EFFECT CONDITION

print("FIXED Condition + Subject & Item random intercepts")
mm.CrSQ <- glmer(accuracy ~ pretty_condition + (1|subject) + (1|q) ,
                data = df_i, family = "binomial")
#summary(mm.CrSQ)

paste("AIC decreases w/ new model", AIC(logLik(mm.rSQ)) > AIC(logLik(mm.CrSQ)) )
test_lrt(mm.rSQ,mm.CrSQ) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rSQ,mm.CrSQ))$p[2])

# control = glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

```

*A likelihood ratio test indicates adding CONDITION as a fixed effect to a logistic regression model including a fixed effect random intercepts for SUBJECT and QUESTION explains more variance in the data than random-effects only model.*

##### Describe

```{r}
#| label: MODEL-DESC-ACC

#::::::::: SETUP
m <- mm.CrSQ

#::::::::: PRINT MODEL 

print("PREDICTOR MODEL")
summary(m)
print("SIGNIFICANCE TEST [non directional]")
car::Anova(m)
paste("MODEL INFO")
glance(m)

#:::::::: MANUAL ONE-SIDED SIGTEST 
#note: anova and chi square are always one-tailed, but that is independent of being one-sided
#https://www.ibm.com/support/pages/can-one-get-one-tailed-tests-logistic-regression-dividing-significance-levels-half
# one-sided (right tail) z test for B COEFFICIENT
#SANITY CHECK 2-tailed test should match the model output
# tt <- 2*pnorm(summary(m)$coefficients[2,3], lower.tail = F)
# paste("p value for two-tailed test, null B = 0 : ",round(tt,5))
# ot <- pnorm(summary(m)$coefficients[2,3], lower.tail = F)
# paste("BUT we want a one  directional, null: B <= 0: ",round(ot,5))

#:::::::: INTERPRET COEFFICIENTS

paste("LOG ODDS")
# se <- sqrt(diag(stats::vcov(m)))
# (tab <- cbind(Est = fixef(m),
#               LL = fixef(m) - 1.96 * se,
#               UL = fixef(m) + 1.96 * se))
tidy(m,   conf.int = TRUE, conf.level = 0.95, conf.method = "Wald")

paste("ODDS RATIOS")
# (e <- exp(tab))
tidy(m,   conf.int = TRUE, conf.level = 0.95, conf.method = "Wald", exponentiate = TRUE)

paste("PROBABILITIES")
#probability control = plogis(intercept)
#probability impasse = plogis(intercept + coefficient)

#FROM predict()
# newdata <- df_i %>% dplyr::select(pretty_condition, subject, q)
# preds <- predict(m, newdata = newdata, type = "response")
# preds <- cbind(newdata, preds)
# p <- preds %>% 
#   dplyr::select(pretty_condition, preds) %>% 
#   group_by(pretty_condition) %>% 
#   summarise(
#     median = median(preds),
#     se = sd(preds)/sqrt(n()),
#     lwr = median - 1.96*se,
#     upr = median + 1.96*se)
    
  
#FROM merTools
#setup df 
newdata <- df_i %>% dplyr::select(pretty_condition, subject, q)
#make predictions
preds <- predictInterval(m, newdata = newdata,
                              which = "fixed", #full, fixed or random for those only
                              type = "probability", #linear.prediction
                              stat = "median",
                              n.sims = 1000,
                              level = 0.80) #width of prediction interval
#join predictions to the new dataframe
preds <- cbind(newdata, preds)
#summarize
(summ_preds <- preds %>% 
  dplyr::select(pretty_condition, fit, lwr, upr) %>% 
  group_by(pretty_condition) %>% 
  summarise(
    median = median(fit),
    lower = median(lwr),
    upper = median(upr)
  )) 

```

##### INFERENCE

***(In Dissertation)***

We fit a mixed effects binomial logistic regression model with random intercepts for subjects and questions.  Note that we choose to model these data at the item (i.e. question) rather than than subject (i.e. total score) level because the structure of a mixed effects model allows us to differentiate between random variance introduced by individual subjects and questions, versus the expected systematic variance of CONDITION. A likelihood ratio test indicates that a model including a fixed effect of CONDITION explains significantly more variance in the data than an intercepts-only baseline model ($\chi^2 (3) = 18.66, p < 0.001$). The explanatory power of the entire model is substantial ($conditional \ R^2 = 0.89$) and the part related to the fixed effect CONDITION ($marginal \ R^2$) explains 15\% of variance. **Consistent with our hypothesis, the impasse condition substantially increases the odds of a correct response.** Across the entire task participants in the impasse condition were 75 times more likely to offer a correct response, compared with those in the control condition ( $e^{\beta_1} = 75.51, p < 0.001$, $95 \% \  CI [8.46, 674]$). Based on the fixed effect of CONDITION, The model predicts that the probability of a correct response in the control condition is effectively 0% (95% CI [6.8e-4, 0.02]), and the probability of a correct response in the impasse condition increases to 24% (95% CI [0.06, 0.60]). 

```{r}
#| label: MODEL-TBL-ACC

#SJPLOT | MODEL | TABLE
tab_model(m)

# #MODEL SUMMARY | save latex table
# models <- list("odds ratios" = m, "(log odds)" = m)
# notes = list("* p < 0.05, ** p < 0.01, *** p < 0.001",
#              '$sigma^{2}$ = 3.29" N(subject) = 126 $\tau_{00}$(subject) = 22.22 N(question) = 13 $\tau_{00}$(question) = 0.31'
#                )
# 
# modelsummary(models,
#              exponentiate = c(TRUE, FALSE),
#              shape = term ~ model + statistic,
#              fmt = 2, #two digits w/ trailing zero
#              estimate  = "{estimate} {stars}",
#              statistic = "conf.int",
#              gof_map = c("AIC", "sigma"),
#              gof_omit = 'RMSE|ICC|BIC',
#              coef_rename = c("pretty_conditionimpasse" = "Condition[impasse]"),
#              title = 'Accuracy ~ Condition (Mixed Logistic Regression)',
#              notes = notes)
#              # output = "analysis/SGC3A/models/tables/GLMER_OverallAccuracy_lab.tex")
# #              # coef_omit = "Intercept",




```

##### Visualize

```{r}
#| label: MODEL-VIS-ACC

## | PLOT PARAMETERS 


#SJPLOT | MODEL | lOG ODDS
plot_model(m, transform = NULL,
           vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.05, #manually adjust to account for directional test
           ci.lvl = 0.95 ) + #manually adjusted for directional test   
  labs(title = "Model ESTIMATE | Odds Ratio",
       subtitle = "",
       x = "Condition")


#SJPLOT | MODEL | ODDS RATIO
plot_model(m, transform = "exp",
           vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.05, #manually adjust to account for directional test
           ci.lvl = 0.95 ) + #manually adjusted for directional test   
  labs(title = "Model ESTIMATE | Log Odds",
       subtitle = "",
       x = "Condition")

#EASYSTATS | MODEL | ODDS RATIO
# result <- model_parameters(m, exponentiate = TRUE, component = "all")
# plot(result) + labs(
#     title = "Model ESTIMATE | ODDS RATIO"
#   )

## | PLOT TESTS
# result <- equivalence_test(m, rule = "classic", ci=0.9) #classic[tost], , bayes
# plot(result)

## | PLOT PREDICTIONS

#SJPLOT | MODEL | PROBABILITIES
# plot_model(m, type="eff",
#            show.intercept = TRUE,
#            show.values = TRUE,
#            title = "Model Prediction | Probability of Accurate Response",
#            axis.title = c("Condition","Probability of Accurate Response"))

#SJPLOT | MODEL | PROBABILITIES
plot_model(m, type = "pred")[[1]]  +
  ylim(0,1) +
  labs(
    title = "Model PREDICTION | Probability of Accurate Response",
    subtitle = "Impasse increases probability of correct response",
    y = "Probability of Correct Response", x = "Condition"
  )

#GGDIST | MODEL | PREDICTED PROBABILITIES
preds %>% 
  ggplot(aes( x = fit, y = pretty_condition, fill = pretty_condition)) + 
  stat_halfeye(alpha = 0.5, normalize = "xy") + 
  xlim(0,0.3) + theme_clean() + labs(
    title = "Model PREDICTION | Probability of Accurate Response",
    subtitle = "TODO check preds to see if fixed or includes random"
  )

# SIMULATE FIXED EFFECTS
# simulate values of fixed effects 
# (feEx <- FEsim(m,  oddsRatio = FALSE, n.sims = 1000))
# PLOT estimates of fixed effects
# plotFEsim(feEx) +
#   theme_bw() + labs(title = "Coefficient Plot of InstEval Model",
#                     x = "Median Effect Estimate")

# SIMULATE RANDOM EFFECTS
# simulate values of random effects
# reEx <- REsim(m)
# PLOT estimates of random effects
# plotREsim(reEx)
```


##### Diagnostics

```{r}
#| label: MODEL-DIAG-ACC
#| warning: false
#| message: false

# print("SANITY CHECK REPORTING")
# report(m)

# print("MODEL PERFORMANCE")
# performance(m)

print("DIAGNOSTICS")
check_model(m)

```


##### Sanity Check :: Bayesian

```{r}
#| label: BAYES-MODEL-FIT-ACC


# ## 0 | SETUP
# #confirm 13 items [all discriminating items]
# nrow(df_i) / nrow(df_s) == 13
# #confirm all factors 
# is.factor(df_i$q) && is.factor(df_i$subject) && is.factor(df_i$pretty_condition) && is.factor(df_i$accuracy)
# 
#
# 
# print("FIXED Condition + Subject & Item random intercepts")
# Bmm.CrSQ <- brm( accuracy ~ pretty_condition + (1|subject) + (1|q), 
#                  data = df_i, 
#                  family = "bernoulli",
#                  chains = 4, iter = 2000, warmup = 1000,
#                  cores = 4, seed = 1234,
#                  backend = "cmdstanr",
#                  file ="analysis/SGC3A/models/sgc3a_brms_acc_Bmm.CrSQ_LAB.rds")
# 
# #get Priors 
# # describe_priors(Bmm.CrSQ)
# 
# #GRAPHICAL POSTERIOR PREDICTION CHECKS
# pp_check(Bmm.CrSQ)
# 
# #DESCRIBE MODEL
# (d <- describe_posterior(ci=.95, Bmm.CrSQ))
# 
# #SEE MODEL
# plot(pd(Bmm.CrSQ))
# #convert to a pd value
# (pds <- pd_to_p(d$pd))

```

*A likelihood ratio test indicates adding CONDITION as a fixed effect to a logistic regression model including a fixed effect random intercepts for SUBJECT and QUESTION explains more variance in the data than random-effects only model.*


```{r}

# #::::: GGDIST POSTERIOR PROBABILITY OF RESPONSE
# ##WORKING
# ## VIS probability of correct response
# #TAKES A REALLY LONG TIME
# 
# #1 | get draws
# draws <- df_i %>%
#   data_grid(pretty_condition, subject, q) %>%
#   add_fitted_draws(Bmm.CrSQ,
#                    # n = 100,
#                    # dpar = TRUE,
#                    # transform = TRUE, #gives prob%, otherwise OR
#                    re_formula = NA)
# # draws %>% write_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")
# 
# #OR load from file
# # draws <- read_rds("analysis/SGC3A/models/draws_BB.catCrSQ.rds")
# 
# #2| VISUALIZE PREDICTIONS | GGDIST
# ##TODO figure out height normalization.
# ##do it with much smaller number of draws 
# #TODO adjust bandwidth/smoothing? + put on same line + 
# #TAKES A REAAALY LONG TIME
# # draws %>% sample_n(1000) %>% 
# #   ggplot(aes(x = .value,  y = 0, fill = pretty_condition)) +
# #   stat_slab(width = c(.95), alpha = 1, normalize="xy") +
# #   #normalize = all, panels, xy, groups, none
# #   xlim(0,1) + labs(
# #     title = "Model Predicted Probability of Correct Response",
# #     x = "probability of correct response",
# #     y = "Interpretation"
# #   ) +  theme_clean() #+ ggeasy::easy_remove_legend() + ggeasy::easy_remove_y_axis()
# # #TO PLOT ON THE SAME LINE, INCLUDE Y = 0 in aes and ggeasy::remove_y_axis()



```

## H1A \| OVERALL INTERPRETATION STATE

**Do Ss in the IMPASSE condition offer less-orthogonal interpretations across the test phase questions?**

While absolute accuracy score tells us whether a participant successfully interpreted the coordinate system, it doesn't allow us to differentiate between different kinds of incorrect interpretations. Here we examine the (categorical) interpretation state type based on the nature of subject's response, and determine if these interpretations differ by experimental condition. State is a 3-category variable that groups the following interpretations:

-   **"orthogonal"** \[reference category\] includes orthogonal and satisficing responses ==\> indicates a primarily *orthogonal* state of coordinate system understanding

-   **"other"** includes: blank, reference point, responses that can't be classified (including selecting all datapoints), =\> indicates an uncertain or unidentifiable state of coordinate system understanding, but one that is distinctly *not* orthogonal nor triangular

-   **"angular"** includes 'lines connecting' responses as well as responses that include both orthogonal *and* triangular answers =\> indicates some degree of angular/triangular coordinate understanding

-   **"triangular"** includes correct triangular and 'lines connecting' responses as well as responses that include both orthogonal *and* triangular answers =\> indicates some degree of angular/triangular coordinate understanding


+-----------------------+-------------------------------------------------------------------------------------------------------------------------------------------------+
| Research Question     | Does Ss in the impasse condition produce less orthogonal responses across questions?                                          |
+=======================+=================================================================================================================================================+
| **Hypothesis**        | H1A \| Ss in the IMPASSE condition will have a higher likelihood of producing unknown and triangle-like response states across all items |
+-----------------------+-------------------------------------------------------------------------------------------------------------------------------------------------+
| **Data**              | -   data: `df_items` where `q nin 6,9` (13 discriminant test phase items)                                                              |
|                       | -   outcome: `state` ( 3 level factor from high_interpretation )                                                                                |
|                       | -   predictor: `condition` \[between-subjects factor\]                                                                                          |
+-----------------------+-------------------------------------------------------------------------------------------------------------------------------------------------+
| **Analysis Strategy** | 1.  MIXED Multinomial (Logistic) Regression on state predicted by condition                                                                     |
|                       |                                                                                                                                                 |
|                       | Alternative:                                                                                                                                    |
|                       |                                                                                                                                                 |
|                       | -   MIXED Ordinal regression on state (doesn't meet proportional odds assumption-I think)                                                       |
|                       | -   MIXED Multinomial or Ordinal regression on high_interpretation (some cells are 0, produces problems)                                        |
+-----------------------+-------------------------------------------------------------------------------------------------------------------------------------------------+

#### Setup

```{r}
#| label: SETUP-STATE

df_i = df_items %>% filter(q %nin% c(6,9)) %>% dplyr::select(q,subject,state,pretty_condition)

```

#### Visualize
```{r}
#:::::::: STACKED BAR CHART
df_i %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = state)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("ggthemes::calc", 4))+
  # facet_wrap(~pretty_mode) + 
  theme(legend.position = "bottom")+
   labs(title = "DISTRIBUTION of Interpretation",
       x = "Condition",
       y = "Proportion of Questions",
       fill = "",
       subtitle="")

#:::::::: STACKED BAR CHART
df_i %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = state)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("ggthemes::calc", 4))+
  facet_wrap(~q) +
   labs(title = "Interpretation by Question",
       x = "Condition",
       fill = "",
       subtitle="")

```

A proportional bar chart visualizing the proportion of each interpretations type by condition for each data collection modality (left/right facet) reveals that the pattern of responses are similar regardless of the data collection modality, by differ by condition. In the impasse condition, there are more triangular responses than in control. In the impasse condition, there are also more positive transition (i.e. triangle-like) and neutral (ie. blank or uncertain response types) than in the control condition.

#### Describe
```{r}

#::::::::::::DESCRIPTIVES

table(df_i$state, df_i$pretty_condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns


(t <- table(df_i$state, df_i$pretty_condition) %>% addmargins(2) %>% #display sum for row
      addmargins(1)) #sanity check sum of columns

```

#### MIXED MULTINOMIAL REGRESSION


*Does condition affect the response state of of items across the task?*

*Fit a MIXED logistic regression predicting interpretation state (k=3) by condition(k = 2).*

-   Can use mclogit mblogit() with random effect *or* bayesian brms package b/c nlme, lme4 don't support random effects on multinomial (ie no categorical family on glmer())

-   Alternative would be to manually run [k-1] X binomial mixed models \[should compare outcomes\]

-   [k-1] equations will be estimated (# categories - 1); each representing the odds of answering in that particular interpretation (vs) the reference category (orthogonal answer) \[essentially a series of binary logistic regressions, but instead of incorrect/correct, comparing \[reference category\] vs \[this category\])

-   For *each* equation:

    -   $\beta_{0}$ *= Log Odds of \[this category type vs. reference category type) response in CONTROL condition*
    -   $e^{\beta_{0}}$ *= ODDS of \[this category type vs. reference category type\] response in CONTROL condition*
    -   $\beta_{1}$ *=* $\beta_{1impasse}$ *Log Odds (Log OR; change in odds for \[this category\] type response in impasse (vs) control \[log scale\])*
    -   $e^{\beta_{1}}$ *= ODDS RATIO of \[this. vs reference category type\] response in IMPASSE (vs) CONTROL*
    -   Two-tailed NHST *Null hypothesis:* $\beta_{impasse} = 0$ *the odds for \[this category of response vs. reference\] are not different for IMPASSE condition*
    -   *Alternative hypothesis:* $\beta_{impasse} \ne 0$ *the odds of \[this category of response vs. reference\] increases or decreases for IMPASSE condition*

##### Fit Model \[mblogit\]

```{r}
#| label: FIT-MBLOGIT-STATE

#https://www.elff.eu/software/mclogit/manual/mblogit/
#"baseline category logit" model matches multinom()

#check reference level 
print("Categories (first is reference)")
levels(df_i$state)

#FIT EMPTY MODEL
# print("EMPTY MODEL")
mm.cat.rSQ <- mblogit(state ~ 1 , 
                      random = list( ~ 1|subject, ~1|q), 
                      data = df_i)
#summary(mm.cat.rSQ)

#FIT PREDICTOR MODEL
# print("PREDICTOR MODEL")
mm.cat.CrSQ <- mblogit(state ~ pretty_condition , 
                  random = list( ~ 1|subject, ~1|q), 
                  data = df_i)
# summary(mm.cat.CrSQ)

#COMPARE MODEL FIT
paste("AIC wth predictor is lower than empty model?", AIC(mm.cat.rSQ) > AIC(mm.cat.CrSQ))
test_lrt(mm.cat.rSQ, mm.cat.CrSQ)
```
**NOTE: The empty model DOES NOT CONVERGE**
##### Describe

```{r}
#| label: DESC-MBLOGIT-STATE

m <- mm.cat.CrSQ

#DESCRIBE MODEL
summary(m)

#INTERPRET COEFFICIENTS
paste("LOG ODDS")
tidy(m,   conf.int = TRUE, conf.level = 0.95, conf.method = "Wald")
paste("ODDS RATIOS")
tidy(m,   conf.int = TRUE, conf.level = 0.95, conf.method = "Wald", exponentiate = TRUE)

# paste("MODEL INFO")
# glance(m)

#PERFORMANCE
performance(m)

```


##### Inference

**NOTE: For reference only; this model raised NaN errors**


```{r}
#| label: MODEL-TBL-MBLOGIT-STATE

#SJPLOT | MODEL | TABLE
tab_model(m)

# #MODEL SUMMARY | save latex table
# models <- list("odds ratios" = m, "(log odds)" = m)
# notes = list("* p < 0.05, ** p < 0.01, *** p < 0.001",
#              '$sigma^{2}$ = 3.29" N(subject) = 126 $\tau_{00}$(subject) = 22.22 N(question) = 13 $\tau_{00}$(question) = 0.31'
#                )
# 
# modelsummary(models,
#              exponentiate = c(TRUE, FALSE),
#              shape = term ~ model + statistic,
#              fmt = 2, #two digits w/ trailing zero
#              estimate  = "{estimate} {stars}",
#              statistic = "conf.int",
#              gof_map = c("AIC", "sigma"),
#              gof_omit = 'RMSE|ICC|BIC',
#              coef_rename = c("pretty_conditionimpasse" = "Condition[impasse]"),
#              title = 'Accuracy ~ Condition (Mixed Logistic Regression)',
#              notes = notes)
#              # output = "analysis/SGC3A/models/tables/GLMER_OverallAccuracy_lab.tex")
# #              # coef_omit = "Intercept",




```

##### Visualize

```{r}
#| label: VIS-MBLOGIT-STATE


## | PLOT PARAMETERS 

#SJPLOT | MODEL | ODDS RATIO
plot_model(m, 
           transform = "exp",
           vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE)


#TODO SEPARATE THIS BY EQUATION 
# ms <- model_parameters(Bmm.cat.CrSQ, component = "conditional")
# m1 <- ms %>% filter(str_detect(Parameter, "muother"))
# plot(m1)

#EASYSTATS | MODEL | ODDS RATIO
result <- model_parameters(m, exponentiate = TRUE, component = "all")
plot(result, show_labels = TRUE, n_columns = 3)

# result <- simulate_parameters(m)
# plot(result, stack = FALSE)

## | PLOT TESTS
result <- equivalence_test(m, rule = "classic", ci=0.9) #classic[tost], , bayes
plot(result)

## | PLOT PREDICTIONS

#SJPLOT | MODEL | PROBABILITIES
# plot_model(m, type="eff",
#            show.intercept = TRUE,
#            show.values = TRUE,
#            title = "Model Prediction | Probability of Accurate Response",
#            axis.title = c("Condition","Probability of Accurate Response"))
# 
# #PLOT MODEL PREDICTION
# plot_model(m, type = "pred")[[1]] + 
#   ylim(0,1) + labs(
#     title = "Model Prediction | Probability of Accurate Response",
#     subtitle = "Impasse increases Probability of Correct Response"
#   )

#TODO EMMEANS for the estimated marginal means

```

##### Diagnostics

```{r}

# check_model(m)

```



##### Fit Model \[brms\]

```{r}
#| label: FIT-BRMS-STATE

#BAYESIAN RANDOM ONLY
Bmm.cat.rSQ <- brm( state ~ 1 + (1|subject) + (1|q), 
                 data = df_i, 
                 family = "categorical",
                 chains = 4, iter = 2500, warmup = 1000,
                 cores = 4, seed = 1234,
                 save_pars = save_pars(all = TRUE),
                 # backend = "cmdstanr",
                 file ="analysis/SGC3A/models/sgc3a_brms_state_Bmm.cat.rSQ_LAB.rds")


#UNINFORMATIVE PRIOR BAYESIAN MIXED VERSION
# flat_Bmm.cat.CrSQ <- brm( state ~ pretty_condition + (1|subject) + (1|q), 
#                  data = df_i, 
#                  family = "categorical",
#                  chains = 4, iter = 2500, warmup = 1000,
#                  cores = 4, seed = 1234,
#                  save_pars = save_pars(all = TRUE),
#                  # backend = "cmdstanr",
#                  file ="analysis/SGC3A/models/sgc3a_brms_state_FLAT_Bmm.cat.CrSQ_LAB.rds")


# determine default priors 
# prior_summary(flat_Bmm.cat.CrSQ)

#set priors [see justification, below]
inf_priors <- c(
  # too strong?
  # prior(normal(-6.91, 0.201),  class = "Intercept", dpar = "muangular"),
  # prior(normal(-6.91, 0.201),  class = "Intercept", dpar = "muother"),
  # prior(normal(-6.91, 0.201),  class = "Intercept", dpar = "mutriangular"),
  #prior on INTERCEPTS 
  #25% chance of each answer in control, scale = from 0.01 to 62%
  prior(normal(-1.1, 1.5),  class = "Intercept", dpar = "muangular"),
  prior(normal(-1.1, 1.5),  class = "Intercept", dpar = "muother"),
  prior(normal(-1.1, 1.5),  class = "Intercept", dpar = "mutriangular"),
  #prior on COEFFICIENT
  #likely to change odds between 0 and 2.4
  prior(normal(0, 2.42), class = b, coef="pretty_conditionimpasse", dpar = "muangular"),
  prior(normal(0, 2.42), class = b, coef="pretty_conditionimpasse", dpar = "muother"),
  prior(normal(0, 2.42), class = b, coef="pretty_conditionimpasse", dpar = "mutriangular")
)

#INFORMATIVE PRIORS
Bmm.cat.CrSQ <- brm( state ~ pretty_condition + (1|subject) + (1|q), 
                 data = df_i, 
                 family = "categorical",
                 prior = inf_priors,
                 chains = 4, iter = 2500, warmup = 1000,
                 cores = 4, seed = 1234,
                 save_pars = save_pars(all = TRUE),
                 control = list(adapt_delta = 0.98),  # to deal with divergent transitions
                 # backend = "cmdstanr",
                 file ="analysis/SGC3A/models/sgc3a_brms_state_Bmm.cat.CrSQ_LAB.rds"
                 )

#a bayes factor model comparison of the flat vs informative prior models suggest convicing evidence that 
#informative prior model is a better fit
# bayesfactor(Bmm.cat.CrSQ, flat_Bmm.cat.CrSQ)

# PRIORS LOGIC 
# https://www.bayesrulesbook.com/chapter-13.html#building-the-logistic-regression-model

#expectation for probability of _better_ response [in control]?
#very low probability center: 0.1% [very low]; as logodds = logit(0.001) = -6.91
#range from 0 to 55%  logit(0.55) = 0.201
#probability of 0.1 to 55% is equivalent to [logodds] -6.91 +/ 2* 0.201
#therefore... prior for intercept => Normal(−6.91, 0)


#expectation for probability of _better_ response [in impasse]?
#increases probablity from 0 % 
# 0 [very low]; as OR  = exp(0) = 1
#range from 0 to 90%  exp(0.9) = 2.46
#probability of 0 to 90% is equivalent to [ODDS scale] 1 +/ 2* 2.42
#on log odds scale ? [0, ]
#therefore... prior for intercept => Normal(1, 2.42)
                             # prior = normal(0.07, 0.035),


```


##### Describe

```{r}
#| label: DESC-BRMS-STATE

# best model
m <- Bmm.cat.CrSQ

#::::::::: PRINT MODEL 

print("PREDICTOR MODEL")
summary(m)
(d <- describe_posterior(ci=.95, Bmm.cat.CrSQ))

print("BAYES FACTOR [comparison to null]")
#think of this like the anova(model) to get p values for each predictor
#has to recompile the models with rstan. total drag
(b <- bayesfactor(Bmm.cat.rSQ, m))

print("DESCRIBE POSTERIOR")

#:::::::: INTERPRET COEFFICIENTS
# se <- sqrt(diag(stats::vcov(m)))
# # table of estimates with 95% CI
# (tab <- cbind(Est = fixef(m),
#               LL = fixef(m) - 1.96 * se,
#               UL = fixef(m) + 1.96 * se))

paste("LOG ODDS")
(l <- describe_posterior(m))
# (tm <- tidy(m,   conf.int = TRUE))

paste("ODDS RATIOS")
(e <- model_parameters(m, exponentiate = TRUE))
# tidy(m,   conf.int = TRUE, exponentiate = TRUE)
# tm %>% mutate(
#   OR.est = exp(estimate),
#   exp.low = exp(conf.low),
#   exp.high = exp(conf.high)
# ) %>% dplyr::select(effect, component, group, term, OR.est, exp.low, exp.high)

paste("PROBABILITIES")

#PREDICT METHOD
newdata <- df_i %>% dplyr::select(pretty_condition, subject, q)
preds <- predict(m, newdata = newdata, type = "response")
preds <- cbind(newdata, preds)
#lengthen data frame to handle multinomial
preds <- preds %>% 
  dplyr::select(-subject, -q) %>% #marginalize over subject and q
  pivot_longer(
  cols = !pretty_condition,
  values_to = "preds",
  names_to = "state",
) 

(p <- preds %>% 
  group_by(pretty_condition, state ) %>%
  summarise(
    median = median(preds),
    se = sd(preds)/sqrt(n()),
    lwr = median - 1.96*se,
    upr = median + 1.96*se))

##DRAWS METHOD
# GENERATE draws from model
# draws <- df_i %>%
#   data_grid(pretty_condition, subject, q) %>% 
#   add_fitted_draws(Bmm.cat.CrSQ,
#                    # n = 100,
#                    # dpar = TRUE,
#                    # transform = TRUE, #gives prob%, otherwise OR
#                    re_formula = NA)
# # draws %>% write_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")
# 
# #OR load from file
# # draws <- read_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")
# 
# # SUMMARIZE draws from model
# (k <- kable(draws %>%
#   dplyr::select(pretty_condition, .category, .value) %>%
#   group_by(pretty_condition, .category) %>%
#   median_hdci(.value), digits = 4, col.names =
#     c("Condition","Category", "Probability","Lower Cred.I","Upper Cred.I", "CI Width", "Point Type", "Interval Type")) %>%
#   kable_styling())

```


##### INFERENCE

[REPORT POSTERIOR MEDIAN $\exp_{beta}$, 95 \% credible interval, \% probability of direction]

We fit a (bayesian) multinomial logistic regression model with random intercepts for subjects and questions. A Bayes Factor model comparison (against a random intercepts-only model) indicates extreme evidence for a main effect of CONDITION (BF = 9.31e+15). 
**Consistent with our hypothesis, the impasse condition substantially increases the odds of transitional interpretations.** 

Across the entire task participants in the impasse condition were 11 times more likely to offer an 'unknown' rather than orthogonal response compared with those in the control condition ( $e^{\beta_1} = 10.93, 95 \% CI [5.59], pd = 100\%$). Participants in the impasse condition were 11 times more likely to offer an 'angular' rather than orthogonal response compared with those in the control condition ( $e^{\beta_1} = 10.93, 95 \% CI [3.55, 40.54], pd = 100\%$), and 33 times more likely to offer an 'triangular' rather than orthogonal response compared with those in the control condition ( $e^{\beta_1} = 33.38, 95 \% CI [6.36, 194.83], pd = 100\%$). 

The impasse condition decreases the probability of an orthogonal response by 54%
The impasse condition increases the probability of an uknown/uncertain response by 16%
The impasse condition increases the probability of an angular response by 1%
the impasse condition increase the probability of a triangular response by 37%


```{r}
#| label: TBL-BRMS-STATE

#SJPLOT | MODEL | TABLE
tab_model(m)

# #MODEL SUMMARY | save latex table
# models <- list("odds ratios" = m, "(log odds)" = m)
# notes = list("* p < 0.05, ** p < 0.01, *** p < 0.001",
#              '$sigma^{2}$ = 3.29" N(subject) = 126 $\tau_{00}$(subject) = 22.22 N(question) = 13 $\tau_{00}$(question) = 0.31'
#                )
# 
# modelsummary(models,
#              exponentiate = c(TRUE, FALSE),
#              shape = term ~ model + statistic,
#              fmt = 2, #two digits w/ trailing zero
#              estimate  = "{estimate} {stars}",
#              statistic = "conf.int",
#              gof_map = c("AIC", "sigma"),
#              gof_omit = 'RMSE|ICC|BIC',
#              coef_rename = c("pretty_conditionimpasse" = "Condition[impasse]"),
#              title = 'Accuracy ~ Condition (Mixed Logistic Regression)',
#              notes = notes)
#              # output = "analysis/SGC3A/models/tables/GLMER_OverallAccuracy_lab.tex")
# #              # coef_omit = "Intercept",
# 
# 
# ## POINT ESTIMATES IN PROBABILITY
# #1 | get draws
# draws <- df_i %>%
#   data_grid(pretty_condition, subject, q) %>%
#   add_fitted_draws(Bmm.cat.CrSQ,
#                    # n = 100,
#                    # dpar = TRUE,
#                    # transform = TRUE, #gives prob%, otherwise OR
#                    re_formula = NA)
# draws %>% write_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")
# 
# #2 | SUMMARIZE draws 
# k <- kable(draws %>%
#   select(pretty_condition, .category, .value) %>%
#   group_by(pretty_condition, .category) %>%
#   median_hdci(.value), digits = 2, col.names = 
#     c("Condition","Category", "Probability","Lower Cred.I","Upper Cred.I", "CI Width", "Point Type", "Interval Type")) %>% 
#   kable_styling()
# k

#COMPARISONS
# c <- draws %>% 
#   # dplyr::select(pretty_condition, .category, .value) %>%
#   compare_levels(variable = .value, by = pretty_condition, 
#                  comparison = list(c("control","impasse"))) 
#                                    # c("adots","interval"),
#                                    # c("adots","mean"),
#                                    # c("adots","text"),
#                                    # c("density","interval"),
#                                    # c("density","mean"),
#                                    # c("density","text"),
#                                    # c("interval","mean"),
#                                    # c("interval","text"),
#                                    # c("mean","text")
#                                    
# c %>%
#   ggplot(aes(x = .value, y = reorder(x =pretty_condition, X = .value)))+
#   stat_interval(.width = .95, color = "black") +
#   geom_vline(xintercept = 0)+
#   theme_bw()+
#   # coord_cartesian(xlim = c(-.5,1)) +
#   theme_tidybayes() 
# comps

```
##### Visualize

```{r}
#| label: VIS-BRMS-STATE


## | PLOT PARAMETERS 

#SJPLOT | MODEL | ODDS RATIO
# plot_model(m, vline.color = "red",
#            show.intercept = TRUE,
#            show.values = TRUE,
#            p.threshold = 0.1, #manually adjust to account for directional test
#            ci.lvl = 0.90 ) + #manually adjusted for directional test
#   labs(title = "Model Estimate | Odds Ratio",
#        subtitle = "",
#        x = "Condition")


#EASYSTATS | MODEL | ODDS RATIO
result <- model_parameters(Bmm.cat.CrSQ, exponentiate = TRUE, component = "all")
plot(result, show_intercept = TRUE, show_labels = TRUE) 
# + theme_clean()

# 
# result <- estimate_density(m,exponentiate = TRUE)
# plot(result,  stack = FALSE, priors = TRUE)

## | PLOT TESTS
result <- equivalence_test(m, rule = "classic", ci=0.9) #classic[tost], , bayes
plot(result)

result <- rope(m)
plot(result)

(result <- pd(m,exponentiate = TRUE))
plot(result, show_intercept = TRUE, show_labels = TRUE)

## | PLOT PREDICTIONS

#SJPLOT | MODEL | PROBABILITIES
# plot_model(m, type="eff",
#            show.intercept = TRUE,
#            show.values = TRUE,
#            title = "Model Prediction | Probability of Accurate Response",
#            axis.title = c("Condition","Probability of Accurate Response"))
# 
# #PLOT MODEL PREDICTION
plot_model(m, type = "pred")

#TODO EMMEANS for the estimated marginal means

#TODO OUTPUT TABLE 
#https://arelbundock.com/posts/modelsummary_multinomial_logit/
# modelsummary(m)

```

```{r}
#| label: GGDIST-STATE

#::::: GGDIST POSTERIOR PROBABILITY OF RESPONSE
##WORKING
# https://mjskay.github.io/ggdist/reference/stat_slab.html
## VIS probability of correct response
#TAKES A REALLY LONG TIME

#1 | get draws
draws <- df_i %>%
  data_grid(pretty_condition, subject, q) %>%
  add_fitted_draws(Bmm.cat.CrSQ,
                   # n = 100,
                   # dpar = TRUE,
                   # transform = TRUE, #gives prob%, otherwise OR
                   re_formula = NA)
# draws %>% write_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")

#OR load from file
# draws <- read_rds(file = "analysis/SGC3A/models/draws/draws_BB.catCrSQ.rds")

#2| VISUALIZE PREDICTIONS | GGDIST
##TODO figure out height normalization.
##do it with much smaller number of draws 
#TODO adjust bandwidth/smoothing? + put on same line + 
#TAKES A REAAALY LONG TIME
d <- draws %>% sample_n(10) %>% 
  ggplot(aes(x = .value,  y = .category, fill = pretty_condition)) +
  stat_slab(width = c(.95), alpha = 0.5, normalize="xy") +
  # stat_halfeye(.width = c(.95), alpha = .6,interval_alpha = 0,point_alpha = 0, normalize = "groups") +
  #   #normalize = all, panels, xy, groups, none
  xlim(0,1) + labs(
    title = "Model Predicted Probability of Correct Response",
    x = "probability of correct response",
    y = "Interpretation"
  ) +  theme_clean() #+ ggeasy::easy_remove_legend() + ggeasy::easy_remove_y_axis()
# # #TO PLOT ON THE SAME LINE, INCLUDE Y = 0 in aes and ggeasy::remove_y_axis()
# # 
# # ggsave(d, filename = "figures/sgc3a_BBm.cat.CrSQ_lab_posterior.svg", width = 6, height =4)
d
```



##### Diagnostics

```{r}

#CHECK Fit of posterior predictive to data
pp_check(Bmm.cat.CrSQ, ndraws=1000)

#CHECK posterior vs. priors
result <- estimate_density(Bmm.cat.CrSQ)
plot(result, stack = FALSE, priors= TRUE)

#CHECK model
plot(Bmm.cat.CrSQ)

```


##### COMPARE MBLOGIT to BRMS

```{r}
#| label: COMPARE-STATE-MODELS
compare_models(mm.cat.CrSQ, Bmm.cat.CrSQ)
```
The predictions of the manual, frequentist mixed multinomial and bayesian mixed multinomial models are comparable. 


## H1B \| Q1 ACCURACY

**Do Ss in the IMPASSE condition have a higher likelihood of producing a correct response to the first question?**

The graph comprehension task includes 15 questions completed in sequence. But the first question the reader encounters (Q1) is the most important, as it is their *first exposure* to the unconventional triangular coordinate system.

+-----------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Research Question     | Does the frequency of correct (vs) incorrect responses on the first question differ by condition? \[Is response accuracy independent of condition?\]                                                                                   |
+=======================+========================================================================================================================================================================================================================================+
| **Hypothesis**        | H1A \| Ss in the IMPASSE condition will have a higher likelihood of correctly responding to the first question than those in the CONTROL condition                                                                                     |
+-----------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Data**              | -   data: `df_items` where `q == 1`                                                                                                                                                                                                    |
|                       | -   outcome: *accuracy* ( factor(incorrect/correct) from `score_niceABS` \[absolute score\]                                                                                                                                            |
|                       | -   predictor: `condition` \[between-subjects factor\]                                                                                                                                                                                 |
+-----------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Analysis Strategy** | 1.  Logistic Regression on `accuracy` predicted by `condition`                                                                                                                                                                         |
|                       |     -   account for difference in odds of correct score by condition                                                                                                                                                                   |
|                       |                                                                                                                                                                                                                                        |
|                       | Alternatives:                                                                                                                                                                                                                          |
|                       |                                                                                                                                                                                                                                        |
|                       | -   Chi-Square test of independence on outcome `accuracy` by `condition`                                                                                                                                                               |
+-----------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Notes**             | -   CHIQ SQR is simplest method to examine independence of two categorical factors; LOGISTIC REGRESSION is recommended for binomial \~ continuous; though with regression we can quantify the size of the effect and overall model fit |
|                       | -   independence assumption : (CHI SQR) as we only consider responses on the first question, each observation corresponds to an individual subject, and are thus independent                                                           |
|                       | -   cell frequency : (CHI SQR) expected frequency in each cell of the contingency table is greater than 5 (more than 5 correct , more than 5 incorrect responses)                                                                      |
+-----------------------+----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

#### Setup
```{r}
#| label: SETUP-Q1ACC

df <- df_items %>% filter(q==1)  %>% dplyr::select(accuracy, pretty_condition)

```

#### Describe 
```{r}
#| label: VIS-Q1ACC

#:::::::: STACKED BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = accuracy)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  # facet_wrap(~pretty_mode) + 
   labs(#y = "Correct Response on Q 1",
       title = "Q1 Accuracy",
       x = "Condition",
       fill = "",
       subtitle="Impasse Condition yields a greater proportion of correct responses")

```

A proportional bar chart visualizing the proportion of incorrect (vs) correct responses in each condition for each data collection modality (left/right facet) reveals that the pattern of responses appear the same regardless of the data collection modality. In both data collection sessions, the proportion of incorrect responses is much greater than the proportion of correct responses, regardless of condition (marginal probability of incorrect). In the impasse condition, the difference in proportions is smaller than the control condition (conditional probability of success in impasse; (i.e) There are more correct responses in the impasse condition than the control condition).

```{r}
#| label: DESC-Q1ACC

#::::::::::::DESCRIPTIVES

paste("Proportions of Correct Responses by Condition")
table(df$accuracy, df$pretty_condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns

paste("Number of Correct Responses by Condition")
table(df$accuracy, df$pretty_condition) %>% addmargins(2) %>% #display sum for row
      addmargins(1) #sanity check sum of columns
```

#### LOGISTIC REGRESSION

Fit a logistic regression predicting accuracy (absolute score) (n = `r nrow(df)`) by condition (k = 2).\

-   Parameter estimate: $\beta_{0}$ = Log Odds of (correct) responses in CONTROL condition
-   $e^{\beta_{0}}$ = ODDS of correct response in CONTROL condition
-   Parameter estimate: $\beta_{1}$ = $\beta_{1impasse}$ Log Odds (Log OR; change in odds for correct response in impasse (vs) control \[log scale\])
-   $e^{\beta_{1}}$ = ODDS RATIO of correct response in IMPASSE (vs) CONTROL
-   **Null hypothesis**:$\beta_{impasse} \le 0$ the odds for a correct response does not change, or decreases
-   **Alternative hypothesis:** $\beta_{impasse} \gt 0$ the odds of a correct response increases

##### Fit Model

*First, we fit a logistic regression with condition as predictor, and compare its fit to an empty (intercept-only) model.*

```{r}
#| label: MODEL-FIT-Q1ACC
#| warning: false
#| message: false

# MODEL FITTING ::::::::

#: 1 EMPTY MODEL baseline glm model intercept only
m0 = glm(accuracy ~ 1, data = df, family = "binomial")
# print("EMPTY MODEL")
# summary(m0)

#: 2 CONDITION model
m1 <- glm( accuracy ~ pretty_condition, data = df, family = "binomial")
# print("PREDICTOR MODEL")
summary(m1)
car::Anova(m1)

#: 3 TEST SUPERIOR FIT
paste("AIC wth predictor is lower than empty model?", m0$aic > m1$aic)
test_lrt(m0,m1) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(m0,m1))$p[2])
```

*The Condition predictor decreases AIC, but the Likelihood Ratio Test is marginal. We proceed to examine the predictor model, as we plan to do a 1-tailed NHST .*

##### Describe

```{r}
#| label: MODEL-DESC-Q1ACC

#best model 
m <- m1

# DESCRIBE MODEL ::::::::::::::::::::::::::::::::::::: 

print("PREDICTOR MODEL [default two-tailed sig test]")
summary(m)
car::Anova(m)

# MANUAL ONE-SIDED SIGTEST ::::::::::::::::::::::::::: 

# one-sided (right tail) z test for B COEFFICIENT
#https://stats.stackexchange.com/questions/330655/strategy-for-a-one-sided-test-of-glms-coefficients

#SANITY CHECK 2-tailed test should match the model output
##TODO need to do for each coeff
# tt <- 2*pnorm(summary(m)$coefficients[2,3], lower.tail = F)
# paste("p value for two-tailed test, null B = 0 : ",round(tt,3))
# ot <- pnorm(summary(m)$coefficients[2,3], lower.tail = F)
# paste("BUT we want a one tailed directional, null: B <= 0: ",round(ot,3))
# paste("adjusted confint for directional hypothesis")
# (dcint <- confint(m, level = 0.90)) # get 90% for right side))
# https://stats.stackexchange.com/questions/20734/is-a-1-sided-90-prediction-interval-equivalent-to-a-2-sided-95-prediction-inte

#:::::::: INTERPRET COEFFICIENTS

# print("Confidence Interval —- LOG ODDS")
# confint(m1) #not adjusted for 1-tailed
# print("Coefficients —- ODDS RATIOS")
# (e <- cbind( exp(coef(m1)), exp(confint(m1)))) #exponentiated, not adjusted
# (e <- cbind( exp(coef(m1)), exp(dcint))) #exponentiated, adjusted

paste("LOG ODDS")
# (l <- describe_posterior(m))
(tm <- tidy(m,   conf.int = TRUE))

paste("ODDS RATIOS")
(e <- model_parameters(m, exponentiate = TRUE))

print("MODEL PREDICTIONS")
# Retrieve predictions as probabilities 
# (for each level of the predictor)
pred.control <- predict(m1,data.frame(pretty_condition="control"),type="response")
#this should match : plogis(intercept coefficient)
paste("Probability of success in control,", pred.control)
pred.impasse <- predict(m1,data.frame(pretty_condition="impasse"),type="response")
#this should match : plogis(intercept coefficient + predictor coeff)
paste("Probability of success in impasse,", pred.impasse)
```
##### Inference


We fit a logistic regression model to explore the effect of experimental condition on probability of a correct answer on the first question. In this model, the effect of condition is statistically significant (z = 1.79, p = 0.04, one-tailed). The model predicts that the odds of a correct response on the first question in the impasse condition increase by nearly 120% ($e^{beta_{impasse}}$ = 2.19, 95% CI \[1.08, 4.63\]) over the control condition. The intercept** $\beta_{0}$ **parameter is also significant, ($e^{b_{0}}$ = 0.19, p \< 0.001, 95% CI \[0.11, 0.33\]) indicating that the odds of a correct response in the control condition are significantly less than even (less than 50/50 chance of correct response in control condition).

*Equivalent statements:*

-   being in impasse condition increases log odds of correct response by 0.79 (over control)
-   being in impasse increases odds of correct response in impasse over control increases by factor of 2.19
-   probability of correct response in impasse predicted as 30%, vs only 16% in control condition


##### Visualize

```{r}
#| label: MODEL-VIS-Q1ACC
#| message: false
#| warning : false

#SET MODEL
m <- m1

#GGSTATS | MODEL | LOG ODDS 
# ggcoefstats(m1, output = "plot", 
#               conf.level = 0.90) + 
#   labs(x = "Log Odds Estimate", 
#        title = "LOGODDS | Q1 Accuracy ~ condition",
#        subtitle = "(p is for two tailed test)")


#PARAMETERS | MODEL | SIMULATED PARAMETERS
# similar to bayesian dist of estimate
# result <- simulate_parameters(m1)
# #rename params so intercept is plotted 
# result$Parameter[1] <- "condition [control]"
# result$Parameter[2] <- "condition [impasse]"
# plot(result) 

#EQUIVALENCE TEST [not sure if appropriate for log model?]
# https://journals.sagepub.com/doi/10.1177/2515245918770963#:~:text=Consequently%2C%20when%20reporting%20an%20equivalence,values%20is%20smaller%20than%20alpha.
# https://easystats.github.io/parameters/reference/equivalence_test.lm.html
# (result <- equivalence_test(m1, rule = "classic", component = c("all")))
# plot(result,   show_intercept = TRUE) + 
#   scale_y_discrete(labels = c("impasse", "control")) + 
#   labs( title = "Equivalence Test for Model Parameter Estimates")


#PARAMETERS | MODEL | ODDS RATIO 
# result <- model_parameters(m1,exponentiate = TRUE)
# #rename params 
# result$Parameter[1] <- "condition [control]"
# result$Parameter[2] <- "condition [impasse]"
# plot(result,   show_intercept = TRUE) +  labs(
#   title = "Model Parameter Estimates"
# ) + theme_clean() + theme(legend.position="blank")

#SJPLOT | MODEL | ODDS RATIO
#library(sjPlot)
plot_model(m, type = "est",
           vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.1, #manually adjust to account for directional test
           ci.lvl = 0.90 ) +  #manually adjusted for directional test   
  scale_y_continuous() + #remove to put on log scale x axis 
  scale_x_discrete(labels=c("control","impasse"))+
  labs(title = "MODEL ESTIMATE | Q1 Accuracy ~ condition",
       subtitle = "Impasse increases odds of correct response on Q1",
       x = "Condition") + theme_clean()

  
#SJPLOT | MODEL | PROBABILITIES
plot_model(m, type="pred")[[1]] +
  ylim(0,1) + #scale y axis to actual range
  labs(title = "MODEL PREDICTION  | Q1 Accuracy ~ condition",
       subtitle = "Impasse increases probability of correct response on Q1",
       x = "Condition") + theme_clean()

```

```{r}
#| label: MODEL-TBL-Q1ACC
#SJPLOT | MODEL | TABLE
tab_model(m)

# #MODEL SUMMARY | save latex table
# models <- list("odds ratios" = m1, "(log odds)" = m1)
# notes = list("* p < 0.05, ** p < 0.01, *** p < 0.001",
#              '$sigma^{2}$ = 3.29" N(subject) = 126 $\tau_{00}$(subject) = 22.22 N(question) = 13 $\tau_{00}$(question) = 0.31'
#                )
# 
# modelsummary(models,
#              exponentiate = c(TRUE, FALSE),
#              shape = term ~ model + statistic,
#              fmt = 2, #two digits w/ trailing zero
#              estimate  = "{estimate} {stars}",
#              statistic = "conf.int",
#              gof_map = c("AIC", "sigma"),
#              gof_omit = 'RMSE|ICC|BIC',
#              coef_rename = c("pretty_conditionimpasse" = "Condition[impasse]"),
#              title = 'Accuracy ~ Condition (Mixed Logistic Regression)', 
#              notes = notes,
#              output = "analysis/SGC3A/models/tables/GLMER_OverallAccuracy_lab.tex")
# #              # coef_omit = "Intercept",

```

##### Diagnostics

```{r}
#| label: MODEL-DIAG-Q1ACC
#| message: false
#| warning: false

# print("SANITY CHECK REPORTING")
# report::report(m1)

#print("MODEL PERFORMANCE")
# performance(m1)

print("MODEL DIAGNOSTICS")
check_model(m1)
```


## H1B \| Q1 INTERPRETATION STATE

**Do Ss in the IMPASSE condition offer less-orthogonal interpretations on first question?**

While absolute accuracy score tells us whether a participant successfully interpreted the coordinate system, it doesn't allow us to differentiate between different kinds of incorrect interpretations. Here we examine the (categorical) interpretation state type based on the nature of subject's response, and determine if these interpretations differ by experimental condition. State is a 3-category derived response variable that groups the following interpretations:

-   **"orthogonal"** \[reference category\] includes orthogonal and satisficing responses ==\> indicates a primarily *orthogonal* state of coordinate system understanding

-   **"other"** includes: blank, reference point, responses that can't be classified (including selecting all datapoints), =\> indicates an uncertain or unidentifiable state of coordinate system understanding, but one that is distinctly *not* orthogonal nor triangular

-   **"angular"** includes 'lines connecting' responses as well as responses that include both orthogonal *and* triangular answers =\> indicates some degree of angular/triangular coordinate understanding

-   **"triangular"** includes correct triangular and 'lines connecting' responses as well as responses that include both orthogonal *and* triangular answers =\> indicates some degree of angular/triangular coordinate understanding

+-----------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Research Question     | Does Ss in the impasse condition produce less orthogonal responses on the first question?                                                                                                                                                                                                                                                                 |
+=======================+===========================================================================================================================================================================================================================================================================================================================================================+
| **Hypothesis**        | H1A \| Ss in the IMPASSE condition will have a higher likelihood of producing unknown and and triangle-like response states, relative to orthogonal response states, on the first question                                                                                                                                                                |
+-----------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Data**              | -   data: `df_items` where `q == 1`                                                                                                                                                                                                                                                                                                                       |
|                       | -   outcome: `state` ( 4 level factor from 5 level high_interpretation )                                                                                                                                                                                                                                                                                  |
|                       | -   predictor: `condition` \[between-subjects factor\]                                                                                                                                                                                                                                                                                                    |
+-----------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Analysis Strategy** | 1.  Multinomial (Logistic) Regression on state predicted by condition                                                                                                                                                                                                                                                                                     |
|                       |                                                                                                                                                                                                                                                                                                                                                           |
|                       | Alternative:                                                                                                                                                                                                                                                                                                                                              |
|                       |                                                                                                                                                                                                                                                                                                                                                           |
|                       | -   Ordinal regression on state; but model doesn't satisfy proportional odds assumption (parallel slopes)                                                                                                                                                                                                                                                 |
|                       | -   Multinomial or Ordinal regression on high_interpretation (5 category interpretation state which distinguishes between uncertain (blank, reference) unclassifiable, triangle-like and true triangular.) There are some cells with zeros, however (no uncertain responses in control) which means the model can't accurately estimate those comparisons |
+-----------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

#### Setup 
```{r}
#| label: SETUP-Q1STATE

df <- df_items %>% filter(q==1) %>% dplyr::select(pretty_condition, state)

```


#### Describe 
```{r}
#| label: VIS-Q1STATE

#:::::::: STACKED BAR CHART
df %>% 
  ggplot(data = .,
         mapping = aes(x = pretty_condition,
                       fill = state)) +
  geom_bar(position = "fill" ) + #,color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  # facet_wrap(~pretty_mode) + 
   labs(#y = "Correct Response on Q 1",
       title = "Q1 Interpretation",
       x = "Condition",
       fill = "",
       subtitle="")

```

A proportional bar chart visualizing the proportion of each interpretations type by condition for each data collection modality (left/right facet) reveals that the pattern of responses are similar regardless of the data collection modality, by differ by condition. In the impasse condition, there are more triangular responses than in control. We see that around half of the 'incorrect' (i.e. not triangular) responses in the impasse condition are not orthogonal-like, but "other/unknown".


```{r}
#| label: DESC-Q1STATE

#::::::::::::DESCRIPTIVES

table(df$state, df$pretty_condition) %>% addmargins(2) %>% #display sum for row
      prop.table(margin=2) %>%  #return proportion (of column)
      addmargins(1) #sanity check sum of columns


(t <- table(df$state, df$pretty_condition) %>% addmargins(2) %>% #display sum for row
      addmargins(1)) #sanity check sum of columns

```

#### MULTINOMIAL REGRESSION

*Does condition affect the response state of Q1?*

*Fit a logistic regression predicting interpretation state (k=3) by condition(k = 2).*

-   3 equations will be estimated (# categories - 1); each representing the odds of answering in that particular interpretation (vs) the reference category (orthogonal answer) \[essentially a series of binary logistic regressions, but instead of incorrect/correct, comparing \[reference category\] vs \[this category\])

-   For *each* equation:

    -   $\beta_{0}$ *= Log Odds of \[this category type vs. reference category type) response in CONTROL condition*
    -   $e^{\beta_{0}}$ *= ODDS of \[this category type vs. reference category type\] response in CONTROL condition*
    -   $\beta_{1}$ *=* $\beta_{1impasse}$ *Log Odds (Log OR; change in odds for \[this category\] type response in impasse (vs) control \[log scale\])*
    -   $e^{\beta_{1}}$ *= ODDS RATIO of \[this. vs reference category type\] response in IMPASSE (vs) CONTROL*
    -   Two-tailed NHST *Null hypothesis:* $\beta_{impasse} = 0$ *the odds for \[this category of response vs. reference\] are not different for IMPASSE condition*
    -   *Alternative hypothesis:* $\beta_{impasse} \ne 0$ *the odds of \[this category of response vs. reference\] increases or decreases for IMPASSE condition*

##### TODO REDO BRMS Fit Model \[MBLOGIT\]

```{r}
#| label: MODEL-FIT-Q1STATE

#check reference level 
print("Categories (first is reference)")
levels(df$state)

#FIT EMPTY MODEL
# print("EMPTY MODEL")
mbl.0 <- mblogit(state ~ 1, data = df)
# summary(catm.0)

#FIT PREDICTOR MODEL
# print("PREDICTOR MODEL")
mbl.C <- mblogit(formula = state ~ pretty_condition, data = df, model = TRUE)
#IDENTICAL TO MULTINOM() FROM nnet.. but multinom doesn't handle mixed
#so use mblogit for consistency, but need catm for anova test
catm <- multinom(formula = state ~ pretty_condition, data = df, model = TRUE)


#COMPARE MODEL FIT
paste("AIC wth predictor is lower than empty model?", AIC(mbl.0) > AIC(mbl.C))
test_lrt(mbl.0, mbl.C)

##compare bayesian version
#library(brms)
# b.cat <- brm( state2 ~ pretty_condition, data = df, family = "categorical", backend = "cmdstanr")
# summary(b.cat)
# plot_model(b.cat)
# report(b.cat)
# coefficient estimates are very simliar to catm. super cool!

##compare mclogit version
#"baseline-category logit model
# https://www.elff.eu/software/mclogit/manual/mblogit/
# blm1 <- mblogit(state2 ~ pretty_condition , data = df)
# summary(blm1)
#identical to catm. super cool!

```

*AIC in predictor model is less than empty model, and likelihood ratio test indicates predictor model is significantly better fit to the sample data than the empty (intercept only) model.*


##### Describe

```{r}
#| label: MODEL-DESC-Q1STATE

#set model
m <- mbl.C

#confirm models match 
compare_models(mbl.C, catm)

#::::::::: PRINT MODEL 
paste("MODEL SUMMARY")
summary(m)

paste("SIGNIFICANCE TEST")
# car::Anova(mbl.C)
car::Anova(catm)

#:::::::: INTERPRET COEFFICIENTS

paste("LOG ODDS")
# (l <- describe_posterior(m))
(tm <- tidy(m,   conf.int = TRUE))

paste("ODDS RATIOS")
(e <- model_parameters(m, exponentiate = TRUE))

paste("PROBABILITIES")
newdata <- df %>% dplyr::select(pretty_condition)
preds <- predict(catm, newdata = newdata, type = "probs")
preds <- cbind(newdata, preds)
#LENGTHEN TO HANDLE MULTINOMIAL
preds <- preds %>%
  # dplyr::select(-subject, -q) %>% #marginalize over subject and q
  pivot_longer(
  cols = !pretty_condition,
  values_to = "preds",
  names_to = "state",
)

(p <- preds %>%
  group_by(pretty_condition, state ) %>%
  summarise(
    median = median(preds),
    se = sd(preds)/sqrt(n()),
    lwr = median - 1.96*se,
    upr = median + 1.96*se))


```

##### Inference TODO 

-   Being in the IMPASSE condition *increases* the odds of giving an unknown (potentially nonsense) or blank/uncertain response rather than an orthogonal (or satisficing) response by a factor of 45 (z = 3.81, p \< 0.001) . **Participants in the impasse condition were 45x as likely to give an unknown/uncertain response rather than an orthogonal response compared to participants in control**.

-   Being in the IMPASSE condition *increases* the odds of giving 'triangle-like' response rather than an orthogonal (or satisficing) response by a factor of 17.5 (z = 2.60, p \< 0.001 ). **Participants in the impasse condition were almost 6x as likely to give an a triangular response rather than an orthogonal response compared to participants in control**.


-   Being in the IMPASSE condition *increases* the odds of giving 'triangle-like' response rather than an orthogonal (or satisficing) response by a factor of 4.8 (z = 3.30, p \< 0.001 ). **Participants in the impasse condition were almost 6x as likely to give an a triangular response rather than an orthogonal response compared to participants in control**.


-   As with the (binary) logistic regression on accuracy \~ condition, significant model intercepts indicate that the odds of being in any particular response state (vs) orthogonal are significantly different than 1 *in the control condition.* (i.e. not 1:1 odds or 50% chance of being in that response state. Orthogonal is a much more probable response state in control)

-   \[need to to double check interpretation, but I *think* that the OR intercepts converted to probabilities equate to the marginal probability of being in each state in the control condition. which makes sense. I think.?\]

-   IF I change reference category for condition... then the intercepts should no longer be significant. The b1 coefficients should still be significant, but with changed sign (much less likely) \[Yup! this works!\]

##### Visualize

```{r}
#| label: MODEL-VIS-Q1STATE

## | PLOT PARAMETERS 

#SJPLOT | MODEL | ODDS RATIO
plot_model(m, vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.1, #manually adjust to account for directional test
           ci.lvl = 0.90 ) + #manually adjusted for directional test   
  labs(title = "Model Estimate | Odds Ratio",
       subtitle = "",
       x = "Condition")

#EASYSTATS | MODEL | ODDS RATIO
# result <- model_parameters(m, exponentiate = TRUE, component = "all")
# plot(result)


## | PLOT TESTS

result <- equivalence_test(m, rule = "classic", ci=0.9) #classic[tost], , bayes
plot(result)

## | PLOT PREDICTIONS
#TODO NEED FIX THIS
# (d <- p %>%
#   ggplot(aes(y = median,  x = state, fill = pretty_condition)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = lwr, ymax = upr))
#     # stat_slab(width = c(.95), alpha = 0.5, normalize="xy")
#   #   #normalize = all, panels, xy, groups, none
#   # xlim(0,1) + labs(
#   #   title = "Model Predicted Probability of Correct Response",
#   #   x = "probability of correct response",
#   #   y = "Interpretation"
#   # ) +  theme_clean() #+ ggeasy::easy_remove_legend() +
# )
```

```{r}
#| label: MODEL-TBL-Q1STATE

#SJPLOT | MODEL | TABLE
tab_model(m)

# #MODEL SUMMARY | save latex table
# models <- list("odds ratios" = m1, "(log odds)" = m1)
# notes = list("* p < 0.05, ** p < 0.01, *** p < 0.001",
#              '$sigma^{2}$ = 3.29" N(subject) = 126 $\tau_{00}$(subject) = 22.22 N(question) = 13 $\tau_{00}$(question) = 0.31'
#                )
# 
# modelsummary(models,
#              exponentiate = c(TRUE, FALSE),
#              shape = term ~ model + statistic,
#              fmt = 2, #two digits w/ trailing zero
#              estimate  = "{estimate} {stars}",
#              statistic = "conf.int",
#              gof_map = c("AIC", "sigma"),
#              gof_omit = 'RMSE|ICC|BIC',
#              coef_rename = c("pretty_conditionimpasse" = "Condition[impasse]"),
#              title = 'Accuracy ~ Condition (Mixed Logistic Regression)', 
#              notes = notes,
#              output = "analysis/SGC3A/models/tables/GLMER_OverallAccuracy_lab.tex")
# #              # coef_omit = "Intercept",

# modelsummary(mixcat.1, s)
#TODO OUTPUT TABLE 
#https://arelbundock.com/posts/modelsummary_multinomial_logit/


```

##### Diagnostics

```{r}
#| label: MODEL-DIAG-Q1STATE

#EXAMINE PREDICTIONS
# #create sample data frame
# test <- data.frame(pretty_condition = c("control", "impasse"))
# pred <- predict(catm, newdata = test, "probs")
# paste("Predicted Probability of Being in Each State")
# ( x <- cbind(test, pred))

print("MODEL PERFORMANCE")
performance(catm)
DescTools::PseudoR2(catm, which = c("McFadden", "CoxSnell", "Nagelkerke"))


```

