---
subtitle: 'Study SGC3A | 5 Exploratory Analyses'
---

\newpage

# Exploratory Analyses {#sec-SGC3A-exploration}

**TODO**\
- response consistency - clarify core questions being asked\
- review models already created in ARCHIVE?\
- explore response consistency - fix references
- what predicts consistency?
- consider zero inflated / hurdles with ospan score as preditor

*The purpose of this notebook is exploratory analyses of data collected for study SGC3A.*


```{r}
#| label: SETUP
#| warning : false
#| message : false

#misc utilities
library(Hmisc) # %nin% operator
library(broom)
library(modelr)
library(distributional)
library(jtools)
library(pwr) #power analysis
library(mosaic) #favstats

#visualization
library(ggpubr) #arrange plots
library(cowplot) #arrange shift function plots
library(ggformula) #easy graphs
library(vcd) #mosaic plots
library(vcdExtra) #mosaic plots
library(kableExtra) #printing tables 
library(sjPlot) #visualize model coefficients
library(ggdist) #uncertainty viz 
library(ggstatsplot) #plots with stats

#models and performance
library(report) #easystats reporting
library(see) #easystats visualization
library(performance) #easystats model diagnostics
library(qqplotr) #confint on qq plot
library(gmodels) #contingency table and CHISQR
library(equatiomatic) #extract model equation
library(pscl) #zeroinfl / hurdle models 
library(lme4) #mixed effects models 
library(lmerTest) #for CIs in glmer 
library(ggeffects) #visualization log regr models
library(nnet) #multinomial logistic regression [not mixed]
library(mclogit) #frequentist mixed multinomial logistic regression [mblogit]
library(brms) #bayesian mixed multinomials [+ other bayesian reg models]

library(tidyverse) #ALL THE THINGS

#OUTPUT OPTIONS
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
options(ggplot2.summarise.inform = FALSE)
options(scipen=1, digits=3)

#GRAPH THEMEING
theme_set(theme_minimal()) 
```

```{r}
#| label: IMPORT-DATA
#| warning : false
#| message : false

#IMPORT DATA 
df_items <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds')
df_subjects <- read_rds('analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds')
#weird... doesn't respect project level execute dir in yml unless included in index; otherwise breaks render

```

Exploratory Questions

Consistency \| How consistent are learners in their interpretation of the graph? Do they adopt an interpretation on the first question and hold constant? Or do they change interpretations from question to question? Are there any interpretations that serve as 'absorbing states' (i.e. once encountered, the learner does not exist this state).

Time Course of Exploration \| What is the relationship between response accuracy (and interpretation) and time spent on each item?

Can exploration strategies be derived from mouse cursor activity?

-   does response time predict interpretation vs. non interpretation?

TODO: - does impasse yield different exploration behavior? (characterize mouse) - does impasse yield more time on task? (characterize response time ? number of answers then de-selected?)

TODO: Think about characterizing how variable the interpretations are across a participant. Do they form an interpretation and hold it constant? Or do they change question to question.

## WORKING MEMORY 

During the 2021-2022 online replication study, participants completed a working memory task —the OSPAN task- after completing the SGC3A study.  We added this task to explore possible sources of variance between individuals on the graph comprehension task.  

**Here we address the question: Does working memory (as measured by the OSPAN task) help explain performance on the interval graph comprehension task?**

_First we join the load data for the subset of SGC3A participants who completed the OSPAN task. Note that this is a slightly lower n than the online replication, as a couple of participants failed to finish the OSPAN task, and will be excluded from this analysis._

```{r}

#LOAD OSPAN PARTICIPANTS
df_ospan <- read_csv("analysis/SGC3A/data/1-study-level/sgc3a_ospan.csv") #not scored
sgc3a <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_participants.rds") #not scored
sgc3a_items <- read_rds("analysis/SGC3A/data/2-scored-data/sgc3a_scored_items.rds") #not scored

##ADD OSPAN DATA TO SUBJECT LEVEL DATA

#select needed ospan columns
df_ospan <- df_ospan %>% rename( OSPAN.weighted = weighted, OSPAN.math_acc = math_acc,
                                 OSPAN.order_acc = order_acc) %>% dplyr::select(
                                 subject, OSPAN.weighted, OSPAN.math_acc, OSPAN.order_acc) %>% 
  mutate(
      z_ospan = zscore(OSPAN.weighted),
      ospan_split = as.factor(OSPAN.weighted > median(OSPAN.weighted)),
      ospan_split = recode_factor(ospan_split, "FALSE" = "low-memory", "TRUE" = "high-memory")
    
  )

#get only the participants for whom we have ospan data 
df_sgc3a <- sgc3a %>% filter(subject %in% df_ospan$subject)

#add q1 state field [3 category summary of interepretation]
df_sgc3a <- df_sgc3a %>% mutate(
  item_q1_state =  recode_factor(item_q1_SCALED, 
                         "-1" = "orth-like",
                         "-0.5" = "unknown",
                         "0" = "unknown",
                         "0.5" = "tri-like",
                         "1" = "tri-like"),
)

#join the ospan columns to the scored sgc3a data 
df_sgc3a <- merge(df_sgc3a, df_ospan)


##ADD OSPAN DATA TO ITEM LEVEL DATA
#get only the participants for whom we have ospan data 
df_sgc3a_items <- sgc3a_items %>% filter(subject %in% df_ospan$subject) 
df_sgc3a_items <- merge(df_sgc3a_items, df_ospan)


#define dataset
df_s <- df_sgc3a
df_i <- df_sgc3a_items
```

_What is the distribution of performance on the OSPAN task?_


```{r}
#| label: DESC-SUBJ-OSPAN
 
title = "Descriptive Statistics of OSPAN Task Accuracy"
ospan.stats <- rbind(
  "MATH" = df_s %>% dplyr::select(OSPAN.math_acc) %>% unlist() %>% favstats(),
  "ORDER" = df_s %>%  dplyr::select(OSPAN.order_acc) %>% unlist() %>% favstats(),
  "WEIGHTED" = df_s %>% dplyr::select(OSPAN.weighted) %>% unlist() %>% favstats()

)
ospan.stats %>% kbl (caption = title) %>% kable_classic() %>%
  footnote(general = "MATH = %correct of all math questions;
           ORDER = % correct of OSPAN ordering (out of 30); WEIGHTED = math*ospan number correct", general_title = "Note: ",footnote_as_chunk = T)

```


```{r}
#| label: VIS-SUBJ-OSPAN

# #GGFORMULA | DENSITY HISTOGRAM 
  gf_dhistogram(~z_ospan, data = df_s) +
  labs(x = "OSPAN (weighted) score",
       y = "% of subjects",
       title = "Distribution of OSPAN SCORE",
       subtitle = "")
 
##GGPUBR | HIST+DENSITY SCORE BY CONDITION/MODE
p <- gghistogram(df_s, x = "z_ospan", binwidth = 0.5,
   add = "mean", rug = TRUE,
   fill = "pretty_condition", #, palette = c("#00AFBB", "#E7B800"),
   add_density = TRUE)
facet(p, facet.by=c("pretty_condition")) +
  labs( title = "Distribution of OSPAN Score",
        subtitle ="The shape of the distribution of OSPAN scores is similar across conditions",
        x = "OSPAN (weighted) score", y = "number of subjects") +
  theme_minimal() + theme(legend.position = "blank")


#GGSTATS
ggbetweenstats(y = z_ospan, x = pretty_condition, data = df_s,
               type = "parametric", var.equal = FALSE,
               title = "Independent Samples T-Test indicates little difference between means by Condition")


#PROPORTIONAL BAR CHART OF MED SPLIT BY CONDITION
ggbarstats( x = pretty_condition, y = ospan_split, data = df_s) + labs(
  title = "The Proportion of high (vs) low WM scores by does appear to differ by condition"
)



```
_As we would expect, there is not a meaningful difference between the means (or shape of distribution) of the outcome weighted OSPAN by condition, nor between the proportions of high (vs) low scores after taking a median split of the weighted OSPAN score._



### OSPAN ~ Q1 Accuracy

**Does the OSPAN score help explain variance in Q1 Accuracy?**

```{r}

#SETUP DATA
df_s = df_s #just for reference


##STATSPLOT
ggstatsplot::grouped_ggbarstats(data = df_s, 
                                    y = ospan_split, x = item_q1_NABS, 
                                    grouping.var = pretty_condition,
                                    type = "nonparametric", equal.var = FALSE,
                                    annotation.args = list(
    title = "Potential INTERACTION between condition and WM Score"))
```

_It appears as though there may be an interaction between WM score and CONDITION, such that in the control condition, participants perform equally well (or poorly, as the case may be), regardless of of WM score. In the impasse condition, readers with the impasse condition have a higher proportion of correct responses than readers with low WM scores. One possible explanation of these observations is that perhaps higher WM capacity enables readers to take advantage of the implicit scaffolding the impasse provides._

#### Statistical Tests


To explore the relationship between ACCURACY, CONDITION and OSPAN score, we can run two separate CHI square tests evaluating independence of condition and accuracy, one for each level of working memory (the groups are roughly equal due to the median split). If WM interacts with condition, we should expect to see one chi square indicate independence (low working memory) and one not (high working memory).  Additionally, we can compute a Breslow-Day test for Homogeneity of Odds Ratios, to test the null hypothesis that the odds ratios for accuracy by condition do not differ across levels of working memory. 
```{r}

#IS ACCURACY INDEPENDENT OF CONDITION for low WM?
paste("Accuracy independent of CONDITION for LOW WM?")
x <- df_s %>% filter(ospan_split == "low-memory")
(t <- table(x$item_q1_NABS, x$condition))
# chisq.test(t)
chiSquare(item_q1_NABS ~ condition, data = x)
# YES, IT IS (EXPECTED)

#IS ACCURACY INDEPENDENT OF CONDITION for low WM?
paste("Accuracy independent of CONDITION for HIGH WM?")
x <- df_s %>% filter(ospan_split == "high-memory")
(t <- table(x$item_q1_NABS, x$condition))
# chisq.test(t)
chiSquare(item_q1_NABS ~ condition, data = x)
# YES, IT IS (EXPECTED)

library(DescTools)
#Calculates the Breslow-Day test of homogeneity for a 2 \times 2 \times k2×2×k table, in order to investigate if all #kk strata have the same OR. If OR is not given, the Mantel-Haenszel estimate is used.
t <- table(df_s$item_q1_NABS, df_s$ospan_split, df_s$condition)
BreslowDayTest(t, OR = 1)


```

_A Brewslow-Day Test of homogeneity of Odds Ratios allows us to test the null hypothesis that the odds for each response outcome (correct, incorrect) are the same for each condition across each level of working memory. Essentially, that the odds ratio of a positive response for each condition does not differ by working memory.  The significant p-value for this test allows us to reject this null hypothesis and conclude we have evidence that working memory score changes the odds ratio of a correct vs incorrect response across conditions. 

#### Logistic Regression

**Logistic Regression** Q1 Accuracy ~ Condition + OSPAN

##### Fit Model

```{r}

#DATASETUP
df <- df_i %>% filter(q==1) %>% mutate(
  accuracy = as.factor(score_niceABS),
  l_condition = fct_relevel(condition, "121"),
  l_split = fct_relevel(ospan_split, "high-memory")
)


#:::::::: FIT MODELS

#fit empty model
m.0 <- glm(  accuracy ~ 1, data = df, family = "binomial")
# summary(m.0)

#fit condition as predictor 
m.c <- glm( accuracy ~ pretty_condition, data = df, family = "binomial")
#summary(m.c)

#is condition better than empty model?
print("Is Condition model better than empty model?")
print("Expect YES")
test_lrt(m.0, m.c)

#fit OSPAN as predictor 
m.o <- glm( accuracy ~ ospan_split, data = df, family = "binomial")
#summary(m.o)

#is ospan better than empty model?
print("Is OSPAN model better than empty model?")
print("Expect NO")
test_lrt(m.0, m.o)


# fit condition AND OSPAN MAIN EFFECTS
m.co <- glm( accuracy ~ pretty_condition + ospan_split, data = df, family = "binomial")
summary(m.co)
# car::Anova(m.cw)

#is CONDITION + OSPAN better than CONDITION model?
print("Is CONDITION + OSPAN model better than CONDITION model?")
print("Expect NO")
test_lrt(m.c, m.co)

# fit condition AND OSPAN with INTERACTION 
m.cio <- glm( accuracy ~  pretty_condition:ospan_split + pretty_condition + ospan_split , data = df, family = "binomial")
# summary(m.cio)
# car::Anova(m.ciw)

#is CONDITION + OSPAN IXN better than CONDITION model?
print("Is CONDITION + OSPAN IXN model better than CONDITION model?")
print("Expect YES")
test_lrt(m.cio, m.c)

# MANUAL ONE-SIDED SIGTEST ::::::::::::::::::::::::::: 

# one-sided (right tail) z test for B COEFFICIENT
#https://stats.stackexchange.com/questions/330655/strategy-for-a-one-sided-test-of-glms-coefficients

#SANITY CHECK 2-tailed test should match the model output
# tt <- 2*pnorm(m.3$coefficients, lower.tail = F)
# paste("p value for two-tailed test, null B = 0 : ",round(tt,3))
# ot <- pnorm(m.3$coefficients, lower.tail = F)
# paste("BUT we want a one tailed directional, null: B <= 0: ",round(ot,3))
# paste("adjusted confint for directional hypothesis")
# (dcint <- confint(m1, level = 0.90)) # get 90% for right side))
# # https://stats.stackexchange.com/questions/20734/is-a-1-sided-90-prediction-interval-equivalent-to-a-2-sided-95-prediction-inte

```

##### Describe

```{r}

summary(m.cio)
car::Anova(m.cio)
```
##### Inference

Significant main effect condition, non significant main effect of OSPAN, non significant interaction.

Although we can see (ref plots below) that high_working memory participants were more likely to correctly answer Q1, the difference in proportions for this sample size (~30 per cell in the 2x2 contingency table of condition X OSPAN score) failed to reach statistical significance at the 0.05 alpha level.

##### Visualize
```{r}
plot_model(m.c, type = "eff")
plot_model(m.cio, type = "int")
```

##### Diagnostics
```{r}

#PERFORMANCE
performance(m.cio)

#DIAGNOSTICS
check_model(m.cio)
```


### OSPAN ~ Q1 Interpretation

**Does the OSPAN score help explain variance in Q1 interpetation?**

```{r}
##STATSPLOT
ggstatsplot::grouped_ggbarstats(data = df_s, 
                                    y = ospan_split, x = item_q1_state, 
                                    grouping.var = pretty_condition,
                                    type = "nonparametric", equal.var = FALSE,
                                    annotation.args = list(
    title = "Potential INTERACTION between condition and WM Score"))
```
_Again we see a potential interaction between WM score and CONDITION looking at the interpretation indicated by responses on the first question. In the control condition, low and high working memory participants offered very similar proportions of each type of interpretation (with the high working memory subjects offering slightly more unknown and tri-like interpretations). Alternatively in the impasse condition, high working memory Ss have substantially fewer orthogonal interpretations, with far more triangular and unknown interpretations._

#### Statistical Tests


To explore the relationship between INTERPREATION STATE, CONDITION and OSPAN score, we can run two separate CHI square tests evaluating independence of condition and interpretation, one for each level of working memory (the groups are roughly equal due to the median split). If WM interacts with condition, we should expect to see one chi square indicate independence (low working memory) and one not (high working memory).  Additionally, we can compute a Breslow-Day test for Homogeneity of Odds Ratios, to test the null hypothesis that the odds ratios for accuracy by condition do not differ across levels of working memory. 
```{r}

#IS INTERPRETATION STATE INDEPENDENT OF CONDITION for low WM?
paste("Accuracy independent of CONDITION for LOW WM?")
x <- df_s %>% filter(ospan_split == "low-memory")
(t <- table(x$item_q1_state, x$condition))
# chisq.test(t)
chiSquare(item_q1_state ~ condition, data = x)
# NOT expected

#IS ACCURACY INDEPENDENT OF CONDITION for HIGH WM?
paste("Accuracy independent of CONDITION for HIGH WM?")
x <- df_s %>% filter(ospan_split == "high-memory")
(t <- table(x$item_q1_state, x$condition))
# chisq.test(t)
chiSquare(item_q1_state ~ condition, data = x)
# YES, IT IS (EXPECTED)

library(DescTools)
#Calculates the Breslow-Day test of homogeneity for a 2 \times 2 \times k2×2×k table, in order to investigate if all #kk strata have the same OR. If OR is not given, the Mantel-Haenszel estimate is used.
t <- table(df_s$item_q1_state, df_s$ospan_split, df_s$condition)
BreslowDayTest(t, OR = 1)


```

A Brewslow-Day Test of homogeneity of Odds Ratios allows us to test the null hypothesis that the odds for each response outcome (orth-like, unknown, tri-like) are the same for each condition across each level of working memory. Essentially, that the odds ratio of a positive response for each condition does not differ by working memory.  

_We must reject the null hypothesis that (for low-working memory participants) condition and interpretation are independent. They do appear to be related, such that even for low-working memory participants, the impasse condition yields many more unknown and triangle-like responses than the control condition.  Together with the non-significant p-value for that does not allow us to reject the null hypothesis of homogeneous odds ratios across levels of working memory score, we do not have evidence of an interaction between working memory score and INTERPRETATION on question 1. 

#### Logistic Regression

**Logistic Regression** Q1 Accuracy ~ Condition + OSPAN

##### Fit Model

```{r}

#DATASETUP
df <- df_i %>% filter(q==1) %>% mutate(
  accuracy = as.factor(score_niceABS),
  l_condition = fct_relevel(condition, "121"),
  l_split = fct_relevel(ospan_split, "high-memory")
)

#:::::::: FIT MODELS

#fit empty model
m.0 <- glm(  accuracy ~ 1, data = df, family = "binomial")
# summary(m.0)

#fit condition as predictor 
m.c <- glm( accuracy ~ pretty_condition, data = df, family = "binomial")
#summary(m.c)

#is condition better than empty model?
print("Is Condition model better than empty model?")
print("Expect YES")
test_lrt(m.0, m.c)

#fit OSPAN as predictor 
m.o <- glm( accuracy ~ ospan_split, data = df, family = "binomial")
#summary(m.o)

#is ospan better than empty model?
print("Is OSPAN model better than empty model?")
print("Expect NO")
test_lrt(m.0, m.o)


# fit condition AND OSPAN MAIN EFFECTS
m.co <- glm( accuracy ~ pretty_condition + ospan_split, data = df, family = "binomial")
summary(m.co)
# car::Anova(m.cw)

#is CONDITION + OSPAN better than CONDITION model?
print("Is CONDITION + OSPAN model better than CONDITION model?")
print("Expect NO")
test_lrt(m.c, m.co)

# fit condition AND OSPAN with INTERACTION 
m.cio <- glm( accuracy ~  pretty_condition:ospan_split + pretty_condition + ospan_split , data = df, family = "binomial")
# summary(m.cio)
# car::Anova(m.ciw)

#is CONDITION + OSPAN IXN better than CONDITION model?
print("Is CONDITION + OSPAN IXN model better than CONDITION model?")
print("Expect YES")
test_lrt(m.cio, m.c)

# MANUAL ONE-SIDED SIGTEST ::::::::::::::::::::::::::: 

# one-sided (right tail) z test for B COEFFICIENT
#https://stats.stackexchange.com/questions/330655/strategy-for-a-one-sided-test-of-glms-coefficients

#SANITY CHECK 2-tailed test should match the model output
# tt <- 2*pnorm(m.3$coefficients, lower.tail = F)
# paste("p value for two-tailed test, null B = 0 : ",round(tt,3))
# ot <- pnorm(m.3$coefficients, lower.tail = F)
# paste("BUT we want a one tailed directional, null: B <= 0: ",round(ot,3))
# paste("adjusted confint for directional hypothesis")
# (dcint <- confint(m1, level = 0.90)) # get 90% for right side))
# # https://stats.stackexchange.com/questions/20734/is-a-1-sided-90-prediction-interval-equivalent-to-a-2-sided-95-prediction-inte

```

##### Describe

```{r}
summary(m.cio)
car::Anova(m.cio)
```
##### Inference

Main effect of condition, no main effect of OSPAN and non significant interaction. 

Although we can see (ref plots below) that high_working memory participants were more likely to correctly answer Q1, the difference in proportions for this sample size (~30 per cell in the 2x2 contingency table of condition X OSPAN score) failed to reach statistical significance at the 0.05 alpha level.

##### Visualize
```{r}
plot_model(m.c, type = "eff")
plot_model(m.cio, type = "int")
```

##### Diagnostics
```{r}

#PERFORMANCE
performance(m.cio)

#DIAGNOSTICS
check_model(m.cio)
```





### OSPAN ~ Test Phase Accuracy 

**Does the OSPAN score help explain variance in test phase accuracy?**

```{r}

##SETUP DATA
df <- df_s

##STATSPLOT
ggstatsplot::grouped_ggbetweenstats(data = df, 
                                    y = item_test_NABS, x = ospan_split, 
                                    grouping.var = pretty_condition,
                                    type = "nonparametric", equal.var = FALSE,
                                    annotation.args = list(
    title = "Potential INTERACTION between condition and WM Score"))
```
_There appears to be an interaction between CONDITION and OSPAN score such that only participants with high working memory scores had better performance in the impasse condition. This could potential be driving the effect of the impasse condition._

#### Linear Regression

##### Fit Model
```{r}

#PREP DATA
df_s <- df_s %>% mutate(
  c_acc = scale(item_test_NABS), #mean centered accuracy score
  med_acc = scale(item_test_NABS, center = median(item_test_NABS), scale = F)
)

#:::::::: FIT MODELS

#empty model
m.0 <- lm( data = df_s, item_test_NABS ~ 1)
# summ(m.0)

#condition predictor
m.c <- lm( data = df_s, item_test_NABS ~ pretty_condition)
#summ(m.c)

paste("Condition predictor improves fit?")
test_lrt(m.c, m.0)

#add ospan predictor
m.co <- lm( data = df_s, item_test_NABS ~ pretty_condition + ospan_split)
summary(m.co)

paste("Condition + OSPAN improves fit?")
test_lrt(m.c, m.co)

#add interaction term  
m.cio <- lm( data = df_s, item_test_NABS ~ pretty_condition + ospan_split + pretty_condition:ospan_split)
summary(m.cio)

paste("Condition : OSPAN interaction improves fit?")
test_lrt(m.co, m.cio)
```
##### Describe 

```{r}

#Describe 
summary(m.cio)
anova(m.cio)
```

$\beta_0$ intercept is MEAN when both predictors are at reference: (control condition low memory) [M = 1.129]
$\beta_{1impasse}$ intercept is DIFFERENCE to (IMPASSE condition low memory) [M 1.72 - 1.129 = 0.591]
$\beta_{1high}$ intercept is DIFFERENCE to (CONTROL condition HIGH memory) [M 1.21 - 1.129 = 0.079]
$\beta_{1 impasse : high}$ is DIFFERENCE to ?? no idea can't figure this one out. its not to reference levels

BUT don't fret re: significance of predictors in the model, rather, to see if effects are significant, just run anova on the model. 

##### Visualize

**By comparing the average performance by condition ...**
```{r}
#PLOT EFFECTS
plot_model(m.c, type = "pred")
#TODO GET + ylim(0,8) to work
```

**...to the average by condition broken out by working memory, we see that perhaps working memory score drives some of the effect of improved performance in the impasse condition.**

```{r}
plot_model(m.cio, type = "int") 
#+ ylim(0,8)
```
##### Diagnostics
```{r}

#PERFORMANCE
performance(m.cio)

#DIAGNOSTICS
check_model(m.cio)

```



#### Mixed Logistic Regression

```{r}
#PREP DATA
df_i <- df_sgc3a_items %>% filter(q %nin% c(6,9)) %>% mutate(
  accuracy = as.factor(score_niceABS),
  ospan = as.factor(ospan_split)
) #test phase discrim only

```
##### Fit Model

```{r}

## 1 | SETUP RANDOM INTERCEPT SUBJECT

#:: EMPTY MODEL (baseline, no random effect)
print("Empty fixed model")
m0 = glm(accuracy ~ 1, family = "binomial", data = df_i) 
# summary(m0)

#:: RANDOM INTERCEPT SUBJECT
print("Subject intercept random model")
mm.rS <- glmer(accuracy ~ (1|subject), data = df_i, family = "binomial",
               control=glmerControl(optimizer="bobyqa",
            optCtrl=list(maxfun=100000)))
# summary(mm.rS)

# :: TEST random effect
paste("AIC decreases w/ new model?", m0$aic > AIC(logLik(mm.rS)))
test_lrt(m0,mm.rS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(m0,mm.rS))$p[2])


## 2 | ADD RANDOM INTERCEPT ITEM?

#:: RANDOM INTERCEPT SUBJECT + INTERCEPT Q
# print("Subject & Question random intercepts")
# mm.rSQ <- glmer(accuracy ~ (1|subject) + (1|q), data = df_i, family = "binomial")
# # summary(mm.rSQ)
# 
# # :: TEST random effect
# paste("AIC decreases w/ new model?", AIC(logLik(mm.rS)) > AIC(logLik(mm.rSQ)) )
# test_lrt(mm.rS,mm.rSQ) #same as anova(m0, m1, test = "Chi")
# paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rS,mm.rSQ))$p[2])

## 3 | ADD FIXED EFFECT CONDITION

# print("FIXED Condition + Subject & Question random intercepts")
# mm.CrSQ <- glmer(accuracy ~ pretty_condition + (1|subject) + (1|q), 
#                 data = df_i, family = "binomial")
# summary(mm.CrSQ)
# 
# paste("AIC decreases w/ new model", AIC(logLik(mm.rSQ)) > AIC(logLik(mm.CrSQ)) )
# test_lrt(mm.rSQ,mm.CrSQ) #same as anova(m0, m1, test = "Chi")
# paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rSQ,mm.CrSQ))$p[2])

print("FIXED Condition + Subject random intercepts")
mm.CrS <- glmer(accuracy ~ pretty_condition + (1|subject) ,
                data = df_i, family = "binomial")
# summary(mm.CrS)

paste("AIC decreases w/ new model", AIC(logLik(mm.rS)) > AIC(logLik(mm.CrS)) )
test_lrt(mm.rS,mm.CrS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rS,mm.CrS))$p[2])


## 4 | ADD FIXED EFFECT OSPAN only 
print("FIXED OSPAN + Subject random intercepts")
mm.OrS <- glmer(accuracy ~ ospan + (1|subject) ,
                data = df_i, family = "binomial")
# summary(mm.OrS)
paste("OSPAN decreases AIC over random?", AIC(logLik(mm.rS)) > AIC(logLik(mm.OrS)) )
test_lrt(mm.rS,mm.OrS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.rS,mm.OrS))$p[2])

## 5 | MODEL OF INTEREST! OSPAN + CONDITION + suject

print("FIXED OSPAN + CONDITION + IXN + Subject random intercepts")
mm.COrS <- glmer(accuracy ~ pretty_condition + ospan + (1|subject) ,
                data = df_i, family = "binomial")
# summary(mm.COrS)
paste("OSPAN decreases AIC over CONDITION alone?", AIC(logLik(mm.COrS)) > AIC(logLik(mm.CrS)) )
test_lrt(mm.COrS,mm.CrS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.COrS,mm.CrS))$p[2])


print("FIXED OSPAN + CONDITION + IXN + Subject random intercepts")
mm.CIOrS <- glmer(accuracy ~ pretty_condition + ospan + pretty_condition:ospan + (1|subject) ,
                data = df_i, family = "binomial")

paste("IXN decreases AIC over main effects alone?", AIC(logLik(mm.CIOrS)) > AIC(logLik(mm.COrS)) )
test_lrt(mm.CIOrS,mm.COrS) #same as anova(m0, m1, test = "Chi")
paste("Likelihood Ratio test is significant? p = ",(test_lrt(mm.CIOrS,mm.COrS))$p[2])

```

##### Describe

```{r}

# best model
m1 <- mm.CIOrS

#::::::::: PRINT MODEL 

print("PREDICTOR MODEL")
summary(m1)
car::Anova(m1)

#:::::::: MANUAL ONE-SIDED SIGTEST 

# one-sided (right tail) z test for B COEFFICIENT
#SANITY CHECK 2-tailed test should match the model output
# tt <- 2*pnorm(summary(m1)$coefficients[2,3], lower.tail = F)
# paste("p value for two-tailed test, null B = 0 : ",round(tt,5))
# ot <- pnorm(summary(m1)$coefficients[2,3], lower.tail = F)
# paste("BUT we want a one tailed directional, null: B <= 0: ",round(ot,5))

#:::::::: INTERPRET COEFFICIENTS

se <- sqrt(diag(stats::vcov(m1)))
# table of estimates with 95% CI
paste("LOG ODDS")
(tab <- cbind(Est = fixef(m1), LL = fixef(m1) - 1.96 * se, UL = fixef(m1) + 1.96 *
    se))
paste("ODDS RATIOS")
(e <- exp(tab))

```

##### Inference

TODO

##### Visualize

```{r}
#:::::::: PLOT

#GGSTATS | MODEL | LOG ODDS 
ggcoefstats(m1, output = "plot", 
              conf.level = 0.90) + 
  labs(x = "Log Odds Estimate", 
       subtitle = "p is for two tailed test")

#SJPLOT | MODEL | ODDS RATIO
#library(sjPlot)
plot_model(m1, vline.color = "red", 
           show.intercept = TRUE, 
           show.values = TRUE,
           p.threshold = 0.1, #manually adjust to account for directional test
           ci.lvl = 0.90 ) + #manually adjusted for directional test   
  labs(title = "Model Predicted Odds Ratio",
       subtitle = "",
       x = "Condition")

#SJPLOT | MODEL | PROBABILITIES
plot_model(m1, type="int",
           show.intercept = TRUE,
           show.values = TRUE,
           title = "Model Predicted Probability of Accuracy",
           axis.title = c("Condition","Probability of Accurate Response"))

#SJPLOT | MODEL | TABLE
#tab_model(m1)

```

##### Narrative

**By comparing the average performance by condition ...**
```{r}
#PLOT EFFECTS
plot_model(mm.CrS, type = "pred")
#TODO GET + ylim(0,8) to work
```

**...to the average by condition broken out by working memory, we see that perhaps working memory score drives some of the effect of improved performance in the impasse condition.**

```{r}
plot_model(mm.CIOrS, type = "int") 
#+ ylim(0,8)
```

_It is compelling to see that only high_working memory participants achieved the high probabilites of correct response in impasse_.




##### Diagnostics

```{r}
#| warning: false
#| message: false

print("SANITY CHECK REPORTING")
report(m1)

print("MODEL PERFORMANCE")
performance(m1)

print("DIAGNOSTICS")
check_model(m1)
```




### TODO OSPAN ~ Test Phase Interpretation 

**Does the OSPAN score help explain variance in test phase interpretation?**

```{r}
##STATSPLOT
ggstatsplot::grouped_ggbarstats(data = df_i, 
                                    y = ospan_split, x = pretty_condition, 
                                    grouping.var = score_STATE,
                                    type = "nonparametric", equal.var = FALSE,
                                    annotation.args = list(
    title = "Potential INTERACTION between condition and WM Score"))
```








## TODO SEQUENCE
## TODO BEST FIT

## ARCHIVE
### Q1 ACCURACY 

**What explains Q1 Accuracy?**

##### Multiple Logistic Regression [Q1 Absolute]

```{r}
# #CREATE DATAFRAME OF Q1
# df <- df_items %>% filter(q == 1) %>% mutate( accuracy = as.factor(score_niceABS))
# 
# #MODEL
# m2 <- glm( accuracy ~ rt_s, data = df, family = "binomial")
# summary(m2)
# confint(m2)
# performance(m2)
# report(m2)
# 
# library(effects)
# plot(allEffects(m2))
# 
# 
# m3 <- glm( accuracy ~ condition + rt_s, data = df, family = "binomial")
# summary(m3)
# confint(m3)
# performance(m3)
# report(m3)
# 
# plot(allEffects(m3))
# 
# compare_performance(m,m2,m3)
```

```{r}
#evaluate model using kfold CV
# https://www.statology.org/k-fold-cross-validation-in-r/

# #specify the cross-validation method
# ctrl <- trainControl(method = "cv", number = 5)
# #fit a regression model and use k-fold CV to evaluate performance
# mk <- train( accuracy ~ condition, data = df, method = "glm", trControl = ctrl, family = "binomial")
# print(mk)
# 


```



#### Mass Movement

"movement of mass" from one mode to another

Considering only families of unimodal distributions, the most probably distribution (as predicted by package `performance`) is negative-binomial.

```{r}
#| label: MASS-111
# df <- df_subjects %>% filter(condition==111)
# multimode::modetest(df$s_NABS)
# n_modes = multimode::nmodes(data = df$s_NABS, bw=2)
# multimode::locmodes(df$s_NABS,mod0 =  n_modes, display = TRUE)

```

```{r}
#| label: MASS-121
# df <- df_subjects %>% filter(condition==121)
# multimode::modetest(df$s_NABS)
# n_modes = multimode::nmodes(data = df$s_NABS, bw=2)
# multimode::locmodes(df$s_NABS,mod0 =  n_modes, display = TRUE)

```


### RESPONSE LATENCY

-   [TODO: Investigate super high and super low response times.]{style="color: red;"}.
-   [TODO: Investigate appropriate models for response time data. (see: https://lindeloev.github.io/shiny-rt/)]{style="color: red;"}.
-   Especially see https://lindeloev.github.io/shiny-rt/ for ideas on modelling reaction time data

+-----------------------+----------------------------------------+
| Research Question     |                                        |
+=======================+========================================+
| **Hypothesis**        |                                        |
+-----------------------+----------------------------------------+
| **Analysis Strategy** |                                        |
+-----------------------+----------------------------------------+
| **Alternatives**      |                                        |
+-----------------------+----------------------------------------+
| **Inference**         |                                        |
+-----------------------+----------------------------------------+

### Q1 Response Latency

#### Linear Regression (Log Transform)

##### (In Person)

###### Visualization

```{r}
#| label: VIS-TEST-Q1TIME
# 
# df_lab <- df_subjects %>% filter(mode == "lab-synch")
#  
# #HISTOGRAM
# stats = df_lab %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(item_q1_rt))
# gf_dhistogram(~log(item_q1_rt), fill = ~pretty_condition, data = df_lab) %>% gf_facet_grid(~pretty_condition) +
#   # gf_vline(data = stats, xintercept = ~mean, color = "red") +
#   labs(title = "(LAB) First Question Response Time",,
#        # x = "Response Time (seconds)",
#        # y = "proportion of participants",
#        subtitle = "") + 
#   theme_minimal()

```

###### Model

```{r}
#| label: MODEL-Q1TIME-LAB
# 
# #SCORE predicted by CONDITION
# lab.q1t.lm1 <- lm(log(item_q1_rt) ~ pretty_condition, data = df_lab)
# paste("Model")
# summary(lab.q1t.lm1)
# paste("Partition Variance")
# anova(lab.q1t.lm1)
# paste("Confidence Interval on Parameter Estimates")
# confint(lab.q1t.lm1)
# report(lab.q1t.lm1) #sanity check
# #print model equation
# eq <- extract_eq(lab.q1t.lm1, use_coefs = TRUE)
```

```{r}
# #| label: VISMODEL-Q1-LATENCY-LAB
# 
# #MODEL ESTIMATES WITH UNCERTAINTY
# 
# #setup references 
# #lab.q1t.lm1 <- lm(log(item_q1_rt) ~ condition, data = df_lab)
# m <- lab.q1t.lm1
# df <- df_lab 
# call <- m$call %>% as.character()
# 
# # uncertainty model visualization
# df <- df  %>%
#   data_grid(pretty_condition) %>%
#   augment(m, newdata = ., se_fit = TRUE) 
# 
# #transform log
# df$.fitted <- exp(df$.fitted)
# df$.se.fit <- exp(df$.se.fit)
# 
# df %>% 
#   ggplot(aes(y = pretty_condition, color = pretty_condition)) +
#   stat_halfeye( scale = .5,
#       aes(
#         xdist = dist_student_t(df = df.residual(m), mu = .fitted, sigma = .se.fit),
#         fill = stat(cut_cdf_qi(cdf, 
#                 .width = c(.90, .95),
#                 labels = scales::percent_format())))) +
#   scale_fill_brewer(direction = -1) + 
#   labs (title = "(LAB) Q1 Response Latency ~ Condition", 
#         x = "model predicted mean (seconds)", y = "Condition", fill = "Interval",
#         subtitle = paste("lm(",call[2],")"),
#         caption = "note: model log(predictions) have exponentiated to original scale") + theme(legend.position = "blank")

```

###### Diagnostics

```{r}
#| label: DIAG-Q1TIME-lab
# 
# #model diagnostics
# check_model(lab.q1t.lm1, panel = TRUE)
```



###### Inference

OLS Linear Regression on Q1 response time shows that condition explains a small but statistically significant amount of variance (impasse \> control). However, the model violates the assumption of normally distributed residuals.

##### (Online Replication)

###### Visualization

```{r}
#| label: VIS-TEST-Q1TIME-online
# 
# df_online <- df_subjects %>% filter(mode == "asynch")
#  
# #HISTOGRAM
# stats = df_online %>% group_by(pretty_condition) %>% dplyr::summarise(mean = mean(item_q1_rt))
# gf_dhistogram(~log(item_q1_rt), fill = ~pretty_condition, data = df_lab) %>% gf_facet_grid(~pretty_condition) +
#   # gf_vline(data = stats, xintercept = ~mean, color = "red") +
#   labs(title = "(ONLINE) First Question Response Time",
#        # x = "Response Time (seconds)",
#        # y = "proportion of participants",
#        subtitle = "") + 
#   theme_minimal()

```

###### Model

```{r}
#| label: MODEL-Q1TIME-online

# #SCORE predicted by CONDITION
# rep.q1t.lm1 <- lm(log(item_q1_rt) ~ pretty_condition, data = df_online)
# paste("Model")
# summary(rep.q1t.lm1)
# paste("Partition Variance")
# anova(rep.q1t.lm1)
# paste("Confidence Interval on Parameter Estimates")
# confint(rep.q1t.lm1)
# report(rep.q1t.lm1) #sanity check
# #print model equation
# eq <- extract_eq(rep.q1t.lm1, use_coefs = TRUE)
```

```{r}
#| label: VISMODEL-Q1-LATENCY-ONLINE

#MODEL ESTIMATES WITH UNCERTAINTY
# 
# #setup references 
# # rep.q1t.lm1 <- lm(log(item_q1_rt) ~ condition, data = df_online)
# m <- rep.q1t.lm1
# df <- df_online 
# call <- m$call %>% as.character()
# 
# # uncertainty model visualization
# df <- df  %>%
#   data_grid(pretty_condition) %>%
#   augment(m, newdata = ., se_fit = TRUE) 
# 
# #transform log
# df$.fitted <- exp(df$.fitted)
# df$.se.fit <- exp(df$.se.fit)
# 
# df %>% 
#   ggplot(aes(y = pretty_condition, color = pretty_condition)) +
#   stat_halfeye( scale = .5,
#       aes(
#         xdist = dist_student_t(df = df.residual(m), mu = .fitted, sigma = .se.fit),
#         fill = stat(cut_cdf_qi(cdf, 
#                 .width = c(.90, .95),
#                 labels = scales::percent_format())))) +
#   scale_fill_brewer(direction = -1) + 
#   labs (title = "(ONLINE) Q1 Response Latency ~ Condition", 
#         x = "model predicted mean (seconds)", y = "Condition", fill = "Interval",
#         subtitle = paste("lm(",call[2],")"),
#         caption = "note: model log(predictions) have exponentiated to original scale") + theme(legend.position = "blank")

```

###### Diagnostics

```{r}
#| label: DIAG-Q1TIME-online

#model diagnostics
# check_model(rep.q1t.lm1, panel = TRUE)
```



###### Inference

OLS Linear Regression on Q1 response time shows that condition explains a small but statistically significant amount of variance (impasse \> control). However, the model violates the assumption of normally distributed residuals.

### TODO RESPONSE CONSISTENCY

